### 2021-07-19 chicago-spacial-covid.tex
Began report

\documentclass[10pt,letterpaper,draft]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}

\author{Trent Gerew}
\title{Reaction-diffusion spacial modeling of COVID-19 in Chicago}

\begin{document}
	\begin{align}
		S_t &= - \beta_{SA} SA - \beta_{SI} SI - \mu S, \\
		E_t &= \beta_{SA} SA - \beta_{SI} - (\sigma_A + \sigma_I) E, \\
		A_t &= \sigma_A E - M_{AR} A, \\
		AR_t &= M_{AR} A, \\
		I_t &= \sigma_I E - MI, \\
		H_t &= \gamma M I - (1 - \omega) \chi H - \omega \psi H, \\
		R_t &= (1 - \gamma) M I + (1 - \omega) \chi H, \\
		D_t &= \omega \psi H.
	\end{align}
\end{document}

### 2021-07-19 local_model.m

@@ -1,0 +1,1 @@
+[date cases d_deaths hosp c_deaths] = csvread('chicago_data.csv');

### 2021-07-19 chicago-spacial-covid.tex
Started section for ODE model

@@
 
+\newcommand{\N}{\mathcal{N}}
+\newcommand{\num}{\text{num}}
+\newcommand{\obs}{\text{obs}}
+
 \author{Trent Gerew}
@@
 \begin{document}
+
 	\begin{align}
@@
 	\end{align}
+
+	We determine the optimal parameters by minimizing the Euclidean distance $\N$ between the time series generated by the model (num) and the corresponding observed time series (obs),
+	\begin{equation}
+		\N = \sum_i \left( \left| \log(C_{\num}(t_i)) - \log(C_{\obs}(t_i)) \right|^2
+		+ \left| \log(D_{\num}(t_i)) - \log(D_{\obs}(t_i)) \right|^2 \right)
+	\end{equation}
+	where the index $i$ identifies a point in the time series.
+	The optimization tries to reproduce the number of total reported cases ($C(t) = I(t) + H(t) + R(t) + D(t)$) and the total number of deceased ($D(t)$).
+
+	We can account for a change in the parameters due to the lockdown by imposing a time dependence on the transmission rates $\beta$:
+	\begin{align}
+		\beta_{IS} (t) &= \beta_{IS} \left[ \eta_{IS} + (1 - \eta_{IS}) \frac{1 - \tanh[2(t - t_q)]}{2} \right], \\
+		\beta_{AS} (t) &= \beta_{AS} \left[ \eta_{AS} + (1 - \eta_{AS}) \frac{1 - \tanh[2(t - t_q)]}{2} \right].
+	\end{align}
+	This causes the transmission rates $\beta_{IS}$ and $\beta_{AS}$ to decrease abruptly by a factor of $\eta_{IS}$ and $\eta_{AS}$ respectively at the time $t_q$ when the lockdown was imposed.
 \end{document}

### 2021-07-19 chicago-spacial-covid.tex
Started parameter table

@@
 \documentclass[10pt,letterpaper,draft]{article}
+
+\usepackage{geometry}
+\geometry{
+	letterpaper,
+	total={170mm,257mm},
+	left=20mm,
+	top=20mm,
+}
 
@@
 	This causes the transmission rates $\beta_{IS}$ and $\beta_{AS}$ to decrease abruptly by a factor of $\eta_{IS}$ and $\eta_{AS}$ respectively at the time $t_q$ when the lockdown was imposed.
+
+	\begin{center}
+	\begin{tabular}{ c c c c c }
+		\hline
+		\hline
+			&	&	Median ($t_q = 3$)	&	Median ($t_q = 16$)	&	Initial value \\
+		\hline
+		Population	&	$N$	&	\multicolumn{3}{ c }{2,693,976} \\
+		Initial populations & $(I_0, H_0, D_0)$	&	\multicolumn{3}{ c }{$(297, 134, 6)$} \\
+		Non COVID-19 death rate [per day]	&	$\mu$	&	\multicolumn{3}{ c }{$5.79 \times 10^{-7}$} \\
+		Transmission rate, $S \rightarrow I$ [per day]	&	$\beta_{IS}$	&	$0.35 (0.28-0.41)$	&	$0.32 (0.29-0.35)$	&	$c \in U[0,1]$ \\
+		\hline
+		\hline
+	\end{tabular}
+	\end{center}
 \end{document}

### 2021-07-20 data_collection.py
Fixed spelling

@@ -1,6 +1,7 @@
 # Imports
 import pandas as pd
 import datetime
+import matplotlib.pyplot as plt
 
 # Read in csv and convert dates to DateTime format
 df = pd.read_csv("COVID-19_Daily_Cases__Deaths__and_Hospitalizations.csv", parse_dates=["Date"])
@@ -27,3 +28,7 @@ df = df.loc["2020-03-01":"2021-07-16"]
 
 # Export to a csv
 df.to_csv("chicago_data.csv")
+
+# Create plot
+#df.plot()
+#plt.show()

### 2021-07-20 draft/chicago-spatial-covid-draft.tex
organized repo and created draft

\documentclass[11pt]{article}

%imports
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}

%images and image path
\usepackage{graphicx}
\graphicspath{ {./images/} }

%custom commands
\newcommand{\N}{\mathcal{N}}
\newcommand{\num}{\text{num}}
\newcommand{\obs}{\text{obs}}
\newcommand{\D}{\mathfrak{D}}

%margins
\usepackage[letterpaper, total={6.5in, 9.5in}]{geometry}

%opening
\title{Reaction-diffusion spatial modeling of COVID-19 in Chicago}
\author{Trent Gerew}

\begin{document}

\maketitle

\begin{abstract}
	We examine an application of the Law of Mass Action to model the outbreak of COVID-19 in Chicago.
	We begin with a zero-dimension (ODE) compartmental epidemiological model consisting of Susceptible, Infected, and Removed populations (SIR model).
	We then develop a spatially distributed version of the model in the form of reaction-diffusion equations (PDE).
\end{abstract}

\section{ODE and PDE Model Setup}
We begin by explaining the ODE model which is simply obtained from the full PDE model by removing the diffusion terms in the reaction-diffusion equations.

We start with a population of \textit{susceptibles} ($S$), which may become \textit{infected} ($I$) upon the emergence of the virus within the population.
Infected individuals can interact with the susceptibles at rate $\beta$ to draw new members into the group $I$ of individuals infected by the virus.
We note that the transmission rate $\beta$ incorporates the total population size (ODE model) or the total population density (PDE model).
A fraction of the infected population recover or die at a rate $\gamma$, giving rise to the \textit{removed} ($R$) population.

These parameters specify the ODE process.
They represent processes that occur in a ``well mixed'' situation when no spatial dependence is assigned,
or processes that happen locally at every point in space for the PDE's.
We discuss parameter selection in the next section.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{sir_schematic}
	\caption{Schematic diagram of the SIR model. The dashed line denotes the interaction of the infectious population with the susceptible population that leads to infection.}
	\label{fig:schematic}
\end{figure}

The population model at the PDE level is an autonomous diffusion with a source:
\begin{align}
	S_t	&=	\Delta (\D_s \Delta S) - \beta SI,	\\
	I_t	&=	\Delta (\D_I \Delta I) + \beta SI - \gamma I,	\\
	R_t	&=	\gamma I.
\end{align}
In principle the $R$ population can have a diffusion, but since we assume this population has immunity we assign $\D_R = 0$.

At the PDE level, we have not incorporated any functional forms for directional spreading.
In principle, these terms can represent ``daily practices'' (work commutes, etc.), as well as longer temporal or spatial scales (major thoroughfare traffic, etc.).
We simply allow diffusion to perform the relevant spreading.
Arriving infected individuals are assumed to form local hotspots within the susceptible population.
This initial seeding roughly approximates long-range transportation.

\end{document}

### 2021-07-20 draft/chicago-spatial-covid-draft.tex
Include bibliography

@@
 
-At the PDE level, we have not incorporated any functional forms for directional spreading.
+As in \cite{Kevrekidis_2021}, we have not incorporated any functional forms for directional spreading at the PDE level.
 In principle, these terms can represent ``daily practices'' (work commutes, etc.), as well as longer temporal or spatial scales (major thoroughfare traffic, etc.).
@@
 
+The next step is to identify the parameters at the ODE level.
+To do this, we use a nonlinear optimization algorithm.
+We determine the optimal parameters by minimizing the Euclidean distance between the time series generated by the model ($\num$) and the corresponding observed ($\obs$) time series,
+\begin{equation}
+	\N = \sum_i \left( \left| \log(C_\num (t_i)) - \log(C_\obs (t_i)) \right|^2 + \left| \log(R_\num (t_i)) - \log(R_\obs (t_i)) \right|^2 \right)
+\end{equation}
+where the index $i$ identifies a point in the time series.
+The parameters are optimized to reproduce the reported total number of infected cases $C(t) = I(t) + R(t)$, and the total number of removed $R(t)$.
+
+\bibliographystyle{abbrv}
+\bibliography{chicago-spatial-covid-draft}
+
 \end{document}

### 2021-07-20 draft/chicago-spatial-covid-draft.tex
Update chicago-spatial-covid-draft.tex

@@
 \end{abstract}
+
+\section{Introduction and Similar Work}
 
@@
 
+Since the transition rate $\gamma$ is an ``individual'' parameter, and thus acts ``on average'', we preserve the same value at the PDE level as at the ODE level.
+The transmission rate $\beta$ is more complicated, since at the PDE level the quantities $S$, $I$, and $R$ are densities rather than populations,
+which integrate over the spatial surface to the true population of each category.
+
 \bibliographystyle{abbrv}

### 2021-07-20 data/data_collection.py
Created time series data and ode function

@@ -18,7 +18,7 @@ df = df.sort_index()
 df = df.rename(columns = {"Cases - Total": "Daily Cases", "Deaths - Total": "Daily Deaths", "Hospitalizations - Total": "Daily Hospitalizations"})
 
 # Create a cumulative deaths column
-df["Cum Deaths"] = df["Daily Deaths"].cumsum()
+df["Cumulative Deaths"] = df["Daily Deaths"].cumsum()
 
 # Drop any rows with missing values (note this ordering preserves the original cum deaths)
 df = df.dropna()
@@ -30,5 +30,5 @@ df = df.loc["2020-03-01":"2021-07-16"]
 df.to_csv("chicago_data.csv")
 
 # Create plot
-#df.plot()
-#plt.show()
+df.plot(figsize=(12,8))
+plt.show()

### 2021-07-20 data/time_series.py

@@ -1,0 +1,11 @@
+# Imports
+import pandas as pd
+import numpy as np
+
+# Read in csv and convert dates to DateTime format
+df = pd.read_csv("chicago_data.csv", parse_dates=["Date"])
+
+cases = df["Daily Cases"] + df["Daily Hospitalizations"] + df["Cumulative Deaths"]
+removed = df["Cumulative Deaths"]
+
+t_values = np.linspace(0,1,len(df.loc[:, "Date"]))

### 2021-07-20 local-model/odefcn.m

@@ -1,0 +1,4 @@
+function dydt = odefcn(t,y,b,g)
+  dydt = zeros(2,1);
+  dydt(1) = - b * y(1) * y(2);
+  dydt(2) = b * y(1) * y(2) - g * y(2);

### 2021-07-20 local_model.m

@@ -1,1 +1,0 @@
-[date cases d_deaths hosp c_deaths] = csvread('chicago_data.csv');

### 2021-07-20 draft/chicago-spatial-covid-draft.tex
Created time series data and ode function
Added parameter table

@@
 \usepackage{amssymb}
+\usepackage{footnote}
 
@@
 
+\section{Results}
+\subsection{ODE Model}
+
+\begin{savenotes}
+\begin{table}[h]
+	\centering
+	\caption{ODE parameters}
+	\label{tab:parameters}
+	\begin{tabular}{ c c c c }
+		\hline
+		\hline
+			&	&	Median	&	Initial value \\
+		\hline
+		Population	&	$N$	&	2,695,598 \\
+		Initial population	&	$I_0$	&	200 \\
+		Transmission rate, $S \rightarrow I$ [per day]	&	$\beta$\footnote{The transmission rate $\beta$ must be divided by $N$ when used in the ODE model.}	&	0.30	&	$c \in U[0,1]$ \\
+		Transition rate, $I \rightarrow R$ [per day]	&	$\gamma$	&	0.55	&	$c \in U[0.25,0.75]$ \\
+		Diffusivity, $S$ [km$^2$/day]	&	$\D_S$	&	10	&	\\
+		Diffusivity, $I$ [km$^2$/day]	&	$\D_I$	&	100	&	\\
+		\hline
+		\hline
+	\end{tabular}
+\end{table}
+\end{savenotes}
+
 \bibliographystyle{abbrv}

### 2021-07-21 data/time_series.py
Export time series for matlab

@@ -5,7 +5,18 @@ import numpy as np
 # Read in csv and convert dates to DateTime format
 df = pd.read_csv("chicago_data.csv", parse_dates=["Date"])
 
-cases = df["Daily Cases"] + df["Daily Hospitalizations"] + df["Cumulative Deaths"]
-removed = df["Cumulative Deaths"]
+# Create a new dataframe
+series = pd.DataFrame()
 
+# Create uniform time values and put them in the dataframe
 t_values = np.linspace(0,1,len(df.loc[:, "Date"]))
+series["Time"] = pd.Series(t_values)
+
+# Create cases values and put them in the dataframe
+series["Cases"] = df["Daily Cases"] + df["Daily Hospitalizations"] + df["Cumulative Deaths"]
+
+# Create removed values and put them in the dataframe
+series["Removed"] = df["Cumulative Deaths"]
+
+# Export new dataframe to a csv
+series.to_csv("time_series.csv", header=False, index=False)

### 2021-07-21 local-model/localoptim.m

@@ -1,0 +1,12 @@
+%clear command window and workspace
+clear
+clc
+
+%import the time series csv file
+series = load('time_series.csv');
+
+%redefine variables for ease
+time = series(:,1);
+cases = series(:,2);
+removed = series(:,3);
+

### 2021-07-21 local-model/odefcn.m

@@ -1,4 +1,5 @@
 function dydt = odefcn(t,y,b,g)
+  %sir system of ode's
   dydt = zeros(2,1);
   dydt(1) = - b * y(1) * y(2);
   dydt(2) = b * y(1) * y(2) - g * y(2);

### 2021-07-21 local-model/odefcn.m
Removed time dependence

@@ -1,4 +1,4 @@
-function dydt = odefcn(t,y,b,g)
+function dydt = odefcn(y,b,g)
   %sir system of ode's
   dydt = zeros(2,1);
   dydt(1) = - b * y(1) * y(2);

### 2021-07-21 local-model/odefcn.m
Added comments

@@ -1,5 +1,14 @@
-function dydt = odefcn(y,b,g)
-  %sir system of ode's
+function dydt = odefcn(y,p)
+  %ODEFCN   SIR system of ode's.
+  %   dydt = odefcn(y,p) where
+  %   y(1) = S(t), y(2) = I(t),
+  %   p(1) = beta, p(2) = gamma
+  
+  %initialize dydt array
   dydt = zeros(2,1);
-  dydt(1) = - b * y(1) * y(2);
-  dydt(2) = b * y(1) * y(2) - g * y(2);
+  
+  %dydt(1) = dSdt
+  dydt(1) = - p(1) * y(1) * y(2);
+  %dydt(2) = dIdt
+  dydt(2) = p(1) * y(1) * y(2) - p(2) * y(2);
+end

### 2021-07-21 local-model/cost.m
Create cost function

@@ -1,0 +1,15 @@
+function N = cost(num, obs)
+  %COST   Cost function for minimization.
+  %   N = cost(num, obs) where
+  %   num(1) = casesNum, num(2) = removedNum,
+  %   obs(1) = casesObs, obs(2) = removedObs
+  
+  %create new variables for ease
+  casesNum = num(1);
+  removedNum = num(2);
+  casesObs = obs(1);
+  removedObs = obs(2);
+  
+  %calculate the Euclidean norm
+  N = norm(casesNum - casesObs) + norm(removedNum - removedObs);
+end

### 2021-07-21 draft/chicago-spatial-covid-draft.tex
Added ODE dynamics

@@
 \newcommand{\D}{\mathfrak{D}}
+\newcommand{\Ro}{\mathcal{R}_0}
 
@@
 
+\subsection{ODE Dynamics}
+Here we briefly discuss some dynamics of the ODE system.
+Note that for the simply SIR model as we have here, $R$ is completely determined when $S$ and $I$ are known, so we need only consider
+\begin{align*}
+	S_t	&=	- \tilde{\beta} SI, \\
+	I_t	&=	\tilde{\beta} SI - \gamma I,
+\end{align*}
+where $\tilde{\beta} = \beta / N$.
+
+Immediately we see there are infinitely many non-isolated fixed points of the form $(S,0)$.
+Linearizing, we get
+\begin{equation*}
+	\mathbf{A}_{(S,0)} = \begin{bmatrix} 0 & - \tilde{\beta} S \\ 0 & \beta S - \gamma \end{bmatrix}
+\end{equation*}
+Thus, as seen in Figure \ref{fig:localdyn}, for $S < \frac{\gamma}{\tilde{\beta}}$ the nodes are stable, and for $S > \frac{\gamma}{\tilde{\beta}}$ they are unstable.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.25\textwidth]{local_dynamics}
+	\caption{Stream plot for the local ODE dynamics.}
+	\label{fig:localdyn}
+\end{figure}
+We can then conclude that an epidemic will occur only if $S_0 = S(t=0) > \frac{\gamma}{\tilde{\beta}}$ \cite{Wang_SIR}.
+Hence, a transcritical bifurcation occurs at $\Ro = \tilde{\beta} S_0 / \gamma$, which is known as the \textit{basic reproduction number}.
+
+Note that the SIR model can be viewed as a generalized chemical reaction
+\begin{align*}
+	S + I	&\longrightarrow	2 I, \\
+	I		&\longrightarrow	R.
+\end{align*}
+However, these two reactions are not reversible, so we are not able to write the energy-dissipation law for the SIR model \cite{Forster_2013, Wang_2020}.
+Indeed, taking $\frac{I_t}{S_t} = - 1 + \frac{\gamma}{\tilde{\beta} S}$ and solving this differential equation gives $I = - S + \frac{\gamma}{\tilde{\beta}} \ln(S) + C$.
+Thus,
+\begin{equation}
+	E(S,I) = S + I - \frac{\gamma}{\tilde{\beta}} \ln(S)
+\end{equation}
+is a conserved quantity for the system.
+
 \section{Results}
@@
 		\hline
-			&	&	Median	&	Initial value \\
+			&	&	Optimal	&	Initial value \\
 		\hline

### 2021-07-21 draft/chicago-spatial-covid-draft.tex
Updated dynamics section
Fixed parameter table

@@
 	\includegraphics[width=0.25\textwidth]{local_dynamics}
-	\caption{Stream plot for the local ODE dynamics.}
+	\caption{Stream plot for the local ODE dynamics. Note the transcritical bifurcation around $S_0 = 0.4$.}
 	\label{fig:localdyn}
@@
 We can then conclude that an epidemic will occur only if $S_0 = S(t=0) > \frac{\gamma}{\tilde{\beta}}$ \cite{Wang_SIR}.
-Hence, a transcritical bifurcation occurs at $\Ro = \tilde{\beta} S_0 / \gamma$, which is known as the \textit{basic reproduction number}.
+Hence, a transcritical bifurcation occurs at $\Ro = \tilde{\beta} S_0 / \gamma$, which is known as the \textit{basic reproduction number},.
+and is used to estimate how the epidemic developed initially.
 
@@
 		Population	&	$N$	&	2,695,598 \\
-		Initial population	&	$I_0$	&	200 \\
-		Transmission rate, $S \rightarrow I$ [per day]	&	$\beta$\footnote{The transmission rate $\beta$ must be divided by $N$ when used in the ODE model.}	&	0.30	&	$c \in U[0,1]$ \\
-		Transition rate, $I \rightarrow R$ [per day]	&	$\gamma$	&	0.55	&	$c \in U[0.25,0.75]$ \\
+		Initial population	&	$I_0$	&	 \\
+		Transmission rate, $S \rightarrow I$ [per day]	&	$\beta$\footnote{The transmission rate $\beta$ must be divided by $N$ when used in the ODE model.}	&	&	$c \in U[0,1]$ \\
+		Transition rate, $I \rightarrow R$ [per day]	&	$\gamma$	&		&	$c \in U[0.25,0.75]$ \\
 		Diffusivity, $S$ [km$^2$/day]	&	$\D_S$	&	10	&	\\

### 2021-07-21 draft/chicago-spatial-covid-draft.tex
Fixed punctuation
Include data portal reference
Removed unused command
Fixed punctuation error

@@
 \usepackage{footnote}
+\usepackage{url}
+\usepackage{natbib}
 
@@
 \begin{equation*}
-	\mathbf{A}_{(S,0)} = \begin{bmatrix} 0 & - \tilde{\beta} S \\ 0 & \beta S - \gamma \end{bmatrix}
+	\mathbf{A}_{(S,0)} = \begin{bmatrix} 0 & - \tilde{\beta} S \\ 0 & \beta S - \gamma \end{bmatrix}.
 \end{equation*}
@@
 We can then conclude that an epidemic will occur only if $S_0 = S(t=0) > \frac{\gamma}{\tilde{\beta}}$ \cite{Wang_SIR}.
-Hence, a transcritical bifurcation occurs at $\Ro = \tilde{\beta} S_0 / \gamma$, which is known as the \textit{basic reproduction number},.
-and is used to estimate how the epidemic developed initially.
+Hence, a transcritical bifurcation occurs at $\Ro = \tilde{\beta} S_0 / \gamma$, which is known as the \textit{basic reproduction number}, and is used to estimate how the epidemic developed initially.
 
@@
 \subsection{ODE Model}
+We begin by discussing the zero-dimensional model for Chicago.
+Data for the progression of the pandemic in Chicago were obtained from the Chicago Data Portal \cite{Chicago_2021}.
 

### 2021-07-21 draft/chicago-spatial-covid-draft.tex
Changed wording

@@
 
-Note that the SIR model can be viewed as a generalized chemical reaction
+Note that the ODE model can be viewed as a generalized chemical reaction
 \begin{align*}
@@
 \end{align*}
-However, these two reactions are not reversible, so we are not able to write the energy-dissipation law for the SIR model \cite{Forster_2013, Wang_2020}.
+However, these two reactions are not reversible, so we are not able to write the energy-dissipation law for the ODE model \cite{Forster_2013, Wang_2020}.
 Indeed, taking $\frac{I_t}{S_t} = - 1 + \frac{\gamma}{\tilde{\beta} S}$ and solving this differential equation gives $I = - S + \frac{\gamma}{\tilde{\beta}} \ln(S) + C$.

### 2021-07-21 local-model/odefcn.m
Include time dependence

@@ -1,6 +1,6 @@
-function dydt = odefcn(y,p)
+function dydt = odefcn(t,y,p)
   %ODEFCN   SIR system of ode's.
-  %   dydt = odefcn(y,p) where
+  %   dydt = odefcn(t,y,p) where
   %   y(1) = S(t), y(2) = I(t),
   %   p(1) = beta, p(2) = gamma
   

### 2021-07-21 data/time_series.py
Time series written directly to local model folder

@@ -19,4 +19,4 @@ series["Cases"] = df["Daily Cases"] + df["Daily Hospitalizations"] + df["Cumulat
 series["Removed"] = df["Cumulative Deaths"]
 
 # Export new dataframe to a csv
-series.to_csv("time_series.csv", header=False, index=False)
+series.to_csv("../local-model/time_series.csv", header=False, index=False)

### 2021-07-21 local-model/loadData.m

@@ -1,0 +1,11 @@
+function [time, cases, removed] = loadData()
+  %import the time series csv file
+  series = load('time_series.csv');
+
+  %redefine variables for ease
+  cases = series(:,2)';
+  removed = series(:,3)';
+
+  %create even timesteps
+  time = linspace(0,1,length(cases));
+end

### 2021-07-21 local-model/localoptim.m
Delete localoptim.m

@@ -1,12 +1,0 @@
-%clear command window and workspace
-clear
-clc
-
-%import the time series csv file
-series = load('time_series.csv');
-
-%redefine variables for ease
-time = series(:,1);
-cases = series(:,2);
-removed = series(:,3);
-

### 2021-07-21 local-model/myObj.m
Create objective function

@@ -1,0 +1,16 @@
+function obj = myObj(p)
+  
+  %predicted values
+  y0 = [100, 5];  %initial values [S_0, I_0]
+  [t,y] = ode45(@(t,y)odefc(t,x,p), time, y0);
+  
+  casesNum = y(2);
+  removedNum = y0(1)*ones(1,length(casesNum)) - y(1) - y(2);
+  
+  %measured values
+  casesObs = cases;
+  removedObs = removed;
+  
+  %compute the Euclidean norm
+  obj = norm(casesNum - casesObs) + norm(removedNum - removedObs);
+end

### 2021-07-21 local-model/sim.m
Create test simulation

@@ -1,0 +1,8 @@
+function [] = sim(p0)
+  
+  %predicted values
+  [t,y] = ode45(@(t,y)odefcn(t,y,p0), [0 498], [8413809,431]);
+  
+  %actual values
+  plot(t,y)
+  %plot(t,removed)

### 2021-07-21 local-model/sim.m
Update sim.m

@@ -4,5 +4,7 @@ function [] = sim(p0)
   [t,y] = ode45(@(t,y)odefcn(t,y,p0), [0 498], [8413809,431]);
   
   %actual values
+  hold on
   plot(t,y)
-  %plot(t,removed)
+  plot(t,removed)
+  hold off

### 2021-07-21 data/time_series.py
Select first wave of cases

@@ -4,13 +4,25 @@ import numpy as np
 
 # Read in csv and convert dates to DateTime format
 df = pd.read_csv("chicago_data.csv", parse_dates=["Date"])
+df = df.set_index(["Date"])
+
+# Select the date range
+start = '2020-03-01' # start of data collection
+end = '2020-09-30' # end of first wave
+
+# Reform dataframe to date range
+df = df.loc[start:end]
 
 # Create a new dataframe
 series = pd.DataFrame()
 
+# Set dates to be index of the dataframe
+series["Date"] = df.index
+series = series.set_index(["Date"])
+
 # Create uniform time values and put them in the dataframe
-t_values = np.linspace(0,1,len(df.loc[:, "Date"]))
-series["Time"] = pd.Series(t_values)
+t_values = np.linspace(0,1,len(df.loc[:, "Daily Cases"]))
+series["Time"] = t_values
 
 # Create cases values and put them in the dataframe
 series["Cases"] = df["Daily Cases"] + df["Daily Hospitalizations"] + df["Cumulative Deaths"]

### 2021-07-25 local-model/cost.m
Removed test ode fitting functions

@@ -1,15 +1,0 @@
-function N = cost(num, obs)
-  %COST   Cost function for minimization.
-  %   N = cost(num, obs) where
-  %   num(1) = casesNum, num(2) = removedNum,
-  %   obs(1) = casesObs, obs(2) = removedObs
-  
-  %create new variables for ease
-  casesNum = num(1);
-  removedNum = num(2);
-  casesObs = obs(1);
-  removedObs = obs(2);
-  
-  %calculate the Euclidean norm
-  N = norm(casesNum - casesObs) + norm(removedNum - removedObs);
-end

### 2021-07-25 local-model/loadData.m

@@ -1,11 +1,0 @@
-function [time, cases, removed] = loadData()
-  %import the time series csv file
-  series = load('time_series.csv');
-
-  %redefine variables for ease
-  cases = series(:,2)';
-  removed = series(:,3)';
-
-  %create even timesteps
-  time = linspace(0,1,length(cases));
-end

### 2021-07-25 local-model/myObj.m

@@ -1,16 +1,0 @@
-function obj = myObj(p)
-  
-  %predicted values
-  y0 = [100, 5];  %initial values [S_0, I_0]
-  [t,y] = ode45(@(t,y)odefc(t,x,p), time, y0);
-  
-  casesNum = y(2);
-  removedNum = y0(1)*ones(1,length(casesNum)) - y(1) - y(2);
-  
-  %measured values
-  casesObs = cases;
-  removedObs = removed;
-  
-  %compute the Euclidean norm
-  obj = norm(casesNum - casesObs) + norm(removedNum - removedObs);
-end

### 2021-07-25 local-model/mySim.m

@@ -1,22 +1,0 @@
-function [] = sim(p0)
-  
-  %load data
-  [time, cases, removed] = loadData();
-  
-  %initial conditions
-  y0 = [8413809,431]
-  
-  %predicted values
-  [t,y] = ode45(@(t,y)odefcn(t,y,p0), time, y0);
-  
-  disp(num2str(length(t)));
-  
-  % Calculate total cases and total removed
-  casesNum = y(2);
-  removedNum = y0(1)*ones(1,length(casesNum)) - y(1) - y(2);
-  
-  %plot values
-  hold on
-  plot(t,casesNum,t,removedNum)
-  hold off
-end

### 2021-07-25 local-model/odefcn.m

@@ -1,14 +1,0 @@
-function dydt = odefcn(t,y,p)
-  %ODEFCN   SIR system of ode's.
-  %   dydt = odefcn(t,y,p) where
-  %   y(1) = S(t), y(2) = I(t),
-  %   p(1) = beta, p(2) = gamma
-  
-  %initialize dydt array
-  dydt = zeros(2,1);
-  
-  %dydt(1) = dSdt
-  dydt(1) = - p(1) * y(1) * y(2);
-  %dydt(2) = dIdt
-  dydt(2) = p(1) * y(1) * y(2) - p(2) * y(2);
-end

### 2021-07-25 data/time_series.py
Set start date to March 17

@@ -7,7 +7,7 @@ df = pd.read_csv("chicago_data.csv", parse_dates=["Date"])
 df = df.set_index(["Date"])
 
 # Select the date range
-start = '2020-03-01' # start of data collection
+start = '2020-03-17' # start of data collection
 end = '2020-09-30' # end of first wave
 
 # Reform dataframe to date range
@@ -19,10 +19,6 @@ series = pd.DataFrame()
 # Set dates to be index of the dataframe
 series["Date"] = df.index
 series = series.set_index(["Date"])
-
-# Create uniform time values and put them in the dataframe
-t_values = np.linspace(0,1,len(df.loc[:, "Daily Cases"]))
-series["Time"] = t_values
 
 # Create cases values and put them in the dataframe
 series["Cases"] = df["Daily Cases"] + df["Daily Hospitalizations"] + df["Cumulative Deaths"]

### 2021-07-25 local-model/odeModel.m
Create odeModel.m

@@ -1,0 +1,148 @@
+% Clear windows and workspace
+clear
+clc
+close all
+
+%% Problem Definition
+% Population
+n = 2695598; %total population
+
+% Initial Conditions
+i0 = 127; %infected
+r0 = 2; %recovered
+s0 = n - i0 - r0; %susceptible
+y0 = [s0, i0, r0]; %initial conditions vector
+
+% Time domain
+%t = 1:61; %60 days
+
+%% Observed Data
+% Load time, total cases and total deaths
+[t,Cobs,Dobs] = loadData();
+
+% Convert to a single vector
+Yobs = [Cobs(:); Dobs(:)];
+
+%% Test Data (with noise)
+% % Parameters
+% p = [0.3/n, 0.3]; %[beta, gamma]
+% 
+% % Compute simulation
+% [~,Y] = ode45(@(t,y)sir(t,y,p),t,y0);
+% I = Y(:,2);
+% R = Y(:,3);
+% 
+% % Define total cases and deaths, and add noise
+% noise = 1000*sawtooth(t)';
+% Cobs = I + R + noise;
+% Dobs = R + noise;
+% 
+% % Convert to a single vector
+% Yobs = [Cobs(:); Dobs(:)];
+
+%% Model Solution
+% Define optimization iterations
+iters = 2000;
+
+% Parameter samples
+beta = rndSample([0,1],iters);
+gamma = rndSample([0.25,0.75],iters);
+
+% Initialize estimate vectors
+betaEst = zeros(1,iters);
+gammaEst = zeros(1,iters);
+
+% Sample each initial condition
+for i=1:iters
+    p0 = [beta(i)/n,gamma(i)];
+    % Generate parameter estimate
+    pEst = fminsearch(@(p)objfun(t,Yobs,p,y0),p0);
+    betaEst(i) = pEst(1);
+    gammaEst(i) = pEst(2);
+end
+
+% Calculate median values
+betaMed = median(betaEst);
+gammaMed = median(gammaEst);
+
+% Calculate IQR
+betaIQR = iqr(betaEst);
+gammaIQR = iqr(gammaEst);
+
+%% Present Results
+% Solve with median estimated parameters
+[~,Y] = ode45(@(t,y)sir(t,y,[betaMed,gammaMed]),t,y0);
+
+% Convert to SIR
+I = Y(:,2);
+R = Y(:,3);
+
+% Define total cases and deaths
+Cnum = I + R;
+Dnum = R;
+
+% Print parameters
[…]

### 2021-07-25 draft/chicago-spatial-covid-draft.tex
Include updated initial conditions and optimal parameters

@@
 		\hline
-			&	&	Optimal	&	Initial value \\
+			&	&	Median (interquartile range)	&	Initial value \\
 		\hline
 		Population	&	$N$	&	2,695,598 \\
-		Initial population	&	$I_0$	&	 \\
-		Transmission rate, $S \rightarrow I$ [per day]	&	$\beta$\footnote{The transmission rate $\beta$ must be divided by $N$ when used in the ODE model.}	&	&	$c \in U[0,1]$ \\
-		Transition rate, $I \rightarrow R$ [per day]	&	$\gamma$	&		&	$c \in U[0.25,0.75]$ \\
+		Initial population	&	$(I_0, R_0)$	&	(127, 2) \\
+		Transmission rate, $S \rightarrow I$ [per day]	&	$\beta$\footnote{The transmission rate $\beta$ must be divided by $N$ when used in the ODE model.}	&	0.38206(0.38204-0.38209)	&	$c \in U[0,1]$ \\
+		Transition rate, $I \rightarrow R$ [per day]	&	$\gamma$	&	 0.39656(0.39654-0.39659)	&	$c \in U[0.25,0.75]$ \\
 		Diffusivity, $S$ [km$^2$/day]	&	$\D_S$	&	10	&	\\

### 2021-07-25 draft/chicago-spatial-covid-draft.tex
Updated ODE model figures

@@
 \usepackage{graphicx}
+\usepackage{subfig}
 \graphicspath{ {./images/} }
@@
 We determine the optimal parameters by minimizing the Euclidean distance between the time series generated by the model ($\num$) and the corresponding observed ($\obs$) time series,
-\begin{equation}
-	\N = \sum_i \left( \left| \log(C_\num (t_i)) - \log(C_\obs (t_i)) \right|^2 + \left| \log(R_\num (t_i)) - \log(R_\obs (t_i)) \right|^2 \right)
+\begin{equation} \label{eq:obj}
+	\N = \sum_i^{t_\text{fit}^\text{end}} \left( \left| \log(C_\num (t_i)) - \log(C_\obs (t_i)) \right|^2 + \left| \log(R_\num (t_i)) - \log(R_\obs (t_i)) \right|^2 \right)
 \end{equation}
@@
 Data for the progression of the pandemic in Chicago were obtained from the Chicago Data Portal \cite{Chicago_2021}.
+The relevant results are given in Figure \ref{ode_results}.
+We obtained the optimal ODE model parameters for Chicago from 1,000 optimizations that compared model predictions to jointly the number of cumulative infected and the deceases, as in Equation \ref{eq:obj}.
 
@@
 
+\begin{figure}
+	\centering
+	\subfloat{\includegraphics[width=.475\textwidth]{ode_cases}}
+	\subfloat{\includegraphics[width=.475\textwidth]{ode_deaths}} \\
+	\subfloat{\includegraphics[width=.475\textwidth]{ode_IR}}
+	\caption{ODE model for Chicago with fitting to official data from March 17, 2020 ($t = t_\text{init} = 1$) to September 30, 2020 ($t = t_\text{fit}^\text{end} = 198$).
+		Left panel: Confirmed cases $C(t) = I(t) + R(t)$; Right panel: Number of deaths $D(t)$.
+		Bottom panel shows the other populations: infected $I(t)$ and recovered $R(t)$.}
+	\label{fig:ode_results}
+\end{figure}
+
 \bibliographystyle{abbrv}

### 2021-07-25 draft/chicago-spatial-covid-draft.tex
Fixed spelling and reference errors

@@
 Data for the progression of the pandemic in Chicago were obtained from the Chicago Data Portal \cite{Chicago_2021}.
-The relevant results are given in Figure \ref{ode_results}.
-We obtained the optimal ODE model parameters for Chicago from 1,000 optimizations that compared model predictions to jointly the number of cumulative infected and the deceases, as in Equation \ref{eq:obj}.
+The relevant results are given in Figure \ref{fig:ode_results}.
+We obtained the optimal ODE model parameters for Chicago from 1,000 optimizations that compared model predictions to jointly the number of cumulative infected and the deceased, as in Equation \ref{eq:obj}.
 

### 2021-07-25 draft/chicago-spatial-covid-draft.tex
Update chicago-spatial-covid-draft.tex

@@
 The relevant results are given in Figure \ref{fig:ode_results}.
-We obtained the optimal ODE model parameters for Chicago from 1,000 optimizations that compared model predictions to jointly the number of cumulative infected and the deceased, as in Equation \ref{eq:obj}.
+We obtained the optimal ODE model parameters for Chicago from 1,000 optimizations (implemented in MATLAB via the \verb|fminsearch| function)
+that compared model predictions to jointly the number of cumulative infected and the deceased, as in Equation \ref{eq:obj}.
+For each optimization, the initial guess for each parameter was uniformly sampled within a pre-specified range.
+The parameter ranges were determined from epidemiological information.
+
+Upon performing the optimizations we find that the fitting yields the results summarized in Table \ref{tab:parameters}.
+We show the median parameters, as well as the interquartile range, and the range of variation used to sample the parameters.
+Model predictions (with median parameter values, solid blue lines) are compared graphically to data (black dots) in Figure \ref{fig:ode_results}.
 

### 2021-07-26 draft/chicago-spatial-covid-draft.tex
Moved ODE dynamics to appendix
Update chicago-spatial-covid-draft.tex

@@
 
-\subsection{ODE Dynamics}
+\section{Results}
+\subsection{ODE Model}
+We begin by discussing the zero-dimensional model for Chicago.
+Data for the progression of the pandemic in Chicago were obtained from the Chicago Data Portal \cite{Chicago_2021}.
+The relevant results are given in Figure \ref{fig:ode_results}.
+We obtained the optimal ODE model parameters for Chicago from 1,000 optimizations (implemented in MATLAB via the \verb|fminsearch| function)
+that compared model predictions to jointly the number of cumulative infected and the deceased, as in Equation \ref{eq:obj}.
+For each optimization, the initial guess for each parameter was uniformly sampled within a pre-specified range.
+The parameter ranges were determined from epidemiological information.
+
+Upon performing the optimizations we find that the fitting yields the results summarized in Table \ref{tab:parameters}.
+We show the median parameters, as well as the interquartile range, and the range of variation used to sample the parameters.
+Model predictions (with median parameter values, solid blue lines) are compared graphically to data (black dots) in Figure \ref{fig:ode_results}.
+
+\begin{savenotes}
+\begin{table}[h]
+	\centering
+	\caption{ODE parameters for Chicago: optimal (best-fitting), median and interquartile range, and variation range used in the optimization algorithm.
+Initial parameter guesses were uniformly sampled within these ranges.}
+	\label{tab:parameters}
+	\begin{tabular}{ c c c c }
+		\hline
+		\hline
+			&	&	Median (interquartile range)	&	Initial value \\
+		\hline
+		Population	&	$N$	&	2,695,598	& \\
+		Initial population	&	$(I_0, R_0)$	&	(127, 2)	& \\
+		Transmission rate, $S \rightarrow I$ [per day]	&	$\beta$\footnote{The transmission rate $\beta$ must be divided by $N$ when used in the ODE model.}	&	0.38206(0.38204-0.38209)	&	$c \in U[0,1]$ \\
+		Transition rate, $I \rightarrow R$ [per day]	&	$\gamma$	&	 0.39656(0.39654-0.39659)	&	$c \in U[0.25,0.75]$ \\
+		Diffusivity, $S$ [km$^2$/day]	&	$\D_S$	&	10	&	\\
+		Diffusivity, $I$ [km$^2$/day]	&	$\D_I$	&	100	&	\\
+		\hline
+		\hline
+	\end{tabular}
+\end{table}
+\end{savenotes}
+
+\begin{figure}
+	\centering
+	\subfloat{\includegraphics[width=.475\textwidth]{ode_cases}}
+	\subfloat{\includegraphics[width=.475\textwidth]{ode_deaths}} \\
+	\subfloat{\includegraphics[width=.475\textwidth]{ode_IR}}
+	\caption{ODE model for Chicago with fitting to official data from March 17, 2020 ($t = t_\text{init} = 1$) to September 30, 2020 ($t = t_\text{fit}^\text{end} = 198$).
+		Left panel: Confirmed cases $C(t) = I(t) + R(t)$; Right panel: Number of deaths $D(t)$.
+		Bottom panel shows the other populations: infected $I(t)$ and recovered $R(t)$.}
+	\label{fig:ode_results}
+\end{figure}
+
+\appendix
+\section{ODE Dynamics}
 Here we briefly discuss some dynamics of the ODE system.
@@
 
-\section{Results}
-\subsection{ODE Model}
-We begin by discussing the zero-dimensional model for Chicago.
-Data for the progression of the pandemic in Chicago were obtained from the Chicago Data Portal \cite{Chicago_2021}.
-The relevant results are given in Figure \ref{fig:ode_results}.
-We obtained the optimal ODE model parameters for Chicago from 1,000 optimizations (implemented in MATLAB via the \verb|fminsearch| function)
-that compared model predictions to jointly the number of cumulative infected and the deceased, as in Equation \ref{eq:obj}.
-For each optimization, the initial guess for each parameter was uniformly sampled within a pre-specified range.
-The parameter ranges were determined from epidemiological information.
-
-Upon performing the optimizations we find that the fitting yields the results summarized in Table \ref{tab:parameters}.
-We show the median parameters, as well as the interquartile range, and the range of variation used to sample the parameters.
-Model predictions (with median parameter values, solid blue lines) are compared graphically to data (black dots) in Figure \ref{fig:ode_results}.
-
-\begin{savenotes}
-\begin{table}[h]
-	\centering
-	\caption{ODE parameters}
-	\label{tab:parameters}
-	\begin{tabular}{ c c c c }
-		\hline
-		\hline
-			&	&	Median (interquartile range)	&	Initial value \\
-		\hline
-		Population	&	$N$	&	2,695,598 \\
-		Initial population	&	$(I_0, R_0)$	&	(127, 2) \\
-		Transmission rate, $S \rightarrow I$ [per day]	&	$\beta$\footnote{The transmission rate $\beta$ must be divided by $N$ when used in the ODE model.}	&	0.38206(0.38204-0.38209)	&	$c \in U[0,1]$ \\
-		Transition rate, $I \rightarrow R$ [per day]	&	$\gamma$	&	 0.39656(0.39654-0.39659)	&	$c \in U[0.25,0.75]$ \\
-		Diffusivity, $S$ [km$^2$/day]	&	$\D_S$	&	10	&	\\
-		Diffusivity, $I$ [km$^2$/day]	&	$\D_I$	&	100	&	\\
-		\hline
-		\hline
-	\end{tabular}
-\end{table}
-\end{savenotes}
-
-\begin{figure}
-	\centering
-	\subfloat{\includegraphics[width=.475\textwidth]{ode_cases}}
-	\subfloat{\includegraphics[width=.475\textwidth]{ode_deaths}} \\
-	\subfloat{\includegraphics[width=.475\textwidth]{ode_IR}}
-	\caption{ODE model for Chicago with fitting to official data from March 17, 2020 ($t = t_\text{init} = 1$) to September 30, 2020 ($t = t_\text{fit}^\text{end} = 198$).
-		Left panel: Confirmed cases $C(t) = I(t) + R(t)$; Right panel: Number of deaths $D(t)$.
-		Bottom panel shows the other populations: infected $I(t)$ and recovered $R(t)$.}
-	\label{fig:ode_results}
-\end{figure}
-
 \bibliographystyle{abbrv}

### 2021-07-27 draft/chicago-spatial-covid-draft.tex
Include figure and explanation of 3 waves

@@
 
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=.6\textwidth]{obs_data}
+	\caption{Observed data for Chicago showing the three impact waves.}
+	\label{fig:obs_data}
+\end{figure}
+
+The COVID-19 pandemic had three major impact waves in Chicago, as seen clearly in Figure \ref{fig:obs_data}.
+For simplicity, we analyze only the first major wave, which ends at approximately October 30, 2020.
+Additionally, we begin the analysis at March 17 instead of the first day of data collection (March 1) because of reporting errors early in the pandemic.
+
 Upon performing the optimizations we find that the fitting yields the results summarized in Table \ref{tab:parameters}.

### 2021-07-27 data/data_collection.py

@@ -3,6 +3,11 @@ import pandas as pd
 import datetime
 import matplotlib.pyplot as plt
 
+# LaTeX font for matplotlib
+import matplotlib
+matplotlib.rcParams['mathtext.fontset'] = 'stix'
+matplotlib.rcParams['font.family'] = 'STIXGeneral'
+
 # Read in csv and convert dates to DateTime format
 df = pd.read_csv("COVID-19_Daily_Cases__Deaths__and_Hospitalizations.csv", parse_dates=["Date"])
 # Use only the totals (not demographics)

### 2021-07-27 draft/chicago-spatial-covid-draft.tex
Expand transmission rate discussion

@@
 
+The optimal parameters were determined for a scenario in which restrictive measures (quarantine, lockdown) were not enforced.
+Therefore, we do not impose a time dependence on the transmission rate $\beta$.
+The transmission rate represents the rate of contact of susceptible individuals with infectious individuals that leads to infection by the virus.
+It may be expressed as the product of daily average contacts and the probability of infection given a contact \cite{Drossinos_2020}.
+Thus, restrictive measures such as mobility restrictions, social-distancing, face-mask wearing, and gathering limitations are expected to decrease the transmission rate, an effect not accounted for in this model.
+
 \begin{savenotes}
@@
 \end{savenotes}
+
+The two top panels of Figure \ref{fig:ode_results} show how well we capture the data for the Chicago outbreak.
+The model reproduces reasonable well the number of cases (top left panel) but not the number of deaths (top right panel).
+The model fails to capture the characteristic S-shape in the logarithmic plot of the top right panel (deaths).
 
@@
 We can then conclude that an epidemic will occur only if $S_0 = S(t=0) > \frac{\gamma}{\tilde{\beta}}$ \cite{Wang_SIR}.
-Hence, a transcritical bifurcation occurs at $\Ro = \tilde{\beta} S_0 / \gamma$, which is known as the \textit{basic reproduction number}, and is used to estimate how the epidemic developed initially.
+Hence, a transcritical bifurcation occurs at $\Ro = \tilde{\beta} S_0 / \gamma$, which is known as the \textit{basic reproduction number}, defined as the average number of secondary infections caused by a single infected individual \cite{Drossinos_2020}.
 

### 2021-07-27 draft/chicago-spatial-covid-draft.tex
Fixed authors
Update chicago-spatial-covid-draft.tex

@@
 
+%insitute command
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\institute}[1]{% add institute to \maketitle
+	\apptocmd{\@author}{\end{tabular}
+	\par
+	\begin{tabular}[t]{c}
+		\small \textit{#1}}{}{}
+}
+\makeatother
+
 %margins
@@
 \title{Reaction-diffusion spatial modeling of COVID-19 in Chicago}
-\author{Trent Gerew}
+\author{Trent Gerew\thanks{\texttt{tgerew@hawk.iit.edu}} \and 
+	Chun Liu\thanks{\texttt{cliu124@iit.edu}} \and 
+	Yiwei Wang\thanks{\texttt{ywang487@iit.edu}}}
+\institute{Department of Applied Mathematics, Illinois Institute of Technology, Chicago, Illinois}
 
@@
 \begin{equation} \label{eq:obj}
-	\N = \sum_i^{t_\text{fit}^\text{end}} \left( \left| \log(C_\num (t_i)) - \log(C_\obs (t_i)) \right|^2 + \left| \log(R_\num (t_i)) - \log(R_\obs (t_i)) \right|^2 \right)
+	\N = \sum_i^{t_\text{fit}^\text{end}} \left( \left| \log(C_\num (t_i)) - \log(C_\obs (t_i)) \right|^2 + \left| \log(D_\num (t_i)) - \log(D_\obs (t_i)) \right|^2 \right)
 \end{equation}
 where the index $i$ identifies a point in the time series.
-The parameters are optimized to reproduce the reported total number of infected cases $C(t) = I(t) + R(t)$, and the total number of removed $R(t)$.
+The parameters are optimized to reproduce the reported total number of infected cases $C(t) = I(t) + R(t)$, and the total number of deceased $D(t)$.
 

### 2021-07-27 draft/chicago-spatial-covid-draft.tex
Update chicago-spatial-covid-draft.tex

@@
 where the index $i$ identifies a point in the time series.
-The parameters are optimized to reproduce the reported total number of infected cases $C(t) = I(t) + R(t)$, and the total number of deceased $D(t)$.
+The parameters are optimized to reproduce the reported total number of infected cases $C(t) = I(t) + R(t)$, and the total number of deceased $D(t) = R(t)$.
 

### 2021-08-23 pres-8-27/slides/background.tex
Initial SoReMo presentation

\begin{slide}{Background}
The study of infectious diseases began with Bernoulli in 1760. \\
The study of a critical level of transmissibility for a disease is fundamental in epidemiology. \\
This is useful for
\begin{itemize}
	\item predicting counterintuitive effects of treatments, and
	\item considering containment measures
\end{itemize}
\vfill
The main flavors of models fit under deterministic/stochastic and discrete/continuous. \\

Here we consider a deterministic and continuous compartment model (Kermack-McKendrick model). \\
\begin{columns}[T]
\begin{column}{.6\textwidth}
	\begin{block}{Assumptions}
		\begin{itemize}
			\item Birth rate balances the mortality rate.
			\item No immunity in early stages of epidemic.
			\item Removal provides complete immunity against reinfection.
		\end{itemize}
	\end{block}
\end{column}
\end{columns}

\end{slide}

### 2021-08-23 pres-8-27/slides/basic_r_num.tex
Initial SoReMo presentation

\begin{slide}{Basic Reproduction Number}

Consider the disease-free equilibrium: $S=N$, $I=R=0$.\\
\vfill
Then the Jacobian matrix of the infection subsystem is
$$\mathbf{J_0}	= \begin{bmatrix} \frac{\partial \dot{I}}{\partial I} \end{bmatrix}
			= \begin{bmatrix} - \gamma + \frac{\beta_I S I}{N^2} + \frac{\beta_I S}{N} \end{bmatrix}_{S=N , I = R = 0}
			= - \gamma + \beta_I $$

$\mathbf{J_0}$ can be decomposed into transmission ($\mathbf{T} = \beta_I$) and transition ($\mathbf{\Sigma} = - \gamma$) parts such that $\mathbf{J_0} = \mathbf{T} + \mathbf{\Sigma}$. \\
\vfill
The next-generation matrix is
$$\mathbf{K_L} = - \mathbf{T} \mathbf{\Sigma}^{-1} = \frac{\beta_I}{\gamma}$$
and the basic reproduction number is
$$\mathcal{R}_0 = \rho ( \mathbf{K_L} ) = \frac{\beta_I}{\gamma}$$

\end{slide}

### 2021-08-23 pres-8-27/slides/chicago_dem.tex
Initial SoReMo presentation

\begin{slide}{Demographics and Death Rates}
The City of Chicago Data Portal provides COVID data in demographic categories for age, race, and sex. \\
We begin by fitting $\alpha$ for each category independently, where
\begin{itemize}
	\item $\delta$, $\gamma$, and $\rho$ are constants determined medical literature, and
	\item $R_0$ is determined from the Rt Live model.
\end{itemize}

\begin{columns}[T]
	\begin{column}{.4\textwidth}
		\vspace{0.5in}
		The average death rate of Chicago is then calculated by a weighted mean:
		$$\alpha = \frac{1}{\sum N_i} \sum \alpha_i N_i$$
		\vfill
	\end{column}
	\begin{column}{.5\textwidth}
		\begin{figure}[h]
			\centering
			\includegraphics[height=1.5in]{images/chicago_model}
			\caption{Final model trained on the entire data set and predicted 10 days out (from July 16).}
		\end{figure}
	\end{column}
\end{columns}

\end{slide}

### 2021-08-23 pres-8-27/slides/chicago_r_num.tex
Initial SoReMo presentation

\begin{slide}{Basic Reproduction Number}
\vfill
The website \href{https://rt.live/}{Rt Live} tracked $R_0$ since the beginning of the COVID-19 pandemic using data from \href{https://covidtracking.com/?source=rt}{The COVID Tracking Project}.
\begin{columns}[T]
	\begin{column}{.5\textwidth}
		\begin{figure}[h]
			\centering
			\includegraphics[height=1.25in]{images/chicago_r_num}	
		\end{figure}
	\end{column}
	\begin{column}{.5\textwidth}
		\begin{block}{Model (\cite{rtlive2020})}
			Search for the most likely curve that produced the new cases per day that is observed:
			\begin{itemize}
				\item Assume a seed number of people and a $R_0$ curve,
				\item Distribute these cases into the future using a known delay distribution,
				\item Scale and add noise.
			\end{itemize}
		\end{block}
	\end{column}
\end{columns}
\vfill
Based the calculation for Illinois, we assume a logistic form for $R_0$ over time:
$$R_0 (t) = \frac{R_{\text{start}} - R_{\text{end}}}{1 + \exp ( - k (x_0 - t))} + R_{\text{end}}$$
\vfill
\end{slide}

### 2021-08-23 pres-8-27/slides/local_dynamics.tex
Initial SoReMo presentation

\begin{slide}{Dynamics}

\begin{columns}[T]
	\begin{column}{.7\textwidth}
		Let $k = \frac{\beta_I}{N}$, then $\dot{S} = -k S I$ and $\dot{I} = k S I - \gamma I$

		There are infintely non-isolated fixed points of the form $(a,0)$.
		Linearizing, we get
		$$\mathbf{A_{(a,0)}} = \begin{bmatrix} 0 & -k a \\ 0 & k a - \gamma \end{bmatrix}$$
		so for $a < \frac{\gamma}{k}$ the nodes are stable, and they are unstable for $a > \frac{\gamma}{k}$.
	\end{column}
	\begin{column}{.25\textwidth}
		\centering
		\includegraphics[height=1.5in]{images/local}
	\end{column}
\end{columns}
Taking $\frac{\dot{I}}{\dot{S}} = -1 + \frac{\gamma}{k S}$ and solving this differential equation gives $I = - S + \frac{\gamma}{k} \ln (S) + C$.
Thus
$$E(S,I) = S + I - \frac{\gamma}{k} \ln (S)$$
is a conserved quantity for the system.
So there cannot be any attracting fixed points (\cite{strogatz:2000}).
\end{slide}

### 2021-08-23 pres-8-27/slides/mobility.tex
Initial SoReMo presentation

\begin{slide}{Mobile Contact Probability}
If reliable mobility data exists, we can explicitly define the contact probabilities $C_{ij}^X$.
\vfill
\begin{columns}[T]
\begin{column}{.7\textwidth}
	\begin{definition}[\cite{Gatto10484}]
		$$ C_{ij}^X =
		\begin{cases}
			(1 - p_i) + (1 - r_X) p_i + r_X p_i q_{ij}	&	\text{if } i = j, \\
			r_X p_i q_{ij}					&	\text{otherwise}
		\end{cases} $$
	\end{definition}
\end{column}
\end{columns}
\vfill
\begin{itemize}
	\item $p_i$ is the fraction of mobile people in $i$
	\item $q_{ij}$ is the fraction of mobile people beween $i$ and $j=1, \dots , n$
	\item $r_X (0 \leq r_X \leq 1)$ is the fraction of contacts occuring when people of $X$ travel
\end{itemize}
\vfill
Mobility data can be often be found in the form of Origin-Destination matrices.
\begin{itemize}
	\item The \href{https://www.bts.gov/topics/airlines-and-airports/origin-and-destination-survey-data}{US DoT} and \href{https://libguides.mit.edu/gis}{MIT}
	provide such data on a national scale.
	\item Within Chicago, we can access tranportation datasets provided by the \href{https://data.cityofchicago.org/browse?category=Transportation&page=1}{Chicago Data Portal}.
\end{itemize}

\end{slide}

### 2021-08-23 pres-8-27/slides/model_contact.tex
Initial SoReMo presentation

--- 2021-08-23 pres-8-27/slides/mobility.tex
@@
 \begin{slide}{Mobile Contact Probability}
-If reliable mobility data exists, we can explicitly define the contact probabilities $C_{ij}^X$.
-\vfill
+Using one of the definitions of $\pi_{ij}$ we can define the contact probabilities.
 \begin{columns}[T]
 \begin{column}{.7\textwidth}
-	\begin{definition}[\cite{Gatto10484}]
+	\begin{definition}
 		$$ C_{ij}^X =
 		\begin{cases}
-			(1 - p_i) + (1 - r_X) p_i + r_X p_i q_{ij}	&	\text{if } i = j, \\
-			r_X p_i q_{ij}					&	\text{otherwise}
+			(1 - \sum_{j=1}^n \pi_{ij}) + (1 - r_X)( \sum_{j=1}^n \pi_{ij} ) + r_X \pi_{ij} &	\text{if } i = j, \\
+			r_X \pi_{ij}				&	\text{otherwise}
 		\end{cases} $$
@@
 \begin{itemize}
-	\item $p_i$ is the fraction of mobile people in $i$
-	\item $q_{ij}$ is the fraction of mobile people beween $i$ and $j=1, \dots , n$
 	\item $r_X (0 \leq r_X \leq 1)$ is the fraction of contacts occuring when people of $X$ travel
 \end{itemize}
-\vfill
-Mobility data can be often be found in the form of Origin-Destination matrices.
-\begin{itemize}
-	\item The \href{https://www.bts.gov/topics/airlines-and-airports/origin-and-destination-survey-data}{US DoT} and \href{https://libguides.mit.edu/gis}{MIT}
-	provide such data on a national scale.
-	\item Within Chicago, we can access tranportation datasets provided by the \href{https://data.cityofchicago.org/browse?category=Transportation&page=1}{Chicago Data Portal}.
-\end{itemize}
-
 \end{slide}

### 2021-08-23 pres-8-27/slides/spatial_model.tex
Initial SoReMo presentation

\begin{slide}{Spatial Model}

Now suppose we have a set of $n$ connected communities.
Assume $\gamma$ is community-independent, while $\lambda_i$ is the force of infection for community $i$.
\begin{columns}[T]
	\begin{column}{.5\textwidth}
		\begin{align*}
			\dot{S_i}	&=	-\lambda_i S_i \\
			\dot{I_i}	&=	\lambda_i S_i - \gamma I_i \\
			\dot{R_i}	&=	\gamma I_i
		\end{align*}
		\begin{definition}[\cite{Gatto10484}]
			$$\lambda_i = \sum_{j=1}^n C_{ij}^S \frac{ \sum_{k=1}^n ( \beta_I C_{kj}^I I_k ) }{ \sum_{k=1}^n ( C_{kj}^S S_k + C_{kj}^I I_k + C_{kj}^R R_k ) }$$
		\end{definition}
	\end{column}
	\begin{column}{.4\textwidth}
		\centering
		\includegraphics[height=1.75in]{images/map}	
	\end{column}	
\end{columns}

\vfill

$C_{ij}^X$ (with $X \in \{S,I,R\}$) is the probability that individuals who belong to compartment $X$ and are from community $i$ contact individuals in community $j$.

\end{slide}

### 2021-08-24 pres-8-27/slides/covid-01.tex
SoReMo updates

\begin{slide}{The COVID Problem}
	Identified in Wuhan, China in December 2019. \\

	Caused by the severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2).

	\vspace{1cm}

	In the United States alone, we currently have
	\begin{itemize}
		\item \textbf{37,768,911} total cases
		\item \textbf{626,833} total deaths
	\end{itemize}
\end{slide}

### 2021-08-24 pres-8-27/slides/covid-03.tex
SoReMo updates

\begin{slide}{The COVID Problem}
	\centering
	\begin{columns}[T]
		\begin{column}{.5\textwidth}
			\centering
			COVID-19 Weekly Cases per 100,000 Population by Race/Ethnicity, United States
			\includegraphics[height=4.5cm]{images/ethnicity-trends}
		\end{column}
		\begin{column}{.5\textwidth}
			\centering
			COVID-19 Weekly Case Rate per 100,000 Population by Percentage of County Population in Poverty, United States
			\includegraphics[height=3.7cm]{images/poverty-trends}
		\end{column}
	\end{columns}
\end{slide}

### 2021-08-25 pres-8-27/slides/covid-03.tex
SoReMo updates

@@
 			\centering
-			COVID-19 Weekly Cases per 100,000 Population by Race/Ethnicity, United States
+			COVID-19 weekly cases per 100,000 population by race/ethnicity, United States
 			\includegraphics[height=4.5cm]{images/ethnicity-trends}
@@
 			\centering
-			COVID-19 Weekly Case Rate per 100,000 Population by Percentage of County Population in Poverty, United States
+			COVID-19 weekly case rate per 100,000 population by percentage of county population in poverty, United States
 			\includegraphics[height=3.7cm]{images/poverty-trends}

### 2021-08-25 pres-8-27/slides/obj-04.tex
SoReMo updates

\begin{slide}{Project Context and Objectives}
	{\large \textbf{Idea:} Use reaction-diffusion to build explicit spatial dependence.}
	\begin{columns}[T]
		\begin{column}{.6\textwidth}
			\vspace{2cm}
			\begin{itemize}
				\item The data clearly shows a diffusive pattern
				\item Spatial dependence can approximate demographic differences
				\item Spatially dependent data exists for many scales and regions
			\end{itemize}
		\end{column}
		\begin{column}{.4\textwidth}
			\includegraphics[height=6cm]{images/spatial-grid}
		\end{column}
	\end{columns}
\end{slide}

### 2021-08-26 pres-8-27/slides/work-01.tex
SoReMo updates

\begin{slide}{Current Work}
	Working over the summer with Prof. Chun Liu and Prof. Yiwei Wang. \\

	\begin{enumerate}
		\item Start with a population of \emph{susceptibles} ($S$)
		\item Some of the population may become \emph{infected} ($I$) upon emergence of the virus
		\item Infected individuals interact with susceptibles at rate $\beta$ to draw new members into $I$
	\end{enumerate}
\end{slide}

### 2021-08-26 pres-8-27/slides/work-01.tex
SoReMo updates

@@
 \begin{slide}{Current Work}
-	Working over the summer with Prof. Chun Liu and Prof. Yiwei Wang. \\
+	Original motivation from Summer 2020 IPRO with Prof. Robert Ellis and Mr. David Eads (NPR). \\
+	Expanded over Summer 2021 with Prof. Chun Liu and Prof. Yiwei Wang. \\
 
+	\vspace{.5cm}
+
+	{\large Start by defining the zero-dimension model:}
 	\begin{enumerate}
@@
 		\item Infected individuals interact with susceptibles at rate $\beta$ to draw new members into $I$
+		\item A fraction of $I$ recover or die at a rate $\gamma$ moving to the \emph{removed} ($R$) population
 	\end{enumerate}
+	\begin{figure}
+		\centering
+		\includegraphics[height=3cm]{images/sir_schematic}
+	\end{figure}
+	This represents processes that occur in a ``well mixed'' situation.
 \end{slide}

### 2021-08-26 pres-8-27/slides/work-02.tex
SoReMo updates

\begin{slide}{Current Work}
	The population model at the PDE level is an autonomous diffusion with a source:
	\begin{align*}
	S_t	&=	\nabla (\D_s \nabla S) - \beta SI,	\\
	I_t	&=	\nabla (\D_I \nabla I) + \beta SI - \gamma I,	\\
	R_t	&=	\gamma I.
	\end{align*}

	\vspace{.5cm}
	Assumptions:
	\begin{itemize}
		\item The $R$ population has immunity, so we assign $\D_R = 0$.
		\item Diffusion performs all relevant spreading.
		\item Arriving infected individuals form local hotspots.
	\end{itemize}

\end{slide}

### 2021-08-26 pres-8-27/slides/work-03.tex
SoReMo updates

\begin{slide}{Current Work}
	{\large Next we identify the parameters $\beta$ and $\gamma$ at the ODE level:} \\

	\vspace{.5cm}

	\textbf{Idea:} Minimize the distance between the model time series and the observed time series.
	\begin{equation*}
	\N = \sum_i^{t_\text{fit}^\text{end}} \left( \left| C_\num (t_i) - C_\obs (t_i) \right|^2 + \left| D_\num (t_i) - D_\obs (t_i) \right|^2 \right)
	\end{equation*}
	where the index $i$ identifies a point in the time series.\\

	\vspace{.5cm}

	Optimize the parameters to reproduce:
	\begin{itemize}
		\item the reported \emph{total} number of infected cases $C(t) = I(t) + R(t)$
		\item the \emph{total} number of deceased $D(t) = R(t)$.
	\end{itemize}

\end{slide}

### 2021-08-26 pres-8-27/slides/work-04.tex
SoReMo updates

--- 2021-07-27 draft/chicago-spatial-covid-draft.tex
@@
-\documentclass[11pt]{article}
-
-%imports
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{footnote}
-\usepackage{url}
-\usepackage{natbib}
-
-%images and image path
-\usepackage{graphicx}
-\usepackage{subfig}
-\graphicspath{ {./images/} }
-
-%custom commands
-\newcommand{\N}{\mathcal{N}}
-\newcommand{\num}{\text{num}}
-\newcommand{\obs}{\text{obs}}
-\newcommand{\D}{\mathfrak{D}}
-\newcommand{\Ro}{\mathcal{R}_0}
-
-%margins
-\usepackage[letterpaper, total={6.5in, 9.5in}]{geometry}
-
-%opening
-\title{Reaction-diffusion spatial modeling of COVID-19 in Chicago}
-\author{Trent Gerew}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-	We examine an application of the Law of Mass Action to model the outbreak of COVID-19 in Chicago.
-	We begin with a zero-dimension (ODE) compartmental epidemiological model consisting of Susceptible, Infected, and Removed populations (SIR model).
-	We then develop a spatially distributed version of the model in the form of reaction-diffusion equations (PDE).
-\end{abstract}
-
-\section{Introduction and Similar Work}
-
-\section{ODE and PDE Model Setup}
-We begin by explaining the ODE model which is simply obtained from the full PDE model by removing the diffusion terms in the reaction-diffusion equations.
-
-We start with a population of \textit{susceptibles} ($S$), which may become \textit{infected} ($I$) upon the emergence of the virus within the population.
-Infected individuals can interact with the susceptibles at rate $\beta$ to draw new members into the group $I$ of individuals infected by the virus.
-We note that the transmission rate $\beta$ incorporates the total population size (ODE model) or the total population density (PDE model).
-A fraction of the infected population recover or die at a rate $\gamma$, giving rise to the \textit{removed} ($R$) population.
-
-These parameters specify the ODE process.
-They represent processes that occur in a ``well mixed'' situation when no spatial dependence is assigned,
-or processes that happen locally at every point in space for the PDE's.
-We discuss parameter selection in the next section.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth]{sir_schematic}
-	\caption{Schematic diagram of the SIR model. The dashed line denotes the interaction of the infectious population with the susceptible population that leads to infection.}
-	\label{fig:schematic}
-\end{figure}
-
-The population model at the PDE level is an autonomous diffusion with a source:
-\begin{align}
-	S_t	&=	\Delta (\D_s \Delta S) - \beta SI,	\\
-	I_t	&=	\Delta (\D_I \Delta I) + \beta SI - \gamma I,	\\
-	R_t	&=	\gamma I.
-\end{align}
-In principle the $R$ population can have a diffusion, but since we assume this population has immunity we assign $\D_R = 0$.
-
-As in \cite{Kevrekidis_2021}, we have not incorporated any functional forms for directional spreading at the PDE level.
-In principle, these terms can represent ``daily practices'' (work commutes, etc.), as well as longer temporal or spatial scales (major thoroughfare traffic, etc.).
-We simply allow diffusion to perform the relevant spreading.
-Arriving infected individuals are assumed to form local hotspots within the susceptible population.
-This initial seeding roughly approximates long-range transportation.
-
-The next step is to identify the parameters at the ODE level.
-To do this, we use a nonlinear optimization algorithm.
-We determine the optimal parameters by minimizing the Euclidean distance between the time series generated by the model ($\num$) and the corresponding observed ($\obs$) time series,
-\begin{equation} \label{eq:obj}
-	\N = \sum_i^{t_\text{fit}^\text{end}} \left( \left| \log(C_\num (t_i)) - \log(C_\obs (t_i)) \right|^2 + \left| \log(R_\num (t_i)) - \log(R_\obs (t_i)) \right|^2 \right)
-\end{equation}
-where the index $i$ identifies a point in the time series.
-The parameters are optimized to reproduce the reported total number of infected cases $C(t) = I(t) + R(t)$, and the total number of removed $R(t)$.
-
-Since the transition rate $\gamma$ is an ``individual'' parameter, and thus acts ``on average'', we preserve the same value at the PDE level as at the ODE level.
-The transmission rate $\beta$ is more complicated, since at the PDE level the quantities $S$, $I$, and $R$ are densities rather than populations,
-which integrate over the spatial surface to the true population of each category.
-
-\section{Results}
-\subsection{ODE Model}
-We begin by discussing the zero-dimensional model for Chicago.
-Data for the progression of the pandemic in Chicago were obtained from the Chicago Data Portal \cite{Chicago_2021}.
-The relevant results are given in Figure \ref{fig:ode_results}.
-We obtained the optimal ODE model parameters for Chicago from 1,000 optimizations (implemented in MATLAB via the \verb|fminsearch| function)
-that compared model predictions to jointly the number of cumulative infected and the deceased, as in Equation \ref{eq:obj}.
-For each optimization, the initial guess for each parameter was uniformly sampled within a pre-specified range.
-The parameter ranges were determined from epidemiological information.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=.6\textwidth]{obs_data}
-	\caption{Observed data for Chicago showing the three impact waves.}
-	\label{fig:obs_data}
-\end{figure}
-
-The COVID-19 pandemic had three major impact waves in Chicago, as seen clearly in Figure \ref{fig:obs_data}.
-For simplicity, we analyze only the first major wave, which ends at approximately October 30, 2020.
-Additionally, we begin the analysis at March 17 instead of the first day of data collection (March 1) because of reporting errors early in the pandemic.
-
-Upon performing the optimizations we find that the fitting yields the results summarized in Table \ref{tab:parameters}.
-We show the median parameters, as well as the interquartile range, and the range of variation used to sample the parameters.
-Model predictions (with median parameter values, solid blue lines) are compared graphically to data (black dots) in Figure \ref{fig:ode_results}.
-
-The optimal parameters were determined for a scenario in which restrictive measures (quarantine, lockdown) were not enforced.
-Therefore, we do not impose a time dependence on the transmission rate $\beta$.
-The transmission rate represents the rate of contact of susceptible individuals with infectious individuals that leads to infection by the virus.
-It may be expressed as the product of daily average contacts and the probability of infection given a contact \cite{Drossinos_2020}.
-Thus, restrictive measures such as mobility restrictions, social-distancing, face-mask wearing, and gathering limitations are expected to decrease the transmission rate, an effect not accounted for in this model.
-
-\begin{savenotes}
+\begin{slide}{Current Work}
 	\begin{table}[h]
@@
 			Transition rate, $I \rightarrow R$ [per day]	&	$\gamma$	&	 0.39656(0.39654-0.39659)	&	$c \in U[0.25,0.75]$ \\
-		Diffusivity, $S$ [km$^2$/day]	&	$\D_S$	&	10	&	\\
-		Diffusivity, $I$ [km$^2$/day]	&	$\D_I$	&	100	&	\\
 			\hline
@@
 	\end{table}
-\end{savenotes}
-
-The two top panels of Figure \ref{fig:ode_results} show how well we capture the data for the Chicago outbreak.
-The model reproduces reasonable well the number of cases (top left panel) but not the number of deaths (top right panel).
-The model fails to capture the characteristic S-shape in the logarithmic plot of the top right panel (deaths).
-
-\begin{figure}
-	\centering
-	\subfloat{\includegraphics[width=.475\textwidth]{ode_cases}}
-	\subfloat{\includegraphics[width=.475\textwidth]{ode_deaths}} \\
-	\subfloat{\includegraphics[width=.475\textwidth]{ode_IR}}
-	\caption{ODE model for Chicago with fitting to official data from March 17, 2020 ($t = t_\text{init} = 1$) to September 30, 2020 ($t = t_\text{fit}^\text{end} = 198$).
-		Left panel: Confirmed cases $C(t) = I(t) + R(t)$; Right panel: Number of deaths $D(t)$.
-		Bottom panel shows the other populations: infected $I(t)$ and recovered $R(t)$.}
-	\label{fig:ode_results}
-\end{figure}
-
-\appendix
-\section{ODE Dynamics}
-Here we briefly discuss some dynamics of the ODE system.
-Note that for the simply SIR model as we have here, $R$ is completely determined when $S$ and $I$ are known, so we need only consider
-\begin{align*}
-S_t	&=	- \tilde{\beta} SI, \\
-I_t	&=	\tilde{\beta} SI - \gamma I,
-\end{align*}
-where $\tilde{\beta} = \beta / N$.
-
-Immediately we see there are infinitely many non-isolated fixed points of the form $(S,0)$.
-Linearizing, we get
-\begin{equation*}
-\mathbf{A}_{(S,0)} = \begin{bmatrix} 0 & - \tilde{\beta} S \\ 0 & \beta S - \gamma \end{bmatrix}.
-\end{equation*}
-Thus, as seen in Figure \ref{fig:localdyn}, for $S < \frac{\gamma}{\tilde{\beta}}$ the nodes are stable, and for $S > \frac{\gamma}{\tilde{\beta}}$ they are unstable.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.25\textwidth]{local_dynamics}
-	\caption{Stream plot for the local ODE dynamics. Note the transcritical bifurcation around $S_0 = 0.4$.}
-	\label{fig:localdyn}
-\end{figure}
-We can then conclude that an epidemic will occur only if $S_0 = S(t=0) > \frac{\gamma}{\tilde{\beta}}$ \cite{Wang_SIR}.
-Hence, a transcritical bifurcation occurs at $\Ro = \tilde{\beta} S_0 / \gamma$, which is known as the \textit{basic reproduction number}, defined as the average number of secondary infections caused by a single infected individual \cite{Drossinos_2020}.
-
-Note that the ODE model can be viewed as a generalized chemical reaction
-\begin{align*}
-S + I	&\longrightarrow	2 I, \\
-I		&\longrightarrow	R.
-\end{align*}
-However, these two reactions are not reversible, so we are not able to write the energy-dissipation law for the ODE model \cite{Forster_2013, Wang_2020}.
-Indeed, taking $\frac{I_t}{S_t} = - 1 + \frac{\gamma}{\tilde{\beta} S}$ and solving this differential equation gives $I = - S + \frac{\gamma}{\tilde{\beta}} \ln(S) + C$.
-Thus,
-\begin{equation}
-E(S,I) = S + I - \frac{\gamma}{\tilde{\beta}} \ln(S)
-\end{equation}
-is a conserved quantity for the system.
-
-\bibliographystyle{abbrv}
-\bibliography{chicago-spatial-covid-draft}
-
-\end{document}
+\end{slide}

### 2021-08-26 pres-8-27/slides/work-04.tex
SoReMo updates

@@
 \begin{slide}{Current Work}
+	{\large \textbf{ODE Results}} \\
+
+	\vspace{.5cm}
+
+	Data were obtained from the Chicago Data Portal. \\
+	Optimal ODE model parameters obtained from 1,000 optimizations (implemented in MATLAB). \\
+
+	\vspace{.3cm}
+
 	\begin{table}[h]
 		\centering
-		\caption{ODE parameters for Chicago: optimal (best-fitting), median and interquartile range, and variation range used in the optimization algorithm.
-Initial parameter guesses were uniformly sampled within these ranges.}
 		\label{tab:parameters}
@@
 			Initial population	&	$(I_0, R_0)$	&	(127, 2)	& \\
-			Transmission rate, $S \rightarrow I$ [per day]	&	$\beta$\footnote{The transmission rate $\beta$ must be divided by $N$ when used in the ODE model.}	&	0.38206(0.38204-0.38209)	&	$c \in U[0,1]$ \\
+			Transmission rate, $S \rightarrow I$ [per day]	&	$\beta$	&	0.38206(0.38204-0.38209)	&	$c \in U[0,1]$ \\
 			Transition rate, $I \rightarrow R$ [per day]	&	$\gamma$	&	 0.39656(0.39654-0.39659)	&	$c \in U[0.25,0.75]$ \\
@@
 	\end{table}
+
+	Note the transmission rate $\beta$ must be divided by $N$ when used in the ODE model.
 \end{slide}

### 2021-08-26 pres-8-27/slides/social-01.tex
SoReMo updates

\begin{slide}{Social and Community Connections}
	{\large Original context:} Working with David Eads (NPR) to identify \textit{demographic trends} in COVID-19. \\

	\vspace{.2cm}

	What are the effects of \textbf{age} and \textbf{race/ethnicity} in the Chicago area?
\end{slide}

### 2021-08-27 pres-8-27/slides/work-00.tex
SoReMo updates

\begin{slide}{Current Work}
	{\large \textbf{Idea:}}
	\begin{enumerate}
		\item Define a \textit{local} process that occurs locally at every point in space.
		\item Determine suitable \textit{local} parameters (e.g., transition rates).
		\item Define the characteristics of the \textit{physical space} (e.g., population density).
		\item Integrate the local process over the physical space.
		\item Determine suitable \textit{physical} parameters (e.g., transmission rates).
	\end{enumerate}

	\vspace{.5cm}

	{\large \textbf{Progress:}}
	\begin{itemize}
		\item Start by defining a toy model: \textbf{S}usceptible, \textbf{I}nfected, \textbf{R}emoved.
		\item Local parameters fitted using data from the \alert{Chicago Data Portal}.
		\item Spatial characteristics defined using data from the \alert{U.S. Census Bureau}.
		\item Numerical spatial integration is a work in progress.
	\end{itemize}
\end{slide}

### 2021-10-26 draft/chicago-spatial-covid-draft.tex
Added PDE section
Fixed deltas to nablas
Lots of updates.

@@
 \begin{align}
-	S_t	&=	\Delta (\D_s \Delta S) - \beta SI,	\\
-	I_t	&=	\Delta (\D_I \Delta I) + \beta SI - \gamma I,	\\
+	S_t	&=	\nabla (\D_s \nabla S) - \beta SI,	\\
+	I_t	&=	\nabla (\D_I \nabla I) + \beta SI - \gamma I,	\\
 	R_t	&=	\gamma I.
@@
 \begin{equation} \label{eq:obj}
-	\N = \sum_i^{t_\text{fit}^\text{end}} \left( \left| \log(C_\num (t_i)) - \log(C_\obs (t_i)) \right|^2 + \left| \log(D_\num (t_i)) - \log(D_\obs (t_i)) \right|^2 \right)
+	\N = \sum_i^{t_\text{fit}^{\text{end}}} \sqrt{ (C_\num (t_i) - C_\obs (t_i))^2 + (D_\num (t_i) - D_\obs (t_i))^2}
 \end{equation}
@@
 
+\subsection{PDE Model}
+We write the system in the form of a reaction-diffusion equation
+\begin{equation} \label{eq:rd}
+	\partial_t \mathbf{u} = \D \nabla^2 \mathbf{u} + \mathbf{R (u)}
+\end{equation}
+where $\mathbf{u} = \mathbf{u}(x,y,t)$ is a vector of density variables,  $\mathbf{R(u)}$ is a nonlinear term describing local reaction kinetics, and $\D$ denotes a diagonal diffusion coefficient matrix.
+In this case we have
+\begin{align*}
+	\mathbf{u} &= \begin{pmatrix} S(x,y,t) \\ I(x,y,t) \\ R(x,y,t) \end{pmatrix} &
+	\mathbf{R(u)} &= \begin{pmatrix} - \beta S I \\ \beta S I - \gamma I \\ \gamma I \end{pmatrix} &
+	\D &= \begin{pmatrix} \D_S & 0 & 0 \\ 0 & \D_I & 0 \\ 0 & 0 & 0 \end{pmatrix}
+\end{align*}
 \appendix

### 2021-10-26 pres-8-27/Plaincolortheme.tex
Lots of updates.

% This effectively sets an IIT color theme with no background bars and is
% derived from the beaver theme that is built into beamer
% 
\definecolor{iitred}{rgb}{0.7423,0.0020,0.0192}

\setbeamercolor{section in toc}{fg=black,bg=white}
\setbeamercolor{alerted text}{fg=iitred}
\setbeamercolor*{palette primary}{fg=iitred,bg=white}
\setbeamercolor*{palette secondary}{fg=iitred,bg=white}
\setbeamercolor*{palette tertiary}{bg=white,fg=iitred}
\setbeamercolor*{palette quaternary}{fg=iitred,bg=white}

\setbeamercolor*{sidebar}{fg=iitred,bg=white}

\setbeamercolor*{palette sidebar primary}{fg=iitred}
\setbeamercolor*{palette sidebar secondary}{fg=iitred}
\setbeamercolor*{palette sidebar tertiary}{fg=iitred}
\setbeamercolor*{palette sidebar quaternary}{fg=iitred}

%\setbeamercolor*{titlelike}{parent=palette primary}
\setbeamercolor{titlelike}{parent=palette primary,fg=iitred}
\setbeamercolor{frametitle}{bg=white}
\setbeamercolor{frametitle right}{bg=white}

\setbeamercolor*{separation line}{}
\setbeamercolor*{fine separation line}{}
\setbeamercolor{block body}{parent=normal text,use=block title,bg=red!0,fg=}
\setbeamercolor{local structure}{fg=iitred}

% This sets all text color
%\setbeamercolor{normal text}{fg=iitred}

### 2021-10-26 pres-8-27/slides/future-02.tex
Lots of updates.

\begin{slide}{Future Goals}
	{\large Transition to PDE model (proof of concept)}

	\vspace{.5cm}

	\begin{itemize}
		\item The transition rate $\gamma$ is an ``individual'' parameter, so the value is preserved at the PDE level.
		\item The transmission rate $\beta$ depends on the interaction between individuals, so this must be modified.

		\vspace{.3cm}

		\item The geometry (shape and density distribution) of Chicago must be defined (Census data).
		\item Suitable initial infected distributions must be seeded.
	\end{itemize}
\end{slide}

### 2021-10-26 pres-8-27/slides/social-01.tex
SoReMo updates
Lots of updates.

@@
 \begin{slide}{Social and Community Connections}
-	{\large Original context:} Working with David Eads (NPR) to identify \textit{demographic trends} in COVID-19. \\
+	{\large Summer 2020 context:} IPRO with David Eads (NPR) to identify \textit{demographic trends} in COVID-19. \\
 
@@
 
-	What are the effects of \textbf{age} and \textbf{race/ethnicity} in the Chicago area?
+	\begin{itemize}
+		\item How does the impact of COVID-19 (in Chicago) depend on \textbf{age} and \textbf{race/ethnicity}?
+	\end{itemize}
+
+	\vspace{1cm}
+
+	{\large Present context:} COVID-19 is still here. Where is it impacting and why?
+
+	\vspace{.2cm}
+
+	\begin{itemize}
+		\item Does \textbf{community structure} play a role?
+		\item How effective are \textbf{vaccines}? (Or what is thwarting them?)
+		\item Are \textbf{care facilities} properly allocated?
+		\item \emph{Other ideas?}
+	\end{itemize}
+
+	\vspace{.7cm}
+
+	{\large \textbf{Goal:} Can we design a model to answer these questions?}
 \end{slide}

### 2021-10-26 pres-8-27/slides/work-05.tex
Lots of updates.

\begin{slide}{Current Work}
	{\large ODE model for Chicago with fitting to official data from March 17, 2020 ($t = t_\text{init} = 1$) to September 30, 2020 ($t = t_\text{fit}^\text{end} = 198$)}
	\begin{figure}
		\centering
		\subfloat[Confirmed cases $C(t) = I(t) + R(t)$]{\includegraphics[width=.333\textwidth]{images/ode_cases}}
		\subfloat[Number of deaths $D(t)$]{\includegraphics[width=.333\textwidth]{images/ode_deaths}}
		\subfloat[Infected $I(t)$ and Recovered $R(t)$]{\includegraphics[width=.333\textwidth]{images/ode_IR}}
	\end{figure}
\end{slide}

### 2021-10-26 update-10-22/update-10-22.tex
Lots of updates.

\documentclass[11pt]{article}

%imports
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{footnote}
\usepackage{url}
\usepackage{natbib}

%images and image path
\usepackage{graphicx}
\usepackage{subfig}
\graphicspath{ {./images/} }

%custom commands
\newcommand{\N}{\mathcal{N}}
\newcommand{\num}{\text{num}}
\newcommand{\obs}{\text{obs}}
\newcommand{\D}{\mathfrak{D}}
\newcommand{\Ro}{\mathcal{R}_0}
\newcommand{\lap}{{\mathcal{L}}}
\renewcommand\vec{\mathbf}
\newcommand{\mat}[1]{\mathsf{#1}}

%insitute command
\usepackage{etoolbox}
\makeatletter
\providecommand{\institute}[1]{% add institute to \maketitle
	\apptocmd{\@author}{\end{tabular}
	\par
	\begin{tabular}[t]{c}
		\small \textit{#1}}{}{}
}
\makeatother

%margins
\usepackage[letterpaper, total={6.5in, 9.5in}]{geometry}

%opening
\title{Reaction-diffusion spatial modeling of COVID-19 in Chicago}
\author{Trent Gerew\thanks{\texttt{tgerew@hawk.iit.edu}}}
\institute{Department of Applied Mathematics, Illinois Institute of Technology, Chicago, Illinois}

\begin{document}

\maketitle

\section{Model Setup}

	\begin{figure}[h!]
		\centering
		\includegraphics[height=6cm]{full-model}
		\label{fig:model}
		\caption{Schematic diagram of the model. The dashed lines indicate the interaction of the infected populations with the susceptible populations that leads to infection.}
	\end{figure}

	\begin{align}
		S_t &=	\D_S \Delta S - \beta_{SA} S A - \beta_{SI} S I - \mu S, \\
		E_t	&=	\D_E \Delta E + \beta_{SA} S A + \beta_{SI} S I - (\sigma_A + \sigma_I) E, \\
		A_t	&=	\D_A \Delta A + \sigma_A E - M_{AR} A, \\
		I_t	&=	\sigma_I E - M I, \\
		H_t	&=	\gamma M I - (1 - \omega) \chi H - \omega \chi H, \\
		R_t	&=	(1 - \gamma) M I + (1 - \omega) \chi H, \\
		D_t	&=	\omega \chi H.
	\end{align}

	\begin{table}[h]
		\centering
		\caption{Population values for Chicago.}
		\label{tab:populations}
		\begin{tabular}{ c c c }
			\hline
			\hline
			&	&	Population \\
			\hline
			Total population		&	$N$		&	2,695,598 \\
			Initial infected		&	$I_0$	&	127	\\
			Initial hospitalized	&	$H_0$	&	30 \\
			Initial deceased		&	$D_0$	&	6 \\
			\hline
			\hline
		\end{tabular}
	\end{table}

	\begin{table}[h]
		\centering
		\caption{Parameters for Chicago: optimal (best-fitting), median and interquartile range, and variation range used in the optimization algorithm. Initial parameter guesses were uniformly sampled within these ranges.}
		\label{tab:parameters}
		\begin{tabular}{ c c c c }
			\hline
			\hline
			&	&	Median (interquartile range)	&	Initial value \\
			\hline
			Population	&	$N$	&	2,695,598	& \\
			Initial population	&	$(I_0, R_0)$	&	(127, 2)	& \\
			Transmission rate, $S \rightarrow I$ [per day]	&	$\beta$\footnote{The transmission rate $\beta$ must be divided by $N$ when used in the ODE model.}	&	0.38206(0.38204-0.38209)	&	$c \in U[0,1]$ \\
			Transition rate, $I \rightarrow R$ [per day]	&	$\gamma$	&	 0.39656(0.39654-0.39659)	&	$c \in U[0.25,0.75]$ \\
			Diffusivity, $S$ [km$^2$/day]	&	$\D_S$	&	10	&	\\
			Diffusivity, $I$ [km$^2$/day]	&	$\D_I$	&	100	&	\\
			\hline
			\hline
		\end{tabular}
	\end{table}

\section{ODE Dynamics}
	We want to understand the trajectories of the dynamics of the ODE system under different initial conditions.
	To do this we first find the equilibrium points by solving
	$$S_t = E_t = A_t = I_t = H_t = R_t = D_t = 0$$
	simultaneously for $\vec{x} = (S, E, A, AR, I, H, R, D)$.
	The solutions of this system are of the form $\vec{x^*} = (0, 0 , 0, AR, 0, 0, R, D)$.
	This implies there are infinitely many non-isolated equilibrium points.
	We determine the stability of these equilibrium points by analyzing the linearized system near the points.
	The Jacobian of the system is
	\begin{equation}
		\mat{J} = \begin{pmatrix}
			- A \beta_{SA} - I \beta_{SI} - \mu	&	0						&	- S \beta_{SA}	&	0	&	- S \beta_{SI}	&	0									&	0	&	0	\\
			A \beta_{SA} + I \beta_{SI}			&	- \sigma_A - \sigma_I	&	S \beta_{SA}	&	0	&	S \beta_{SI}	&	0									&	0	&	0	\\
			0									&	\sigma_A				&	- M_{AR}		&	0	&	0				&	0									&	0	&	0	\\
			0									&	0						&	M_{AR}			&	0	&	0				&	0									&	0	&	0	\\
			0									&	\sigma_I				&	0				&	0	&	- M				&	0									&	0	&	0	\\
			0									&	0						&	0				&	0	&	M \gamma		&	- \chi (1 - \omega) - \psi \omega	&	0	&	0	\\
			0									&	0						&	0				&	0	&	M (1 - \gamma)	&	\chi (1 - \omega)					&	0	&	0	\\
			0									&	0						&	0				&	0	&	0				&	\psi \omega							&	0	&	0
		\end{pmatrix}
	\end{equation}
	Now evaluating $\mat{J}$ at the equilibrium point $\vec{x^*}$ and calculating the eigenvalues, we have
	\begin{equation}
		\vec{\lambda} = \{0 , 0, 0, - M, - M_{AR}, - \mu, - \sigma_A - \sigma_I, - \chi + \chi \omega - \psi \omega \}.
	\end{equation}
	Note that the first three eigenvalues are 0, which implies the equilibrium points are non-isolated.
	This agrees with our earlier observation.

	The equilibrium points are stable when $\lambda_i < 0$ for $4 \leq i \leq 8$.
	Since all the system parameters are positive, this implies $\lambda_i < 0$ for $4 \leq i \leq 7$.
	Thus the stability depends on the sign of $\lambda_8$.
	There are two cases when $\lambda_8 = - \chi + \chi \omega - \psi \omega < 0$ is true:
	\begin{enumerate}
		\item $0 < \omega \leq 1$ implies $\lambda_8 < 0$, and
		\item $\omega > 1$ and $\chi < \frac{\psi \omega}{\omega - 1}$ implies $\lambda_8 < 0$.
	\end{enumerate}
	That is, whenever we have either of these conditions the equilibrium points are stable.
	We call this situation \textit{endemic}.
	If $\lambda_8 > 0$, the equilibrium points are unstable and the situation is an \textit{epidemic}.

\bibliographystyle{abbrv}
\bibliography{chicago-spatial-covid-draft}

\end{document}

### 2021-10-26 local-model/odeModel.m

@@ -13,32 +13,12 @@ r0 = 2; %recovered
 s0 = n - i0 - r0; %susceptible
 y0 = [s0, i0, r0]; %initial conditions vector
 
-% Time domain
-%t = 1:61; %60 days
-
 %% Observed Data
 % Load time, total cases and total deaths
 [t,Cobs,Dobs] = loadData();
 
 % Convert to a single vector
 Yobs = [Cobs(:); Dobs(:)];
-
-%% Test Data (with noise)
-% % Parameters
-% p = [0.3/n, 0.3]; %[beta, gamma]
-% 
-% % Compute simulation
-% [~,Y] = ode45(@(t,y)sir(t,y,p),t,y0);
-% I = Y(:,2);
-% R = Y(:,3);
-% 
-% % Define total cases and deaths, and add noise
-% noise = 1000*sawtooth(t)';
-% Cobs = I + R + noise;
-% Dobs = R + noise;
-% 
-% % Convert to a single vector
-% Yobs = [Cobs(:); Dobs(:)];
 
 %% Model Solution
 % Define optimization iterations

### 2021-10-26 local-model/pdeModel.m

@@ -1,0 +1,114 @@
+%% Simple SIR in two spacial dimensions.
+
+clear;
+clc;
+close all;
+
+%% Initialize a two equation PDE model.
+
+model = createpde(2);
+
+%% Establish a model for the polygonal domain
+poly1 = [2 % 2 specifies we are forming a polygon
+    6 % Number of edges
+    1 % x1
+    1 % x2
+    2 % x3
+    2 % x4
+    0 % x5
+    0 % x6
+    1 % y1
+    0 % y2
+    0 % y3
+    2 % y4
+    2 % y5
+    1 % y6
+    ];
+
+gd = [poly1];
+
+ns = char('poly1');
+ns = ns';
+sf = 'poly1';
+[dl,bt] = decsg(gd,sf,ns);
+
+% Plot the computational domain
+fig1 = figure(1);
+pdegplot(dl,'EdgeLabels','on','FaceLabels','on')
+axis equal
+
+% Attach the computational domain to the PDE model
+geometryFromEdges(model,dl);
+% geometryFromEdges(model,@squareg);
+% 
+% pdegplot(model,'EdgeLabels','on')
+% ylim([-1.5,1.5])
+% axis equal
+
+%% Boundary Conditions
+%applyBoundaryCondition(model,'neumann','edge',1:6,'g',[0,0],'q',[0,0]);
+
+%% Specify coefficients for the PDE model
+% Define the parameters of the system
+Ds = 0.001;
+Di = 0.01;
+beta = 0.38206;
+gamma = 0.39656;
+
+% Based on these parameters, specify the PDE coefficients in the toolbox
+% format
+d = [1;1];
+c = [Ds;Di];
+f = @(location,state) [-beta * state.u(1,:) .* state.u(2,:);
+                       beta * state.u(1,:) .* state.u(2,:) - gamma * state.u(2,:)];
+
+specifyCoefficients(model,'m',0,'d',d,'c',c,'a',0,'f',f);
+
+%% Generate the mesh
+% Spacial domain
+generateMesh(model,'Hmax',0.1);
+% Plot the mesh
+fig2 = figure(2);
+pdeplot(model)
+axis equal
+
+% Time domain
+tlist = 0:0.1:200;
+
+%% Initial conditions
+setInitialConditions(model,@u0fun2);
+
+%% Solve the PDE
+model.SolverOptions.ReportStatistics='on';
+result = solvepde(model,tlist);
+
+%% Display the results
[…]

### 2021-10-26 odelModelComplete.m

@@ -1,0 +1,215 @@
+% Clear windows and workspace
+clear
+clc
+close all
+
+%% Problem Definition
+% Population
+n = 2695598; % Total population
+
+% Known initial conditions
+i0 = 99; % Infected
+h0 = 44; % Hospitalized
+d0 = 2; % Deceased
+
+% Initial Ratios
+e0i0 = 1; % Exposed/Infected
+a0i0 = 0.5; % Asymptomatic/Infected
+
+% Calculate initial conditions for other compartments
+e0 = e0i0 * i0; % Exposed
+a0 = a0i0 * i0; % Asymptomatic
+
+ar0 = 0; % Asymptomatic recovered
+r0 = 0; % Recovered
+
+s0 = n - (e0 + a0 + ar0 + i0 + h0 + r0 + d0); % Susceptible
+
+% Initial conditions vector
+y0 = [s0, e0, a0, ar0, i0, h0, r0, d0];
+
+%% Observed Data
+% Load time, total cases, and total deaths
+[t, Cobs, Dobs] = loadData();
+
+% Convert to a single vector
+Yobs = [Cobs(:); Dobs(:)];
+
+%% Model Solution
+% Define optimization iterations
+iters = 10;
+
+% Initialize estimate vectors
+beta_is = rndSample([0,1],iters); % Transmission rate, S -> I
+beta_as = rndSample([0,1],iters); % Transmission rate, S -> A
+sigma_a = 1./rndSample([2,7],iters); % Latent period, E -> A
+sigma_i = 1./rndSample([2,7],iters); % Incubation period, E -> A
+m = 1./rndSample([5,12],iters); % Infectivity period
+m_ar = 1./rndSample([5,12],iters); % Recovery period, A -> AR
+chi = 1./rndSample([5,20],iters); % Recovery period, H -> R
+psi = 1./rndSample([5,20],iters); % Period to deseased, H -> D
+gamma = rndSample([0.25,0.75],iters); % Conversion fraction, I -> H, I -> D
+omega = rndSample([0.1,0.5],iters); % Conversion fraction, H -> D, H -> R
+
+e0i0 = rndSample([1,6],iters); % Initial Exposed/Infected
+a0i0 = rndSample([1,6],iters); % Initial Asymptotic/Infected
+
+% Fixed Parameters
+mu = 5.79e-07;
+
+% Divide beta by n for ODE model
+beta_is = beta_is./n;
+beta_as = beta_as./n;
+
[…]

### 2021-10-26 spatial-model/ZipGeometry.m

@@ -1,0 +1,37 @@
+clear
+clc
+close all;
+
+zips = shaperead('chicago_zips.shp','UseGeoCoords',true);
+h = height(zips);
+faceColors = makesymbolspec('Polygon',...
+    {'INDEX',[1 h],'FaceColor',polcmap(h)});
+
+geoshow(zips,'SymbolSpec',faceColors)
+
+% Create a polygon for each ZIP code
+for zip = 1:length(zips)
+    poly(zip) = simplify(polyshape(zips(zip).Lon, zips(zip).Lat));
+end
+
+
+% % Take the union of the polygons to create the entire domain
+% polyout = simplify(union(poly));
+% 
+% % Create a triangulation representation of the domain
+% tr = triangulation(polyout);
+% 
+% % Create a PDE model
+% model = createpde;
+% 
+% tnodes = tr.Points';
+% telements = tr.ConnectivityList';
+% 
+% geometryFromMesh(model,tnodes,telements);
+% pdegplot(model)
+% 
+% generateMesh(model,'Hgrad',1.3)
+% 
+% figure
+% pdemesh(model)
+% pdegplot(model)

### 2021-10-26 spatial-model/mapTest.m

@@ -1,0 +1,41 @@
+clear
+clc
+close all;
+
+zips = shaperead('chicago_zips.shp','UseGeoCoords',true);
+h = height(zips);
+faceColors = makesymbolspec('Polygon',...
+    {'INDEX',[1 h],'FaceColor',polcmap(h)});
+
+geoshow(zips,'SymbolSpec',faceColors)
+
+%% COVID data by ZIP code
+% Import the shapefile
+covid_zips = shaperead('chicago_covid_zips.shp','UseGeoCoords',true);
+
+% Remove unused fields
+fields = ["time_week_", "time_wee_2", "cases_cumu", "case_rate_", "case_rat_2", ...
+    "tests_week", "tests_cumu", "test_rate_", "test_rat_2", "percent_te", ...
+    "percent__2", "deaths_wee", "death_rate", "death_ra_2", "row_id"];
+covid_zips = rmfield(covid_zips,fields);
+
+
+maxweeks = max([covid_zips.week_numbe]);
+
+for week = 1:maxweeks
+    rows = ismember([covid_zips.week_numbe],week);
+    week_zips = covid_zips(rows,:);
+    
+    rows = ~isnan([week_zips.cases_week]);
+    week_zips = covid_zips(rows,:);
+    
+    maxcases = max([week_zips.cases_week]);
+    fall = flipud(autumn(height(week_zips)));
+    
+    casesColors = makesymbolspec('Polygon',{'cases_week', ...
+        [0 maxcases], 'FaceColor', fall});
+    
+    refresh
+    geoshow(week_zips)
+    pause(.2)
+end

### 2021-10-26 spatial-model/maptest2.m

@@ -1,0 +1,19 @@
+MapLatLimit = [41 48];
+MapLonLimit = [-74 -66];
+
+states = readgeotable('usastatelo.shp');
+n = ["Maine" "New Hampshire" "Vermont" "Massachusetts" "Connecticut" "Rhode Island" "New York"];
+rows = ismember(states.Name,n);
+NEstates = states(rows,:);
+
+maxdensity = max([NEstates.PopDens2000])
+fall = flipud(autumn(height(NEstates)));
+
+densityColors = makesymbolspec('Polygon', {'PopDens2000', ...
+    [0 maxdensity], 'FaceColor', fall});
+
+axesm('MapProjection', 'eqaconic', 'MapParallels', [],...
+  'MapLatLimit', MapLatLimit, 'MapLonLimit', MapLonLimit,...
+  'GLineStyle', '-')
+geoshow(NEstates, 'DisplayType', 'polygon', ...
+    'SymbolSpec', densityColors)

### 2021-10-26 spatial-model/pdeTest2.m

@@ -7,8 +7,8 @@ clear
 clc
 
 %% Step parameters
-J = 20; % Number of discrete steps
-delta_x = pi/20; % Space mesh
+J = 100; % Number of discrete steps
+delta_x = pi/J; % Space mesh
 s = 5/11; % s = delta_t / (delta_x)^2 < 1/2
 
 %% Define matrix

### 2021-10-26 spatial2/chicago_zip.m

@@ -1,0 +1,6 @@
+dom = shaperead('domain');
+landuse = shaperead('landuses');
+streams = shaperead('streams');
+wells = shaperead('wells');
+
+aquif = CSGobj_v2(2,20,200,500,1);

### 2021-10-27 data/data_collection.py
plot updates

@@ -29,7 +29,7 @@ df["Cumulative Deaths"] = df["Daily Deaths"].cumsum()
 df = df.dropna()
 
 # Select a date range (also excludes any missing dates)
-df = df.loc["2020-03-01":"2021-07-16"]
+df = df.loc["2020-03-01":"2021-10-26"]
 
 # Export to a csv
 df.to_csv("chicago_data.csv")
@@ -37,3 +37,7 @@ df.to_csv("chicago_data.csv")
 # Create plot
 df.plot(figsize=(12,8))
 plt.show()
+
+# Export plot
+import tikzplotlib
+tikzplotlib.save("chicago-data.tex")

### 2021-10-27 version2/beta_asfun.m

@@ -4,7 +4,7 @@ function beta = beta_asfun(p,t_q,t)
     eta_as = p(2);
     
     % Create a t_q vector
-    t_q = t_q * ones(1, length(t));
+    t_q = t_q * ones(length(t),1);
     
     % Calculate beta_is at time t
     beta = beta_as * (eta_as + (1 - eta_as) * (1 - tanh(2 * (t -  t_q))) / 2);

### 2021-10-27 version2/sir.m

@@ -1,4 +1,4 @@
-function dydt = sir(t,y,p1)
+function dydt = sir(t,y,p,q)
 % SIR   SIR model function.
 
     % Derivatives with respect to time
@@ -10,28 +10,36 @@ function dydt = sir(t,y,p1)
     
     % System parameters
     % Optimized parameters (p vector)
-    sigma_a = p1(1);
-    sigma_i = p1(2);
-    m_ar = p1(3);
-    m = p1(4);
-    gamma = p1(5);
-    omega = p1(6);
-    chi = p1(7);
-    psi = p1(8);
-    beta_sa = p1(9);
-    beta_si = p1(10);
-    eta_sa = p1(11);
-    eta_si = p1(12);
-    % Fixed parameters (q vector)
-    mu = p1(13);
-    t_q = p1(14);
+    sigma_a = p(1);
+    sigma_i = p(2);
+    m_ar = p(3);
+    m = p(4);
+    gamma = p(5);
+    omega = p(6);
+    chi = p(7);
+    psi = p(8);
+    beta_sa = p(9);
+    beta_si = p(10);
+    eta_sa = p(11);
+    eta_si = p(12);
     
+    % Fixed parameters (q vector)
+    mu = q(1);
+    t_q = q(2);
+    
+    % Calculate time dependent betas
+    beta_si_t = betafun([beta_si,eta_si],t_q,t);
+    beta_sa_t = betafun([beta_sa,eta_sa],t_q,t);
+
     % Define system
     dydt = [
-        -beta_asfun([beta_as,eta_as],t_q,t) .* s * a - beta_
-    
-    
-    
-    
-    
-    
+        -beta_sa_t .* (s * a) - beta_si_t .* (s * i) - mu * s; % S_t
+        beta_sa_t .* (s * a) + beta_si_t .* (s * i) - (sigma_a + sigma_i) * e; % E_t
+        sigma_a * e - m_ar * a; % A_t
+        m_ar * a; % AR_t
+        sigma_i * e - m * i; % I_t
+        gamma * m * i - (1 - omega) * chi * h - omega * psi * h; % H_t
+        (1 - gamma) * m * i + (1 - omega) * chi * h; % R_t
+        omega * psi * h; % D_t
+        ];
+end

### 2021-10-27 version2/test.m

@@ -1,0 +1,59 @@
+clear
+clc
+close all
+
+n = 8414240;
+
+i0 = 297;
+h0 = 134;
+d0 = 6;
+
+mu = 5.79e-7;
+beta_is = 0.32 / n;
+beta_as = 0.28 / n;
+eta_is = 0.39;
+eta_as = 0.38;
+sigma_a = 1 / 2.97;
+sigma_i = 1 / 3.77;
+m = 1 / 5.97;
+m_ar = 1 / 6.85;
+chi = 1 / 6.74;
+psi = 1 / 8.33;
+gamma = 0.55;
+omega = 0.25;
+
+t_q = 16;
+
+e0i0 = 2.69;
+a0i0 = 2.93;
+
+e0 = e0i0 * i0;
+a0 = a0i0 * i0;
+
+ar0 = 0;
+r0 = 0;
+
+s0 = n - (e0 + a0 + ar0 + i0 + h0 + r0 + d0);
+
+p = [sigma_a, sigma_i, m_ar, m, gamma, omega, chi, psi, beta_as, beta_is, eta_as, eta_is];
+q = [mu, t_q];
+
+y0 = [s0, e0, a0, ar0, i0, h0, r0, d0];
+
+tspan = [1, 60];
+
+[T,Y] = ode45(@(t,y)sir(t,y,p,q),tspan,y0);
+
+s = Y(:,1);
+e = Y(:,2);
+a = Y(:,3);
+ar = Y(:,4);
+i = Y(:,5);
+h = Y(:,6);
+r = Y(:,7);
+d = Y(:,8);
+
+Cnum = i + h + r + d;
+
+figure()
+semilogy(T,Cnum)

### 2021-10-27 update-10-22/update-10-22.tex
plot updates
Fixed missing AR equation

@@
 \usepackage{natbib}
+\usepackage{pgfplots}
 
@@
 
+	\begin{figure}[h]
+		\centering
+		\input{chicago-data}
+		\label{fig:data}
+		\caption{no}
+	\end{figure}
+
 \section{Model Setup}
@@
 		E_t	&=	\D_E \Delta E + \beta_{SA} S A + \beta_{SI} S I - (\sigma_A + \sigma_I) E, \\
+		AR_t &= M_{AR} A, \\
 		A_t	&=	\D_A \Delta A + \sigma_A E - M_{AR} A, \\
 		I_t	&=	\sigma_I E - M I, \\
-		H_t	&=	\gamma M I - (1 - \omega) \chi H - \omega \chi H, \\
+		H_t	&=	\gamma M I - (1 - \omega) \chi H - \omega \psi H, \\
 		R_t	&=	(1 - \gamma) M I + (1 - \omega) \chi H, \\
-		D_t	&=	\omega \chi H.
+		D_t	&=	\omega \psi H.
 	\end{align}

### 2021-10-27 update-10-22/update-10-22.tex
Covid progression graph with sources

@@
 \usepackage{url}
-\usepackage{natbib}
-\usepackage{pgfplots}
 
@@
 		\centering
-		\input{chicago-data}
+		\includegraphics[height=12cm]{chicago-data}
 		\label{fig:data}
-		\caption{no}
+		\caption{Timeline of the progression of COVID-19 in Chicago with key public policy events marked.
+			The COVID-19 data was obtained from the City of Chicago Data Portal \cite{Chicago-cases}.
+			The dates of the policy events were gathered from the Illinois.gov press releases \cite{phase-5}, \cite{mask-lift}, \cite{full-vax}, \cite{start-vax}, \cite{phase-4}, the Chicago Tribune \cite{phase-3}, NBC Chicago \cite{phase-2}}
 	\end{figure}
@@
 \bibliographystyle{abbrv}
-\bibliography{chicago-spatial-covid-draft}
+\bibliography{update-10-22}
 

### 2021-10-27 data/data_collection.py

@@ -34,10 +34,33 @@ df = df.loc["2020-03-01":"2021-10-26"]
 # Export to a csv
 df.to_csv("chicago_data.csv")
 
+# Create labels
+label_list = [
+    (pd.to_datetime("2020-03-21"), 'Shelter-in-place\nbegins.', 'm', 1000),
+    (pd.to_datetime("2020-04-24"), 'Mask mandate\nbegins.', 'brown', 1550),
+    (pd.to_datetime("2020-05-5"), 'Phase 2', 'c', 200),
+    (pd.to_datetime("2020-05-29"), 'Shelter-in-place\nends.', 'm', 1000),
+    (pd.to_datetime("2020-06-3"), 'Phase 3', 'c', 550),
+    (pd.to_datetime("2020-06-26"), 'Phase 4', 'c', 200),
+    (pd.to_datetime("2021-01-25"), 'Limited vaccination begins.', 'indigo', 1000),
+    (pd.to_datetime("2021-04-12"), 'General vaccination\nbegins.', 'indigo', 500),
+    (pd.to_datetime("2021-05-18"), 'Mask mandate\nends.', 'brown', 1050),
+    (pd.to_datetime("2021-06-11"), 'Phase 5', 'c', 200),
+    (pd.to_datetime("2021-08-30"), 'Mask mandate\nreinstated.', 'brown', 1000)
+]
+
 # Create plot
 df.plot(figsize=(12,8))
+
+ax = plt.gca()
+for date_point, label, clr, y in label_list:
+    plt.axvline(x=date_point, color=clr)
+    plt.text(date_point, ax.get_ylim()[1]-y, label,
+             horizontalalignment='center',
+             verticalalignment='center',
+             color=clr,
+             bbox=dict(facecolor='white', alpha=0.9))
 plt.show()
 
 # Export plot
-import tikzplotlib
-tikzplotlib.save("chicago-data.tex")
+plt.savefig('chicago-data.pdf')

### 2021-10-27 update-10-22/update-10-22.tex
Grammar correction.

@@
 			The COVID-19 data was obtained from the City of Chicago Data Portal \cite{Chicago-cases}.
-			The dates of the policy events were gathered from the Illinois.gov press releases \cite{phase-5}, \cite{mask-lift}, \cite{full-vax}, \cite{start-vax}, \cite{phase-4}, the Chicago Tribune \cite{phase-3}, NBC Chicago \cite{phase-2}}
+			The dates of the policy events were gathered from the Illinois.gov press releases \cite{phase-5}, \cite{mask-lift}, \cite{full-vax}, \cite{start-vax}, \cite{phase-4}, the Chicago Tribune \cite{phase-3}, and NBC Chicago \cite{phase-2}.}
 	\end{figure}

### 2021-10-28 update-10-22/update-10-22.tex
Updated parameters

@@
 \newcommand{\mat}[1]{\mathsf{#1}}
+\newcommand{\U}{\mathcal{U}}
 
@@
 		\centering
-		\caption{Population values for Chicago.}
+		\caption{Population values for Chicago.
+			Initial populations are determined from March 13, 2020.}
 		\label{tab:populations}
@@
 			Total population		&	$N$		&	2,695,598 \\
-			Initial infected		&	$I_0$	&	127	\\
-			Initial hospitalized	&	$H_0$	&	30 \\
-			Initial deceased		&	$D_0$	&	6 \\
+			Initial infected		&	$I_0$	&	162	\\
+			Initial hospitalized	&	$H_0$	&	38 \\
+			Initial deceased		&	$D_0$	&	3 \\
 			\hline
@@
 		\centering
-		\caption{Parameters for Chicago: optimal (best-fitting), median and interquartile range, and variation range used in the optimization algorithm.
+		\caption{Fitting parameters for Chicago: optimal (best-fitting), median and interquartile range, and variation range used in the optimization algorithm.
 			Initial parameter guesses were uniformly sampled within these ranges.}
@@
 			\hline
-			Population	&	$N$	&	2,695,598	& \\
-			Initial population	&	$(I_0, R_0)$	&	(127, 2)	& \\
-			Transmission rate, $S \rightarrow I$ [per day]	&	$\beta$\footnote{The transmission rate $\beta$ must be divided by $N$ when used in the ODE model.}	&	0.38206(0.38204-0.38209)	&	$c \in U[0,1]$ \\
-			Transition rate, $I \rightarrow R$ [per day]	&	$\gamma$	&	 0.39656(0.39654-0.39659)	&	$c \in U[0.25,0.75]$ \\
-			Diffusivity, $S$ [km$^2$/day]	&	$\D_S$	&	10	&	\\
-			Diffusivity, $I$ [km$^2$/day]	&	$\D_I$	&	100	&	\\
+			Transmission rate, $S \to I$ [per day]												&	$\beta_{SI}$	&									&	$c \in \U [0,1]$ \\
+			Transmission rate, $S \to A$ [per day]												&	$\beta_{SA}$	&									& 	$c \in \U [0,1]$ \\
+			Lockdown effect, $S \to I$															&	$\eta_{SI}$ 	&									& 	$c \in \U [0,1]$ \\
+			Lockdown effect, $S \to A$															&	$\eta_{SA}$		&									& 	$c \in \U [0,1]$ \\
+			Incubation period, $E \to I$ [days]													&	$1 / \sigma_I$	&									& 	$1 / k, k \in \U [2,7]$ \\
+			Latent period, $E \to A$ [days]														&	$1 / \sigma_A$	&									& 	$1 / k, k \in \U [2,7]$ \\
+			Infectivity period [days]															&	$1 / M$			&									& 	$1 / k, k \in \U [5,12]$ \\
+			Recovery period, $A \to AR$ [days]													&	$1 / M_{AR}$	&									& 	$1 / k, k \in \U [5,12]$ \\
+			Recovery period, $H \to R$ [days]													&	$1 / \chi$		&									& 	$1 / k, k \in \U [5,20]$ \\
+			Period to deceased, $H \to D$ [days]												&	$1 / \psi$		&									& 	$1 / k, k \in \U [5,20]$ \\
+			Conversion fraction $(I \xrightarrow{\gamma} H, \ I \xrightarrow{1 - \gamma} R)$	&	$\gamma$		&									& 	$c \in \U [0.25,0.75]$ \\
+			Conversion fraction $(H \xrightarrow{\omega} D, \ H \xrightarrow{1 - \omega} R)$	&	$\omega$		&									& 	$c \in \U [0.1,0.5]$ \\
+			Initial population fraction, exposed												&	$E_0 / I_0$		&									& 	$c \in \U [1,5]$ \\
+			Initial population fraction, asymptomatic											&	$A_0 / I_0$		&									& 	$c \in \U [1,5]$ \\
 			\hline

### 2021-10-28 data/time_series.py

@@ -7,7 +7,7 @@ df = pd.read_csv("chicago_data.csv", parse_dates=["Date"])
 df = df.set_index(["Date"])
 
 # Select the date range
-start = '2020-03-17' # start of data collection
+start = '2020-03-18' # start of data collection
 end = '2020-09-30' # end of first wave
 
 # Reform dataframe to date range

### 2021-10-29 data/time_series.py
Export time series to correct folder.

@@ -27,4 +27,4 @@ series["Cases"] = df["Daily Cases"] + df["Daily Hospitalizations"] + df["Cumulat
 series["Removed"] = df["Cumulative Deaths"]
 
 # Export new dataframe to a csv
-series.to_csv("../local-model/time_series.csv", header=False, index=False)
+series.to_csv("../update-10-22/time_series.csv", header=False, index=False)

### 2021-11-05 update-10-22/update-10-22.tex
Added citations for natural death rates.
Finalized ode model

@@
 \usepackage{url}
+\usepackage{hyperref}
 
@@
 \section{Model Setup}
+	We estimate the non-COVID death rate $\mu$ using the average United States death rate in urban regions from 2019 \cite{death-trends}.
+	That is, $\mu = 1.8997 \times 10^{-5}$ [per day].
+
+	\begin{table}[h]
+		\centering
+		\caption{Time sequence of events and simulation times.}
+		\label{tab:times}
+		\begin{tabular}{ c c c c }
+			\hline
+			\hline
+			Initial simulation time	&	Imposed lockdown	&	Effective lockdown	&	Last fitting day	\\
+			\hline
+			March 18, 2020	&	March 21, 2020	&	 April 1, 2020	&	June 24, 2020	\\
+			$t_i = 1$	&	$t_q = 4$	&	$t_q = 15$	&	$t_f = 99$ \\
+			\hline
+			\hline
+		\end{tabular}
+	\end{table}
 

### 2021-11-05 data/time_series.py

@@ -8,7 +8,7 @@ df = df.set_index(["Date"])
 
 # Select the date range
 start = '2020-03-18' # start of data collection
-end = '2020-09-30' # end of first wave
+end = '2020-06-24' # end of first wave
 
 # Reform dataframe to date range
 df = df.loc[start:end]
@@ -27,4 +27,4 @@ series["Cases"] = df["Daily Cases"] + df["Daily Hospitalizations"] + df["Cumulat
 series["Removed"] = df["Cumulative Deaths"]
 
 # Export new dataframe to a csv
-series.to_csv("../update-10-22/time_series.csv", header=False, index=False)
+series.to_csv("../version2/time_series.csv", header=False, index=False)

### 2021-11-05 local-model/odeModel.m

@@ -1,134 +1,0 @@
-% Clear windows and workspace
-clear
-clc
-close all
-
-%% Problem Definition
-% Population
-n = 2695598; %total population
-
-% Initial Conditions
-i0 = 127; %infected
-r0 = 2; %recovered
-s0 = n - i0 - r0; %susceptible
-y0 = [s0, i0, r0]; %initial conditions vector
-
-%% Observed Data
-% Load time, total cases and total deaths
-[t,Cobs,Dobs] = loadData();
-
-% Convert to a single vector
-Yobs = [Cobs(:); Dobs(:)];
-
-%% Model Solution
-% Define optimization iterations
-iters = 2000;
-
-% Parameter samples
-beta = rndSample([0,1],iters);
-gamma = rndSample([0.25,0.75],iters);
-
-% Initialize estimate vectors
-betaEst = zeros(1,iters);
-gammaEst = zeros(1,iters);
-
-% Sample each initial condition
-for i=1:iters
-    p0 = [beta(i)/n,gamma(i)];
-    % Generate parameter estimate
-    pEst = fminsearch(@(p)objfun(t,Yobs,p,y0),p0);
-    betaEst(i) = pEst(1);
-    gammaEst(i) = pEst(2);
-end
-
-% Calculate median values
-betaMed = median(betaEst);
-gammaMed = median(gammaEst);
-
-% Calculate IQR
-betaIQR = iqr(betaEst);
-gammaIQR = iqr(gammaEst);
-
-%% Present Results
-% Solve with median estimated parameters
-[~,Y] = ode45(@(t,y)sir(t,y,[betaMed,gammaMed]),t,y0);
-
-% Convert to SIR
-I = Y(:,2);
-R = Y(:,3);
-
-% Define total cases and deaths
-Cnum = I + R;
-Dnum = R;
-
-% Print parameters
-disp(['beta = ' num2str(betaMed * n) '(' num2str(n*(betaMed - betaIQR)) '-' num2str(n*(betaMed + betaIQR)) ')'])
-disp(['gamma = ' num2str(gammaMed) '(' num2str(gammaMed - gammaIQR) '-' num2str(gammaMed + gammaIQR) ')'])
-
-%% Plot results
-figure() %total cases plot
-plot(t,Cobs,'k.',t,Cnum,'b-')
-legend('$C_{obs}(t)$','$C_{num}(t)$','Interpreter','latex','Location','southeast')
[…]

### 2021-11-05 local-model/pdeModel.m

@@ -1,114 +1,0 @@
-%% Simple SIR in two spacial dimensions.
-
-clear;
-clc;
-close all;
-
-%% Initialize a two equation PDE model.
-
-model = createpde(2);
-
-%% Establish a model for the polygonal domain
-poly1 = [2 % 2 specifies we are forming a polygon
-    6 % Number of edges
-    1 % x1
-    1 % x2
-    2 % x3
-    2 % x4
-    0 % x5
-    0 % x6
-    1 % y1
-    0 % y2
-    0 % y3
-    2 % y4
-    2 % y5
-    1 % y6
-    ];
-
-gd = [poly1];
-
-ns = char('poly1');
-ns = ns';
-sf = 'poly1';
-[dl,bt] = decsg(gd,sf,ns);
-
-% Plot the computational domain
-fig1 = figure(1);
-pdegplot(dl,'EdgeLabels','on','FaceLabels','on')
-axis equal
-
-% Attach the computational domain to the PDE model
-geometryFromEdges(model,dl);
-% geometryFromEdges(model,@squareg);
-% 
-% pdegplot(model,'EdgeLabels','on')
-% ylim([-1.5,1.5])
-% axis equal
-
-%% Boundary Conditions
-%applyBoundaryCondition(model,'neumann','edge',1:6,'g',[0,0],'q',[0,0]);
-
-%% Specify coefficients for the PDE model
-% Define the parameters of the system
-Ds = 0.001;
-Di = 0.01;
-beta = 0.38206;
-gamma = 0.39656;
-
-% Based on these parameters, specify the PDE coefficients in the toolbox
-% format
-d = [1;1];
-c = [Ds;Di];
-f = @(location,state) [-beta * state.u(1,:) .* state.u(2,:);
-                       beta * state.u(1,:) .* state.u(2,:) - gamma * state.u(2,:)];
-
-specifyCoefficients(model,'m',0,'d',d,'c',c,'a',0,'f',f);
-
-%% Generate the mesh
-% Spacial domain
-generateMesh(model,'Hmax',0.1);
-% Plot the mesh
-fig2 = figure(2);
-pdeplot(model)
-axis equal
-
-% Time domain
-tlist = 0:0.1:200;
-
-%% Initial conditions
-setInitialConditions(model,@u0fun2);
-
-%% Solve the PDE
-model.SolverOptions.ReportStatistics='on';
-result = solvepde(model,tlist);
-
-%% Display the results
[…]

### 2021-11-05 odelModelComplete.m

@@ -1,215 +1,0 @@
-% Clear windows and workspace
-clear
-clc
-close all
-
-%% Problem Definition
-% Population
-n = 2695598; % Total population
-
-% Known initial conditions
-i0 = 99; % Infected
-h0 = 44; % Hospitalized
-d0 = 2; % Deceased
-
-% Initial Ratios
-e0i0 = 1; % Exposed/Infected
-a0i0 = 0.5; % Asymptomatic/Infected
-
-% Calculate initial conditions for other compartments
-e0 = e0i0 * i0; % Exposed
-a0 = a0i0 * i0; % Asymptomatic
-
-ar0 = 0; % Asymptomatic recovered
-r0 = 0; % Recovered
-
-s0 = n - (e0 + a0 + ar0 + i0 + h0 + r0 + d0); % Susceptible
-
-% Initial conditions vector
-y0 = [s0, e0, a0, ar0, i0, h0, r0, d0];
-
-%% Observed Data
-% Load time, total cases, and total deaths
-[t, Cobs, Dobs] = loadData();
-
-% Convert to a single vector
-Yobs = [Cobs(:); Dobs(:)];
-
-%% Model Solution
-% Define optimization iterations
-iters = 10;
-
-% Initialize estimate vectors
-beta_is = rndSample([0,1],iters); % Transmission rate, S -> I
-beta_as = rndSample([0,1],iters); % Transmission rate, S -> A
-sigma_a = 1./rndSample([2,7],iters); % Latent period, E -> A
-sigma_i = 1./rndSample([2,7],iters); % Incubation period, E -> A
-m = 1./rndSample([5,12],iters); % Infectivity period
-m_ar = 1./rndSample([5,12],iters); % Recovery period, A -> AR
-chi = 1./rndSample([5,20],iters); % Recovery period, H -> R
-psi = 1./rndSample([5,20],iters); % Period to deseased, H -> D
-gamma = rndSample([0.25,0.75],iters); % Conversion fraction, I -> H, I -> D
-omega = rndSample([0.1,0.5],iters); % Conversion fraction, H -> D, H -> R
-
-e0i0 = rndSample([1,6],iters); % Initial Exposed/Infected
-a0i0 = rndSample([1,6],iters); % Initial Asymptotic/Infected
-
-% Fixed Parameters
-mu = 5.79e-07;
-
-% Divide beta by n for ODE model
-beta_is = beta_is./n;
-beta_as = beta_as./n;
-
[…]

### 2021-11-05 version2/beta_asfun.m

@@ -1,11 +1,0 @@
-function beta = beta_asfun(p,t_q,t)
-    % Define parameters
-    beta_as = p(1);
-    eta_as = p(2);
-    
-    % Create a t_q vector
-    t_q = t_q * ones(length(t),1);
-    
-    % Calculate beta_is at time t
-    beta = beta_as * (eta_as + (1 - eta_as) * (1 - tanh(2 * (t -  t_q))) / 2);
-end

### 2021-11-05 version2/beta_isfun.m

@@ -1,11 +1,0 @@
-function beta = beta_isfun(p,t_q,t)
-    % Define parameters
-    beta_is = p(1);
-    eta_is = p(2);
-    
-    % Create a t_q vector
-    t_q = t_q * ones(1, length(t));
-    
-    % Calculate beta_is at time t
-    beta = beta_is * (eta_is + (1 - eta_is) * (1 - tanh(2 * (t -  t_q))) / 2);
-end

### 2021-11-05 version2/computemodel.m

@@ -1,0 +1,23 @@
+function [T,y] = computemodel(p,q,trange,pops)
+    % Separate the parameters from the population fractions
+    p1 = p(1:end-2); % Parameters
+    e0i0 = p(end-1); % Exposed/Infected
+    a0i0 = p(end); % Asymptotic/Infected
+
+    % Calculate the initial values
+    n = pops(1);
+    i0 = pops(2);
+    h0 = pops(3);
+    d0 = pops(4);
+    e0 = e0i0 * i0;
+    a0 = a0i0 * i0;
+    ar0 = 0;
+    r0 = 0;
+    s0 = n - (e0 + a0 + ar0 + i0 + h0 + r0 + d0);
+    
+    % Define the initial population vector
+    y0 = [s0, e0, a0, ar0, i0, h0, r0, d0];
+
+    % Compute the solution vector
+    [T,y] = ode45(@(t,y)sir(t,y,p1,q),trange,y0);
+end

### 2021-11-05 version2/main.m

@@ -1,0 +1,122 @@
+%% Main Program
+%
+%
+%% 
+clear
+close all
+clc
+
+%% Fixed Parameters
+
+% Population parameters
+n = 2695598; % Total population
+i0 = 160; % Initial infected
+h0 = 38; % Initial hospitalized
+d0 = 3; % Initial deceased
+
+% Non-COVID deaths
+mu = 1.8997e-5;
+
+% Lockdown start
+t_q = 4;
+
+%% Optimization Parameter Bounds
+
+% Lower bounds
+lb = [0, 0, 0, 0, 2, 2, 5, 5, 5, 5, 0.25, 0.1, 1, 1];
+
+% Upper bounds
+ub = [1, 1, 1, 1, 7, 7, 12, 12, 20, 20, 0.75, 0.5, 5, 5];
+
+%% Load Observed Data
+[time, Cobs, Dobs] = loadData("time_series.csv");
+
+% Create y vector
+Yobs = [Cobs(:); Dobs(:)];
+
+% Define end of simulation time
+t_f = length(time);
+
+%% Optimum Solution
+
+% Define number of iterations
+iters = 1;
+    
+% Define fixed parameters
+q = [mu, t_q];
+pops = [n, i0, h0, d0];
+
+% Linear constraints
+A = [];
+b = [];
+Aeq = [];
+beq = [];
+
+% Perform optimization iterations
+for iter=1:iters
+    tic % Start timer
+
+    % Sample starting parameters
+    p0 = sampleparams(lb,ub,n);
+    
+    % Compute parameter estimate
+    pEst(iter,:) = fminsearch(@(p)objfun(time,Yobs,p,q,pops),p0);
+
+    toc % End timer
+
+    disp("Iteration: " + num2str(iter))
+end
+
+% Compute the final solution vector from the median of the estimates
+pFin = median(pEst,1);
+
+% Compute the interquartile range
+iqrange = iqr(pEst,1);
+
+% Save the results to a file
+save('optimfile.mat','pFin','iqrange')
+
+% Compute the solution vector
+[T,y] = computemodel(pFin,q,time,pops);
+
+% Convert to system vectors
+i = y(:,5);
+h = y(:,6);
+r = y(:,7);
+d = y(:,8);
+
+% Define model cases and deaths
+Cnum = i + h + r + d;
+Dnum = d;
+
+% Convert to a single vector
+Ynum = [Cnum(:); Dnum(:)];
+
+%% Plot Results
[…]

### 2021-11-05 version2/objfun.m

@@ -1,13 +1,11 @@
-function obj = objfun(trange,Yobs,p,q,y0)
+function obj = objfun(trange,Yobs,p,q,pops)
 % OBJFUN   Compute objective function.
 %   obj = objfun(trange,Yobs,p,q,y0) compute norm2(Ynum - Yobs)
 %   where Ynum is the solution to sir(t,y,p1) over trange with
 %   initial conditions y0 and p1 = [p,q].
 
-    % Attach q to parameter vector
-    p1 = [p, q];
     % Compute the solution vector
-    [~,y] = ode45(@(t,y)sir(t,y,p1),trange,y0);
+    [~,y] = computemodel(p,q,trange,pops);
     
     % Convert to system vectors
     i = y(:,5);
@@ -23,5 +21,9 @@ function obj = objfun(trange,Yobs,p,q,y0)
     Ynum = [Cnum(:); Dnum(:)];
     
     % Compute the objective function
-    obj = norm(Ynum - Yobs);
+    try
+        obj = norm(Ynum - Yobs);
+    catch
+        obj = 100000;
+    end
 end

### 2021-11-05 version2/sampleparams.m

@@ -1,0 +1,35 @@
+function p0 = sampleparams(lb,ub,n)
+    % Transmission rates
+    beta_si = unifrnd(lb(1),ub(1)); % S -> I [per day]
+    beta_sa = unifrnd(lb(2),ub(2)); % S -> A [per day]
+    
+    % Lockdown effects
+    eta_si = unifrnd(lb(3),ub(3)); % S -> I
+    eta_sa = unifrnd(lb(4),ub(4)); % S -> A
+    
+    % Incubation period
+    sigma_i = 1 / unifrnd(lb(5),ub(5)); % E -> I [days]
+    
+    % Latent period
+    sigma_a = 1 / unifrnd(lb(6),ub(6)); % E -> I [days]
+    
+    % Infectivity period
+    m = 1 / unifrnd(lb(7),ub(7)); % [days]
+    
+    % Recovery periods
+    m_ar = 1 / unifrnd(lb(8),ub(8)); % A -> AR [days]
+    chi = 1 / unifrnd(lb(9),ub(9)); % H -> R [days]
+    
+    % Period to deceased
+    psi = 1 / unifrnd(lb(10),ub(10)); % H -> D [days]
+    
+    % Conversion fractions
+    gamma = unifrnd(lb(11),ub(11)); % I -> H, I -> R
+    omega = unifrnd(lb(12),ub(12)); % H -> D, H -> R
+    
+    % Initial population fractions
+    e0i0 = unifrnd(lb(13),ub(13)); % Exposed/Infected
+    a0i0 = unifrnd(lb(13),ub(13)); % Asymptomatic/Infected
+
+    p0 = [sigma_a, sigma_i, m_ar, m, gamma, omega, chi, psi, beta_sa / n, beta_si / n, eta_sa, eta_si, e0i0, a0i0];
+end

### 2021-11-05 version2/test.m

@@ -1,59 +1,0 @@
-clear
-clc
-close all
-
-n = 8414240;
-
-i0 = 297;
-h0 = 134;
-d0 = 6;
-
-mu = 5.79e-7;
-beta_is = 0.32 / n;
-beta_as = 0.28 / n;
-eta_is = 0.39;
-eta_as = 0.38;
-sigma_a = 1 / 2.97;
-sigma_i = 1 / 3.77;
-m = 1 / 5.97;
-m_ar = 1 / 6.85;
-chi = 1 / 6.74;
-psi = 1 / 8.33;
-gamma = 0.55;
-omega = 0.25;
-
-t_q = 16;
-
-e0i0 = 2.69;
-a0i0 = 2.93;
-
-e0 = e0i0 * i0;
-a0 = a0i0 * i0;
-
-ar0 = 0;
-r0 = 0;
-
-s0 = n - (e0 + a0 + ar0 + i0 + h0 + r0 + d0);
-
-p = [sigma_a, sigma_i, m_ar, m, gamma, omega, chi, psi, beta_as, beta_is, eta_as, eta_is];
-q = [mu, t_q];
-
-y0 = [s0, e0, a0, ar0, i0, h0, r0, d0];
-
-tspan = [1, 60];
-
-[T,Y] = ode45(@(t,y)sir(t,y,p,q),tspan,y0);
-
-s = Y(:,1);
-e = Y(:,2);
-a = Y(:,3);
-ar = Y(:,4);
-i = Y(:,5);
-h = Y(:,6);
-r = Y(:,7);
-d = Y(:,8);
-
-Cnum = i + h + r + d;
-
-figure()
-semilogy(T,Cnum)

### 2021-11-05 version2/optim.m
Separating optimizing script from results processing script

@@ -1,4 +1,4 @@
-%% Main Program
+%% Optimization Program
 %
 %
 %% 
@@ -18,7 +18,7 @@ d0 = 3; % Initial deceased
 mu = 1.8997e-5;
 
 % Lockdown start
-t_q = 4;
+t_q = 15;
 
 %% Optimization Parameter Bounds
 
@@ -40,7 +40,7 @@ t_f = length(time);
 %% Optimum Solution
 
 % Define number of iterations
-iters = 1;
+iters = 20;
     
 % Define fixed parameters
 q = [mu, t_q];
@@ -65,16 +65,18 @@ for iter=1:iters
     toc % End timer
 
     disp("Iteration: " + num2str(iter))
+    
+    % Compute the final solution vector from the median of the estimates
+    pFin = median(pEst,1);
+    
+    % Compute the interquartile range
+    iqrange = iqr(pEst,1);
+    
+    % Save the results to a file
+    save('optimfile.mat','pFin','iqrange')
 end
 
-% Compute the final solution vector from the median of the estimates
-pFin = median(pEst,1);
 
-% Compute the interquartile range
-iqrange = iqr(pEst,1);
-
-% Save the results to a file
-save('optimfile.mat','pFin','iqrange')
 
 % Compute the solution vector
 [T,y] = computemodel(pFin,q,time,pops);

### 2021-11-05 version2/optim.m
Included ode figures in report.

@@ -8,17 +8,8 @@ clc
 
 %% Fixed Parameters
 
-% Population parameters
-n = 2695598; % Total population
-i0 = 160; % Initial infected
-h0 = 38; % Initial hospitalized
-d0 = 3; % Initial deceased
-
-% Non-COVID deaths
-mu = 1.8997e-5;
-
-% Lockdown start
-t_q = 15;
+% Load fixed parameters
+load('params.mat');
 
 %% Optimization Parameter Bounds
 
@@ -40,21 +31,15 @@ t_f = length(time);
 %% Optimum Solution
 
 % Define number of iterations
-iters = 20;
+iters = 3;
     
 % Define fixed parameters
 q = [mu, t_q];
 pops = [n, i0, h0, d0];
 
-% Linear constraints
-A = [];
-b = [];
-Aeq = [];
-beq = [];
-
 % Perform optimization iterations
+tic % Start timer
 for iter=1:iters
-    tic % Start timer
 
     % Sample starting parameters
     p0 = sampleparams(lb,ub,n);
@@ -62,63 +47,16 @@ for iter=1:iters
     % Compute parameter estimate
     pEst(iter,:) = fminsearch(@(p)objfun(time,Yobs,p,q,pops),p0);
 
-    toc % End timer
-
     disp("Iteration: " + num2str(iter))
-    
-    % Compute the final solution vector from the median of the estimates
-    pFin = median(pEst,1);
-    
-    % Compute the interquartile range
-    iqrange = iqr(pEst,1);
-    
-    % Save the results to a file
-    save('optimfile.mat','pFin','iqrange')
 end
+toc % End timer
 
+%% Export Results
+% Compute the final solution vector from the median of the estimates
+pFin = median(pEst,1);
 
+% Compute the interquartile range
+iqrange = iqr(pEst,1);
 
-% Compute the solution vector
-[T,y] = computemodel(pFin,q,time,pops);
-
-% Convert to system vectors
-i = y(:,5);
-h = y(:,6);
-r = y(:,7);
-d = y(:,8);
-
-% Define model cases and deaths
-Cnum = i + h + r + d;
-Dnum = d;
-
-% Convert to a single vector
-Ynum = [Cnum(:); Dnum(:)];
-
-%% Plot Results
-% Total cases
-cFig = figure(1);
-semilogy(T,Cnum,time,Cobs)
[…]

### 2021-11-05 version2/params.m

@@ -1,0 +1,25 @@
+%% Define Fixed Parameters
+% Set the known parmaeters here.
+% These are exported to the other main programs.
+
+%% Clear MATLAB
+clear
+close all
+clc
+
+%% Set Values
+
+% Population parameters
+n = 2695598; % Total population
+i0 = 160; % Initial infected
+h0 = 38; % Initial hospitalized
+d0 = 3; % Initial deceased
+
+% Non-COVID deaths
+mu = 1.8997e-5;
+
+% Lockdown start
+t_q = 4;
+
+% Export fixed parameters
+save('params.mat')

### 2021-11-05 version2/results.m

@@ -1,0 +1,76 @@
+%% Results Program
+%
+%
+%% Clear MATLAB
+clear
+close all
+clc
+
+%% Fixed Parameters
+
+% Load fixed parameters
+load('params.mat');
+
+% Define fixed parameters
+q = [mu, t_q];
+pops = [n, i0, h0, d0];
+
+%% Load Optimized parameters
+load('optimfile.mat', 'pFin');
+
+%% Load Observed Data
+[time, Cobs, Dobs] = loadData("time_series.csv");
+
+% Create y vector
+Yobs = [Cobs(:); Dobs(:)];
+
+% Define end of simulation time
+t_f = length(time);
+
+%% Compute solution
+
+% Compute the solution vector
+[T,y] = computemodel(pFin,q,time,pops);
+
+% Convert to system vectors
+s = y(:,1);
+e = y(:,2);
+ar = y(:,3);
+a = y(:,4);
+i = y(:,5);
+h = y(:,6);
+r = y(:,7);
+d = y(:,8);
+
+% Define model cases and deaths
+Cnum = i + h + r + d;
+Dnum = d;
+
+%% Plot Results
+% Total cases
+cFig = figure(1);
+semilogy(T,Cnum,'r-',time,Cobs,'k.')
+legend('$C_{num}(t)$','$C_{obs}(t)$','Interpreter','latex','Location','southeast')
+xlabel('$t$ (days from March 18)','interpreter','latex')
+ylabel('$C(t)$','interpreter','latex')
+set(gca,'TickLabelInterpreter','latex')
+
+% Total deaths
+dFig = figure(2);
+semilogy(T,Dnum,'r-',time,Dobs,'k.')
+legend('$D_{num}(t)$','$D_{obs}(t)$','Interpreter','latex','Location','southeast')
+xlabel('$t$ (days from March 18)','interpreter','latex')
+ylabel('$D(t)$','interpreter','latex')
+set(gca,'TickLabelInterpreter','latex')
+
+% Export cases plot
+set(cFig,'Units','Inches');
+pos = get(cFig,'Position');
+set(cFig,'PaperPositionMode','Auto','PaperUnits','Inches','PaperSize',[pos(3), pos(4)])
+print(cFig,'~/Documents/GitHub/chicago-spatial-covid/update-10-22/cases','-dpdf','-r0')
+
+% Export deaths plot
+set(dFig,'Units','Inches');
+pos = get(dFig,'Position');
[…]

### 2021-11-05 version2/sampleparams.m

@@ -31,5 +31,13 @@ function p0 = sampleparams(lb,ub,n)
     e0i0 = unifrnd(lb(13),ub(13)); % Exposed/Infected
     a0i0 = unifrnd(lb(13),ub(13)); % Asymptomatic/Infected
 
+    if 0 < omega && omega <= 1
+        %disp('Condition 1 stable')
+    elseif 1 < omega && chi < (psi * omega) / (omega - 1)
+        %disp('Conditon 2 stable')
+    else
+        disp('Unstable')
+    end
+
     p0 = [sigma_a, sigma_i, m_ar, m, gamma, omega, chi, psi, beta_sa / n, beta_si / n, eta_sa, eta_si, e0i0, a0i0];
 end

### 2021-11-07 update-10-22/update-10-22.tex
Included ode figures in report.
Objective function updates

--- 2021-10-27 update-10-22/update-10-22.tex
@@
 \usepackage{url}
+\usepackage{hyperref}
+\usepackage{subcaption}
+\usepackage{multirow}
 
@@
 \usepackage{graphicx}
-\usepackage{subfig}
+%\usepackage{subfig}
 \graphicspath{ {./images/} }
@@
 		\centering
-		\includegraphics[height=12cm]{chicago-data}
+		\includegraphics[height=12cm]{chicago-data2}
 		\label{fig:data}
@@
 			The COVID-19 data was obtained from the City of Chicago Data Portal \cite{Chicago-cases}.
-			The dates of the policy events were gathered from the Illinois.gov press releases \cite{phase-5}, \cite{mask-lift}, \cite{full-vax}, \cite{start-vax}, \cite{phase-4}, the Chicago Tribune \cite{phase-3}, and NBC Chicago \cite{phase-2}.}
+			The dates of the policy events were gathered from the Illinois.gov press releases \cite{phase-5}, \cite{mask-lift}, \cite{full-vax}, \cite{start-vax}, \cite{phase-4}, the Chicago Tribune \cite{phase-3}, and NBC Chicago \cite{phase-2}.
+			Note the logarithmic scale.}
 	\end{figure}
@@
 \section{Model Setup}
+	We begin by explaining the ODE model.
+	This is obtained from the full PDE model in Equations (\ref{eq:s_t})-(\ref{eq:d_t}) by simply removing the diffusion terms.
+
+	We extend the standard SIR model to incorporate key aspects of the COVID-19 virus.
+
+	Given the time scale of the pandemic, we model a weak effect of the change in the susceptible population from births or other mortality factors with the term $- \mu S$.
+	We estimate $\mu = 1.8997 \times 10^{-5}$ [per day] using the average United States death rate in urban regions from 2019 \cite{death-trends}.
+
+	\begin{table}[h]
+		\centering
+		\caption{Time sequence of events and simulation times.}
+		\label{tab:times}
+		\begin{tabular}{ c c c c }
+			\hline
+			\hline
+			Initial simulation time	&	Imposed lockdown	&	Effective lockdown	&	Last fitting day	\\
+			\hline
+			March 18, 2020	&	March 21, 2020	&	 April 1, 2020	&	June 24, 2020	\\
+			$t_i = 1$	&	$t_q = 4$	&	$t_q = 15$	&	$t_f = 99$ \\
+			\hline
+			\hline
+		\end{tabular}
+	\end{table}
 
@@
 	\begin{align}
+		\label{eq:s_t}
 		S_t &=	\D_S \Delta S - \beta_{SA} S A - \beta_{SI} S I - \mu S, \\
+		\label{eq:e_t}
 		E_t	&=	\D_E \Delta E + \beta_{SA} S A + \beta_{SI} S I - (\sigma_A + \sigma_I) E, \\
+		\label{eq:ar_t}
 		AR_t &= M_{AR} A, \\
+		\label{eq:a_t}
 		A_t	&=	\D_A \Delta A + \sigma_A E - M_{AR} A, \\
+		\label{eq:i_t}
 		I_t	&=	\sigma_I E - M I, \\
+		\label{eq:h_t}
 		H_t	&=	\gamma M I - (1 - \omega) \chi H - \omega \psi H, \\
+		\label{eq:r_t}
 		R_t	&=	(1 - \gamma) M I + (1 - \omega) \chi H, \\
+		\label{eq:d_t}
 		D_t	&=	\omega \psi H.
@@
 
+	To create the PDE model, we first must find suitable parameters using the ODE model as in previous studies.
+	We utilize the MATLAB nonlinear optimization algorithm \verb|fminsearch| for this purpose.
+	The optimal parameters are determined by minimizing the Euclidean distance $\mathcal{N}$ between the time series generated by the model (subscript ``num'') and the corresponding observed data time series (subscript ``obs'')
+	\begin{equation} \label{eq:objfun}
+		\mathcal{N} = \sum_i \left( \left| \ln C_\mathrm{num} (t_i) - \ln C_\mathrm{obs} (t_i) \right| + \left| \ln D_\mathrm{num} (t_i) - \ln D_\mathrm{obs} (t_i) \right| \right)
+	\end{equation}
+	where the index $i$ identifies a point in the time series.
+	The parameters are chosen to reproduce the time series of the total number of cases $C(t) = I(t) + H(t) + R(t) + D(t)$, and the total number of deceased $D(t)$.
+
 	\begin{table}[h]
 		\centering
-		\caption{Population values for Chicago.}
+		\caption{Population values for Chicago.
+			Initial populations are determined from March 13, 2020.}
 		\label{tab:populations}
@@
 
+	To account for changes in virus transmission due to the shelter-in-place order, we impose a time dependence on the transmission rates $\beta$ as in Equations (\ref{eq:beta-si}) and (\ref{eq:beta-sa}).
+	\begin{align}
+		\label{eq:beta-si}
+		\beta_{SI} (t) &= \beta_{SI} \left( \eta_{SI} + (1 - \eta_{SI}) \frac{1 - \tanh [2(t - t_q)]}{2} \right) \\
+		\label{eq:beta-sa}
+		\beta_{SA} (t) &= \beta_{SA} \left( \eta_{SA} + (1 - \eta_{SA}) \frac{1 - \tanh [2(t - t_q)]}{2} \right)
+	\end{align}
+
 	\begin{table}[h]
 		\centering
-		\caption{Parameters for Chicago: optimal (best-fitting), median and interquartile range, and variation range used in the optimization algorithm.
+		\caption{ODE parameters: optimal (best-fitting), median and interquartile range, and variation range used in the optimization algorithm.
 			Initial parameter guesses were uniformly sampled within these ranges.}
@@
 			\hline
-			&	&	Median (interquartile range)	&	Initial value \\
+																								&					&	Median (IQR)	&	Initial value \\
 			\hline
-			Population	&	$N$	&	2,695,598	& \\
-			Initial population	&	$(I_0, R_0)$	&	(127, 2)	& \\
-			Transmission rate, $S \rightarrow I$ [per day]	&	$\beta$\footnote{The transmission rate $\beta$ must be divided by $N$ when used in the ODE model.}	&	0.38206(0.38204-0.38209)	&	$c \in U[0,1]$ \\
-			Transition rate, $I \rightarrow R$ [per day]	&	$\gamma$	&	 0.39656(0.39654-0.39659)	&	$c \in U[0.25,0.75]$ \\
-			Diffusivity, $S$ [km$^2$/day]	&	$\D_S$	&	10	&	\\
-			Diffusivity, $I$ [km$^2$/day]	&	$\D_I$	&	100	&	\\
+			Transmission rate, $S \to I$ [per day]\textsuperscript{a}							&	$\beta_{SI}$	&									&	$c \in \U [0,1]$ \\
+			Transmission rate, $S \to A$ [per day]\textsuperscript{a}							&	$\beta_{SA}$	&									& 	$c \in \U [0,1]$ \\
+			Lockdown effect, $S \to I$															&	$\eta_{SI}$ 	&									& 	$c \in \U [0,1]$ \\
+			Lockdown effect, $S \to A$															&	$\eta_{SA}$		&									& 	$c \in \U [0,1]$ \\
+			Incubation period, $E \to I$ [days]													&	$1 / \sigma_I$	&									& 	$1 / k, k \in \U [2,7]$ \\
+			Latent period, $E \to A$ [days]														&	$1 / \sigma_A$	&									& 	$1 / k, k \in \U [2,7]$ \\
+			Infectivity period [days]															&	$1 / M$			&									& 	$1 / k, k \in \U [5,12]$ \\
+			Recovery period, $A \to AR$ [days]													&	$1 / M_{AR}$	&									& 	$1 / k, k \in \U [5,12]$ \\
+			Recovery period, $H \to R$ [days]													&	$1 / \chi$		&									& 	$1 / k, k \in \U [5,20]$ \\
+			Period to deceased, $H \to D$ [days]												&	$1 / \psi$		&									& 	$1 / k, k \in \U [5,20]$ \\
+			Conversion fraction $(I \xrightarrow{\gamma} H, \ I \xrightarrow{1 - \gamma} R)$	&	$\gamma$		&									& 	$c \in \U [0.25,0.75]$ \\
+			Conversion fraction $(H \xrightarrow{\omega} D, \ H \xrightarrow{1 - \omega} R)$	&	$\omega$		&									& 	$c \in \U [0.1,0.5]$ \\
+			Initial population fraction, exposed												&	$E_0 / I_0$		&									& 	$c \in \U [1,5]$ \\
+			Initial population fraction, asymptomatic											&	$A_0 / I_0$		&									& 	$c \in \U [1,5]$ \\
 			\hline
 			\hline
+			\multicolumn{4}{c}{\small \textsuperscript{a} Note: The transmission rates $\beta$ must be divided by $N$ when used in the ODE model.}
 		\end{tabular}
@@
 
+	\begin{figure}[h!]
+		\centering
+		\begin{subfigure}{0.5\textwidth}
+			\centering
+			\includegraphics[height=6cm]{cases}
+			\label{fig:ode-cases}
+			\caption{Confirmed cases $C(t) = I(t) + R(t) + H(t) + D(t)$.}
+		\end{subfigure}%
+		\begin{subfigure}{0.5\textwidth}
+			\centering
+			\includegraphics[height=6cm]{deaths}
+			\label{fig:ode-deaths}
+			\caption{Number of deaths $D(t)$.}
+		\end{subfigure}
+		\caption{ODE model with fitting to official data from March 18, 2020 ($t_i = 1$) to June 24, 2020 ($t_f = 99$).
+		Here we show the case when $\beta$ changes exactly on the imposed lockdown on March 21, 2020 ($t_q = 4$).}
+		\label{fig:ode-results}
+	\end{figure}
+
 \section{ODE Dynamics}
-	We want to understand the trajectories of the dynamics of the ODE system under different initial conditions.
+	We want to understand the trajectories of the dynamics of the ODE system (\ref{eq:s_t})-(\ref{eq:d_t}) under different initial conditions.
 	To do this we first find the equilibrium points by solving
 	$$S_t = E_t = A_t = I_t = H_t = R_t = D_t = 0$$
-	simultaneously for $\vec{x} = (S, E, A, AR, I, H, R, D)$.
-	The solutions of this system are of the form $\vec{x^*} = (0, 0 , 0, AR, 0, 0, R, D)$.
+	simultaneously for $\vec{x} = (S, E, A, AR, I, H, R, D)^\intercal$.
+	The solutions of this system are of the form $\vec{x^*} = (0, 0 , 0, AR, 0, 0, R, D)^\intercal$.
 	This implies there are infinitely many non-isolated equilibrium points.
@@
 			0									&	0						&	0				&	0	&	0				&	\psi \omega							&	0	&	0
-		\end{pmatrix}
+		\end{pmatrix}.
 	\end{equation}
@@
 	That is, whenever we have either of these conditions the equilibrium points are stable.
-	We call this situation \textit{endemic}.
-	If $\lambda_8 > 0$, the equilibrium points are unstable and the situation is an \textit{epidemic}.
+	If $\lambda_8 > 0$, the equilibrium points are unstable.
+
+	We can further analyze the evolution of the pandemic by calculating the basic reproduction number $R_0$.
+	We use the next generation matrix approach of the system (\ref{eq:s_t})-(\ref{eq:d_t}) without the spatial terms, as in \cite{Diekmann-1990} and \cite{VANDENDRIESSCHE200229}.
+	In particular, we rewrite the model in the form $\vec{x}_t = \vec{F} - \vec{V}$.
+	The components $F_i$ represents the rate of appearance of new infections in compartment $i$.
+	The vector $\vec{V} = \vec{V}^- - \vec{V}^+$, where $V_i^+$ represents the rate of transfer of individuals into compartment $i$ by all other means, and $V_i^-$ represents the rate of transfer of individuals out of compartment $i$.
+	Reordering the compartments so $\vec{x'_t} = (E_t, A_t, I_t, H_t, S_t, AR_t, R_t, D_t)^\intercal$, we have
+	\begin{align*}
+		\vec{F} &= \begin{pmatrix}
+			\beta_{SA} S A + \beta_{SI} S I \\
+			0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0
+		\end{pmatrix}, &
+		\vec{V} &= \begin{pmatrix}
+			(\sigma_A +\sigma_I) E \\
+			- \sigma_A E + M_{AR} A \\
+			- \sigma_I E + M I \\
+			- \gamma M I + (1 - \omega) \chi H + \omega \psi H \\
+			\beta_{SA} S A + \beta_{SI} S I + \mu S \\
+			M_{AR} A \\
+			- (1 - \gamma) M I - (1 - \omega) \chi H \\
+			- \omega \psi H
+		\end{pmatrix}.
+	\end{align*}
+	We focus on just the infectious/infected compartments, $E, A, I, H$, and find the Jacobians of $\vec{F}$ and $\vec{V}$ with respect to these populations in the order in which they appear.
+	Evaluating at the the disease-free equilibrium ($S = S^*, \ E = AR = A = I = H = R = D = 0$) yields
+	\begin{align*}
+		\mat{F} &= \begin{pmatrix}
+			0 & \beta_{SA} S^* & \beta_{SI} S^* & 0 \\
+			0 & 0 & 0 & 0 \\
+			0 & 0 & 0 & 0 \\
+			0 & 0 & 0 & 0
+		\end{pmatrix}, &
+		\mat{V} &= \begin{pmatrix}
+			\sigma_A + \sigma_I & 0 & 0 & 0 \\
+			- \sigma_A & M_{AR} & 0 & 0 \\
+			- \sigma_I & 0 & M & 0 \\
+			0 & 0 & - \gamma M & \chi (1 - \omega) + \psi \omega
+		\end{pmatrix}.
+	\end{align*}
+	The matrix $\mat{F} \mat{V}^{-1}$ is the next-generation matrix.
+	Then $R_0 = \rho (\mat{F} \mat{V}^{-1})$, which is
+	\begin{equation} \label{eq:r0}
+		R_0 = \frac{\beta_{SA} S^* \sigma_A}{m_{AR} (\sigma_A + \sigma_I)} + \frac{\beta_{SI} S^* \sigma_I}{M (\sigma_A + \sigma_I)}.
+	\end{equation}
 

### 2021-11-07 data/data_collection.py

@@ -36,21 +36,21 @@ df.to_csv("chicago_data.csv")
 
 # Create labels
 label_list = [
-    (pd.to_datetime("2020-03-21"), 'Shelter-in-place\nbegins.', 'm', 1000),
-    (pd.to_datetime("2020-04-24"), 'Mask mandate\nbegins.', 'brown', 1550),
-    (pd.to_datetime("2020-05-5"), 'Phase 2', 'c', 200),
-    (pd.to_datetime("2020-05-29"), 'Shelter-in-place\nends.', 'm', 1000),
-    (pd.to_datetime("2020-06-3"), 'Phase 3', 'c', 550),
-    (pd.to_datetime("2020-06-26"), 'Phase 4', 'c', 200),
-    (pd.to_datetime("2021-01-25"), 'Limited vaccination begins.', 'indigo', 1000),
-    (pd.to_datetime("2021-04-12"), 'General vaccination\nbegins.', 'indigo', 500),
-    (pd.to_datetime("2021-05-18"), 'Mask mandate\nends.', 'brown', 1050),
-    (pd.to_datetime("2021-06-11"), 'Phase 5', 'c', 200),
-    (pd.to_datetime("2021-08-30"), 'Mask mandate\nreinstated.', 'brown', 1000)
+    (pd.to_datetime("2020-03-21"), 'Shelter-in-place\nbegins.', 'm', 6000),
+    (pd.to_datetime("2020-04-24"), 'Mask mandate\nbegins.', 'brown', 7600),
+    (pd.to_datetime("2020-05-5"), 'Phase 2', 'c', 1000),
+    (pd.to_datetime("2020-05-29"), 'Shelter-in-place\nends.', 'm', 6000),
+    (pd.to_datetime("2020-06-3"), 'Phase 3', 'c', 3700),
+    (pd.to_datetime("2020-06-26"), 'Phase 4', 'c', 1000),
+    (pd.to_datetime("2021-01-25"), 'Limited vaccination begins.', 'indigo', 2000),
+    (pd.to_datetime("2021-04-12"), 'General vaccination\nbegins.', 'indigo', 7600),
+    (pd.to_datetime("2021-05-18"), 'Mask mandate\nends.', 'brown', 5600),
+    (pd.to_datetime("2021-06-11"), 'Phase 5', 'c', 2000),
+    (pd.to_datetime("2021-08-30"), 'Mask mandate\nreinstated.', 'brown', 5600)
 ]
 
[…]

### 2021-11-07 version2/objfun.m

@@ -1,4 +1,4 @@
-function obj = objfun(trange,Yobs,p,q,pops)
+function obj = objfun(trange,Cobs,Dobs,p,q,pops)
 % OBJFUN   Compute objective function.
 %   obj = objfun(trange,Yobs,p,q,y0) compute norm2(Ynum - Yobs)
 %   where Ynum is the solution to sir(t,y,p1) over trange with
@@ -17,14 +17,6 @@ function obj = objfun(trange,Yobs,p,q,pops)
     Cnum = i + h + r + d;
     Dnum = d;
     
-    % Convert to a single vector
-    Ynum = [Cnum(:); Dnum(:)];
-    
     % Compute the objective function
-    try
-        %obj = norm((Cnum - Cobs).^2 + (Dnum - Dobs).^2);
-        obj = norm(Ynum - Yobs);
-    catch
-        obj = 100000;
-    end
+    obj = sum(vecnorm(log(Cnum) - log(Cobs)) + vecnorm(log(Dnum) - log(Dobs)));
 end

### 2021-11-07 version2/optim.m

@@ -12,6 +12,9 @@ clc
 load('params.mat');
 
 %% Optimization Parameter Bounds
+% Bounds are in the form:
+% [beta_si, beta_sa, eta_si, eta_sa, sigma_i, sigma_a, m, m_ar, chi, psi,
+% gamma, omega, e0/i0, a0/i0]
 
 % Lower bounds
 lb = [0, 0, 0, 0, 2, 2, 5, 5, 5, 5, 0.25, 0.1, 1, 1];
@@ -31,11 +34,14 @@ t_f = length(time);
 %% Optimum Solution
 
 % Define number of iterations
-iters = 3;
+iters = 1;
     
 % Define fixed parameters
 q = [mu, t_q];
 pops = [n, i0, h0, d0];
+
+% Preallocate the estimate matrix
+pEst = zeros(iters,length(lb));
 
 % Perform optimization iterations
 tic % Start timer
@@ -45,7 +51,7 @@ for iter=1:iters
     p0 = sampleparams(lb,ub,n);
     
     % Compute parameter estimate
-    pEst(iter,:) = fminsearch(@(p)objfun(time,Yobs,p,q,pops),p0);
+    pEst(iter,:) = fminsearch(@(p)objfun(time,Cobs,Dobs,p,q,pops),p0);
 
     disp("Iteration: " + num2str(iter))
 end

### 2021-11-07 version2/results.m

@@ -21,9 +21,6 @@ load('optimfile.mat', 'pFin');
 %% Load Observed Data
 [time, Cobs, Dobs] = loadData("time_series.csv");
 
-% Create y vector
-Yobs = [Cobs(:); Dobs(:)];
-
 % Define end of simulation time
 t_f = length(time);
 
@@ -50,7 +47,7 @@ Dnum = d;
 % Total cases
 cFig = figure(1);
 semilogy(T,Cnum,'r-',time,Cobs,'k.')
-legend('$C_{num}(t)$','$C_{obs}(t)$','Interpreter','latex','Location','southeast')
+legend('$C_\mathrm{num}(t)$','$C_\mathrm{obs}(t)$','Interpreter','latex','Location','southeast')
 xlabel('$t$ (days from March 18)','interpreter','latex')
 ylabel('$C(t)$','interpreter','latex')
 set(gca,'TickLabelInterpreter','latex')
@@ -58,7 +55,7 @@ set(gca,'TickLabelInterpreter','latex')
 % Total deaths
 dFig = figure(2);
 semilogy(T,Dnum,'r-',time,Dobs,'k.')
-legend('$D_{num}(t)$','$D_{obs}(t)$','Interpreter','latex','Location','southeast')
+legend('$D_\mathrm{num}(t)$','$D_\mathrm{obs}(t)$','Interpreter','latex','Location','southeast')
 xlabel('$t$ (days from March 18)','interpreter','latex')
 ylabel('$D(t)$','interpreter','latex')
 set(gca,'TickLabelInterpreter','latex')

### 2021-11-11 update-10-22/update-10-22.tex
Introduce non-dimensional parameters

@@
 \usepackage{subcaption}
-\usepackage{multirow}
 
@@
 
+	\begin{align*}
+		S_t &=	\D_S \Delta S - \frac{\beta_{A} A}{N} S - \frac{\beta_{I} I}{N} S - \mu S, \\
+		E_t	&=	\D_E \Delta E + \frac{\beta_{A} A}{N} S + \frac{\beta_{I} I}{N} S - (\sigma_A + \sigma_I) E, \\
+		AR_t &= M_{AR} A, \\
+		A_t	&=	\D_A \Delta A + \sigma_A E - M_{AR} A, \\
+		I_t	&=	\sigma_I E - M I, \\
+		H_t	&=	\gamma M I - (1 - \omega) \chi H - \omega \psi H, \\
+		R_t	&=	(1 - \gamma) M I + (1 - \omega) \chi H, \\
+		D_t	&=	\omega \psi H.
+	\end{align*}
+
+	We begin by reducing the model.
+	First, we ignore non-COVID deaths, so $\mu = 0$.
+	Next we introduce the fractions
+	\begin{align*}
+		s &= \frac{S}{N}, & e &= \frac{E}{N}, & ar &= \frac{AR}{N}, & a &= \frac{A}{N}, \\
+		i &= \frac{I}{N}, & h &= \frac{H}{N}, & r &= \frac{R}{N}, & d &= \frac{D}{N}.
+	\end{align*}
+	With these fractions and $\mu = 0$, we have the conservation law
+	\begin{equation} \label{eq:conservation}
+		s + e + ar + a + i + h + r + d = 1.
+	\end{equation}
+	The ODE system becomes
+	\begin{align*}
+		s_t &=	 - \beta_{A} a s - \beta_{I} i s, \\
+		e_t	&=	\beta_{A} a s + \beta_{I} i s - (\sigma_A + \sigma_I) e, \\
+		ar_t &= M_{AR} a, \\
+		a_t	&=	\sigma_A e - M_{AR} a, \\
+		i_t	&=	\sigma_I e - M i, \\
+		h_t	&=	\gamma M i - (1 - \omega) \chi h - \omega \psi h, \\
+		r_t	&=	(1 - \gamma) M i + (1 - \omega) \chi h, \\
+		d_t	&=	\omega \psi h.
+	\end{align*}
+	At this point it is worth noting the dimensions of the components of the system.
+	Since $[X] = \mathrm{population}$ where $X \in \{N, S, E, AR, A, I, H, R, D\}$, then $[x] = 1$ where $x \in \{s, e, ar, a, i, h, r, d\}$.
+	Then,
+	$$[\beta_A] = [\beta_I] = [\sigma_A] = [\sigma_I] = [M_{AR}] = [M] = [\chi] = [\psi] = \frac{1}{T}$$
+	and $[\omega] = [\gamma] = 1$.
+	We introduce the non-dimensional time variable $\tau = t M$.
+	Then by the chain rule, $x_t = M x_\tau$.
+	Next, we define the following non-dimensional parameters
+	\begin{align*}
+		\alpha_A &= \frac{\beta_A}{M}, & \alpha_I &= \frac{\beta_I}{M}, & \lambda_A &= \frac{\sigma_A}{M}, & \lambda_I &= \frac{\sigma_I}{M}
+	\end{align*}
+
+	Using these non-dimensional parameters and the conservation law in Equation \ref{eq:conservation}, we can replace the ODE system with
 	\begin{align}
-		\label{eq:s_t}
-		S_t &=	\D_S \Delta S - \beta_{SA} S A - \beta_{SI} S I - \mu S, \\
-		\label{eq:e_t}
-		E_t	&=	\D_E \Delta E + \beta_{SA} S A + \beta_{SI} S I - (\sigma_A + \sigma_I) E, \\
-		\label{eq:ar_t}
-		AR_t &= M_{AR} A, \\
-		\label{eq:a_t}
-		A_t	&=	\D_A \Delta A + \sigma_A E - M_{AR} A, \\
-		\label{eq:i_t}
-		I_t	&=	\sigma_I E - M I, \\
-		\label{eq:h_t}
-		H_t	&=	\gamma M I - (1 - \omega) \chi H - \omega \psi H, \\
-		\label{eq:r_t}
-		R_t	&=	(1 - \gamma) M I + (1 - \omega) \chi H, \\
-		\label{eq:d_t}
-		D_t	&=	\omega \psi H.
+		s_\tau &= - \alpha_A a s - \alpha_I i s, \\
+		e_\tau &= \alpha_A a s + \alpha_I i s - (\lambda_A + \lambda_I) e, \\
+		ar_\tau &= \xi a, \\
+		a_\tau &= \lambda_A e - \xi a, \\
+		i_\tau &= \lambda_I e - i
 	\end{align}
@@
 			\hline
-			Transmission rate, $S \to I$ [per day]\textsuperscript{a}							&	$\beta_{SI}$	&									&	$c \in \U [0,1]$ \\
-			Transmission rate, $S \to A$ [per day]\textsuperscript{a}							&	$\beta_{SA}$	&									& 	$c \in \U [0,1]$ \\
+			Transmission rate, $S \to I$							&	$\beta_{SI}$	&									&	$c \in \U [0,1]$ \\
+			Transmission rate, $S \to A$							&	$\beta_{SA}$	&									& 	$c \in \U [0,1]$ \\
 			Lockdown effect, $S \to I$															&	$\eta_{SI}$ 	&									& 	$c \in \U [0,1]$ \\
@@
 			\hline
-			\multicolumn{4}{c}{\small \textsuperscript{a} Note: The transmission rates $\beta$ must be divided by $N$ when used in the ODE model.}
 		\end{tabular}

### 2021-11-11 version2/optim.m

@@ -48,10 +48,11 @@ tic % Start timer
 for iter=1:iters
 
     % Sample starting parameters
-    p0 = sampleparams(lb,ub,n);
+    %p0 = sampleparams(lb,ub,n);
+    p0 = (lb + ub) / 2;
     
     % Compute parameter estimate
-    pEst(iter,:) = fminsearch(@(p)objfun(time,Cobs,Dobs,p,q,pops),p0);
+    pEst(iter,:) = fmincon(@(p)objfun(time,Cobs,Dobs,p,q,pops),p0,[],[],[],[],lb,ub);
 
     disp("Iteration: " + num2str(iter))
 end

### 2021-11-12 update-10-22/update-10-22.tex
Non-dimensional parameters and equations

@@
 	\begin{align*}
-		\alpha_A &= \frac{\beta_A}{M}, & \alpha_I &= \frac{\beta_I}{M}, & \lambda_A &= \frac{\sigma_A}{M}, & \lambda_I &= \frac{\sigma_I}{M}
+		\alpha_A &= \frac{\beta_A}{M}, & \alpha_I &= \frac{\beta_I}{M}, & \lambda_A &= \frac{\sigma_A}{M}, & \lambda_I &= \frac{\sigma_I}{M}, \\
+		\eta &= \frac{M_{AR}}{M}, & \theta &= \frac{\chi}{M}, & \kappa &= \frac{\psi}{M}.
 	\end{align*}
@@
 	\begin{align}
+		\label{eq:s_t}
 		s_\tau &= - \alpha_A a s - \alpha_I i s, \\
+		\label{eq:e_t}
 		e_\tau &= \alpha_A a s + \alpha_I i s - (\lambda_A + \lambda_I) e, \\
+		\label{eq:ar_t}
 		ar_\tau &= \xi a, \\
+		\label{eq:a_t}
 		a_\tau &= \lambda_A e - \xi a, \\
-		i_\tau &= \lambda_I e - i
+		\label{eq:i_t}
+		i_\tau &= \lambda_I e - i, \\
+		\label{eq:r_t}
+		r_\tau &= (1 - \gamma) i + (1 - \omega) \theta h.
 	\end{align}
+	Then by the conservation law, we have
+	\begin{equation} \label{eq:d_t}
+		d = n - (s + e + ar + a + i + r).
+	\end{equation}
 
@@
 		\label{eq:beta-si}
-		\beta_{SI} (t) &= \beta_{SI} \left( \eta_{SI} + (1 - \eta_{SI}) \frac{1 - \tanh [2(t - t_q)]}{2} \right) \\
+		\beta_{I} (t) &= \beta_{I} \left( \eta_{I} + (1 - \eta_{I}) \frac{1 - \tanh [2(t - t_q)]}{2} \right) \\
 		\label{eq:beta-sa}
-		\beta_{SA} (t) &= \beta_{SA} \left( \eta_{SA} + (1 - \eta_{SA}) \frac{1 - \tanh [2(t - t_q)]}{2} \right)
+		\beta_{A} (t) &= \beta_{A} \left( \eta_{A} + (1 - \eta_{A}) \frac{1 - \tanh [2(t - t_q)]}{2} \right)
 	\end{align}
@@
 			\hline
-			Transmission rate, $S \to I$							&	$\beta_{SI}$	&									&	$c \in \U [0,1]$ \\
-			Transmission rate, $S \to A$							&	$\beta_{SA}$	&									& 	$c \in \U [0,1]$ \\
-			Lockdown effect, $S \to I$															&	$\eta_{SI}$ 	&									& 	$c \in \U [0,1]$ \\
-			Lockdown effect, $S \to A$															&	$\eta_{SA}$		&									& 	$c \in \U [0,1]$ \\
+			Transmission rate, $S \to I$														&	$\beta_{I}$	&									&	$c \in \U [0,1]$ \\
+			Transmission rate, $S \to A$														&	$\beta_{A}$	&									& 	$c \in \U [0,1]$ \\
+			Lockdown effect, $S \to I$															&	$\eta_{I}$ 	&									& 	$c \in \U [0,1]$ \\
+			Lockdown effect, $S \to A$															&	$\eta_{A}$		&									& 	$c \in \U [0,1]$ \\
 			Incubation period, $E \to I$ [days]													&	$1 / \sigma_I$	&									& 	$1 / k, k \in \U [2,7]$ \\

### 2021-11-12 update-10-22/update-10-22.tex
Include pde model domain and mesh.

@@
 		\centering
-		\includegraphics[height=12cm]{chicago-data2}
-		\label{fig:data}
+		\includegraphics[height=11cm]{chicago-data2}
+
 		\caption{Timeline of the progression of COVID-19 in Chicago with key public policy events marked.
@@
 			Note the logarithmic scale.}
+		\label{fig:data}
 	\end{figure}
@@
 \section{Model Setup}
-	We begin by explaining the ODE model.
-	This is obtained from the full PDE model in Equations (\ref{eq:s_t})-(\ref{eq:d_t}) by simply removing the diffusion terms.
+	The formulation presented here extends the standard SIR model to incorporate some key aspects of the COVID-19 virus.
+	A schematic of the model is given in Figure \ref{fig:model}.
 
-	We extend the standard SIR model to incorporate key aspects of the COVID-19 virus.
+	We begin with a population of \textit{susceptibles} $S$, which may become \textit{exposed} $E$ due to the appearance of the virus in the population.
+	This represents the feature of COVID-19 to remain latent within the host for some time before the host becomes infectious.
+	After the latent period, exposed individuals may become \textit{asymptomatically infectious} $A$ or \textit{symptomatically infectious} $I$.
+	We assume both $A$ and $I$ can interact with $S$ to draw new members into $E$.
+	A fraction of hosts in $I$ may require hospitalization, generating the \textit{hospitalized} $H$ population.
+	A fraction of the $H$ population is successfully treated, moving to the \textit{recovered} $R$ population.
+	The remaining population in $H$ do not recover, leading the the population of \textit{deceased} $D$.
+	Meanwhile, asymptomatic hosts are assumed to all recover, moving to the \textit{asymptomatic recovered} $AR$ population.
 
-	Given the time scale of the pandemic, we model a weak effect of the change in the susceptible population from births or other mortality factors with the term $- \mu S$.
-	We estimate $\mu = 1.8997 \times 10^{-5}$ [per day] using the average United States death rate in urban regions from 2019 \cite{death-trends}.
+	Importantly, the transfer between population groups is defined by the Law of Mass Action.
+	We consider $I/N$ and $A/N$ to be the probability of selecting an infective or asymptomatic individual, respectively.
+	Then we define the force of infection on $S$ as $\beta_I I / N$ and $\beta_A A / N$.
+	Thus, the rate of new infections are described by $\beta_I I S / N$ and $\beta_A A S / N$.
 
@@
 		\includegraphics[height=6cm]{full-model}
+		\caption{Schematic diagram of the model. The dashed lines indicate the interaction of the infected populations with the susceptible populations that leads to infection.}
 		\label{fig:model}
-		\caption{Schematic diagram of the model. The dashed lines indicate the interaction of the infected populations with the susceptible populations that leads to infection.}
 	\end{figure}
 
+	The full model, including spatial terms, can be described by the following system of equations.
 	\begin{align*}
@@
 		E_t	&=	\D_E \Delta E + \frac{\beta_{A} A}{N} S + \frac{\beta_{I} I}{N} S - (\sigma_A + \sigma_I) E, \\
+		A_t	&=	\D_A \Delta A + \sigma_A E - M_{AR} A, \\
 		AR_t &= M_{AR} A, \\		
-		A_t	&=	\D_A \Delta A + \sigma_A E - M_{AR} A, \\
 		I_t	&=	\sigma_I E - M I, \\
@@
 	\end{align*}
+	The $- \mu S$ term models the net change in the population due to non-COVID factors.
 
+\section{ODE Model}
 	We begin by reducing the model.
-	First, we ignore non-COVID deaths, so $\mu = 0$.
-	Next we introduce the fractions
+	We introduce the fractions
 	\begin{align*}
-		s &= \frac{S}{N}, & e &= \frac{E}{N}, & ar &= \frac{AR}{N}, & a &= \frac{A}{N}, \\
+		s &= \frac{S}{N}, & e &= \frac{E}{N}, & a &= \frac{A}{N}, & ar &= \frac{AR}{N}, \\
 		i &= \frac{I}{N}, & h &= \frac{H}{N}, & r &= \frac{R}{N}, & d &= \frac{D}{N}.
 	\end{align*}
-	With these fractions and $\mu = 0$, we have the conservation law
+	Next, we ignore non-COVID deaths.
+	That is, set $\mu = 0$.
+	This gives the conservation law
 	\begin{equation} \label{eq:conservation}
@@
 	\end{equation}
-	The ODE system becomes
-	\begin{align*}
-		s_t &=	 - \beta_{A} a s - \beta_{I} i s, \\
-		e_t	&=	\beta_{A} a s + \beta_{I} i s - (\sigma_A + \sigma_I) e, \\
-		ar_t &= M_{AR} a, \\
-		a_t	&=	\sigma_A e - M_{AR} a, \\
-		i_t	&=	\sigma_I e - M i, \\
-		h_t	&=	\gamma M i - (1 - \omega) \chi h - \omega \psi h, \\
-		r_t	&=	(1 - \gamma) M i + (1 - \omega) \chi h, \\
-		d_t	&=	\omega \psi h.
-	\end{align*}
 	At this point it is worth noting the dimensions of the components of the system.
@@
 		e_\tau &= \alpha_A a s + \alpha_I i s - (\lambda_A + \lambda_I) e, \\
+		\label{eq:a_t}
+		a_\tau &= \lambda_A e - \xi a, \\
 		\label{eq:ar_t}
 		ar_\tau &= \xi a, \\
-		\label{eq:a_t}
-		a_\tau &= \lambda_A e - \xi a, \\
 		\label{eq:i_t}
 		i_\tau &= \lambda_I e - i, \\
+		\label{eq:h_t}
+		h_\tau &= \gamma i - (1 - \omega) \theta h - \omega \kappa h, \\
 		\label{eq:r_t}
@@
 	\begin{equation} \label{eq:d_t}
-		d = n - (s + e + ar + a + i + r).
+		d = n - (s + e + a + ar + i + r)
 	\end{equation}
+	so we need not solve the differential equation for $d$.
 
@@
 
+	We seed the ODE model using data from March 13, 2020 as collected on the City of Chicago Data Portal \cite{Chicago-cases}.
+	The data collected is in Cases, Hospitalized, and Deceased.
+	In order to get the initial infected population $I_0$, we take $I_0 = \mathrm{Cases} - H_0 - D_0$.
+	The relevant data is shown in Table \ref{tab:populations}.
 	\begin{table}[h]
@@
 			Total population		&	$N$		&	2,695,598 \\
-			Initial infected		&	$I_0$	&	162	\\
+			Initial infected		&	$I_0$	&	121	\\
 			Initial hospitalized	&	$H_0$	&	38 \\
@@
 			\hline
-			Transmission rate, $S \to I$														&	$\beta_{I}$	&									&	$c \in \U [0,1]$ \\
-			Transmission rate, $S \to A$														&	$\beta_{A}$	&									& 	$c \in \U [0,1]$ \\
+			Transmission rate, $S \to I$ [per day]												&	$\beta_{I}$		&									&	$c \in \U [0,1]$ \\
+			Transmission rate, $S \to A$ [per day]												&	$\beta_{A}$		&									& 	$c \in \U [0,1]$ \\
 			Lockdown effect, $S \to I$															&	$\eta_{I}$ 		&									& 	$c \in \U [0,1]$ \\
@@
 			\includegraphics[height=6cm]{cases}
+			\caption{Confirmed cases $C(t) = I(t) + R(t) + H(t) + D(t)$.}
 			\label{fig:ode-cases}
-			\caption{Confirmed cases $C(t) = I(t) + R(t) + H(t) + D(t)$.}
 		\end{subfigure}%
@@
 			\includegraphics[height=6cm]{deaths}
+			\caption{Number of deaths $D(t)$.}
 			\label{fig:ode-deaths}
-			\caption{Number of deaths $D(t)$.}
 		\end{subfigure}
@@
 
+\section{PDE Model}
+	\begin{figure}[h!]
+		\centering
+		\begin{subfigure}{0.5\textwidth}
+			\centering
+			\includegraphics[height=6cm]{domain}
+			\caption{Computational domain for the PDE model.}
+			\label{fig:domain}
+		\end{subfigure}%
+		\begin{subfigure}{0.5\textwidth}
+			\centering
+			\includegraphics[height=6cm]{mesh}
+			\caption{Preliminary mesh of the domain.}
+			\label{fig:mesh}
+		\end{subfigure}
+		\caption{Computational domain and initial meshing for the PDE model.}
+		\label{fig:pde}
+	\end{figure}
+
 \section{ODE Dynamics}
@@
 	To do this we first find the equilibrium points by solving
-	$$S_t = E_t = A_t = I_t = H_t = R_t = D_t = 0$$
-	simultaneously for $\vec{x} = (S, E, A, AR, I, H, R, D)^\intercal$.
-	The solutions of this system are of the form $\vec{x^*} = (0, 0 , 0, AR, 0, 0, R, D)^\intercal$.
+	$$s_\tau = e_\tau = ar_\tau = a_\tau = i_\tau = h_\tau = r_\tau = 0$$
+	simultaneously for $\vec{x} = (s, e, a, ar, i, h, r)^\intercal$.
+	The solutions of this system are of the form $\vec{x^*} = (0, 0 , 0, AR, 0, 0, r)^\intercal$.
 	This implies there are infinitely many non-isolated equilibrium points.
@@
 		\mat{J} = \begin{pmatrix}
-			- A \beta_{SA} - I \beta_{SI} - \mu	&	0						&	- S \beta_{SA}	&	0	&	- S \beta_{SI}	&	0									&	0	&	0	\\
-			A \beta_{SA} + I \beta_{SI}			&	- \sigma_A - \sigma_I	&	S \beta_{SA}	&	0	&	S \beta_{SI}	&	0									&	0	&	0	\\
-			0									&	\sigma_A				&	- M_{AR}		&	0	&	0				&	0									&	0	&	0	\\
-			0									&	0						&	M_{AR}			&	0	&	0				&	0									&	0	&	0	\\
-			0									&	\sigma_I				&	0				&	0	&	- M				&	0									&	0	&	0	\\
-			0									&	0						&	0				&	0	&	M \gamma		&	- \chi (1 - \omega) - \psi \omega	&	0	&	0	\\
-			0									&	0						&	0				&	0	&	M (1 - \gamma)	&	\chi (1 - \omega)					&	0	&	0	\\
-			0									&	0						&	0				&	0	&	0				&	\psi \omega							&	0	&	0
+			- a \alpha_{a} - i \alpha_{i}		&	0						&	- s \alpha_{a}	&	0	&	- s \alpha_{i}	&	0										&	0	\\
+			a \alpha_{a} + i \alpha_{i}			&	- \lambda_a - \lambda_i	&	s \alpha_{a}	&	0	&	s \alpha_{i}	&	0										&	0	\\
+			0									&	\lambda_a				&	- \xi			&	0	&	0				&	0										&	0	\\
+			0									&	0						&	\xi				&	0	&	0				&	0										&	0	\\
+			0									&	\lambda_i				&	0				&	0	&	- 1				&	0										&	0	\\
+			0									&	0						&	0				&	0	&		\gamma		&	- \theta (1 - \omega) - \kappa \omega	&	0	\\
+			0									&	0						&	0				&	0	&	1 - \gamma		&	\theta (1 - \omega)						&	0
 		\end{pmatrix}.
@@
 	\begin{equation}
-		\vec{\lambda} = \{0 , 0, 0, - M, - M_{AR}, - \mu, - \sigma_A - \sigma_I, - \chi + \chi \omega - \psi \omega \}.
+		\vec{\lambda} = \{-1 , 0, 0, 0, - \lambda_a - \lambda_i , - \xi, - \theta + \theta \omega - \kappa \omega\}.
 	\end{equation}
-	Note that the first three eigenvalues are 0, which implies the equilibrium points are non-isolated.
+	Note that three of the eigenvalues are 0, which implies the equilibrium points are non-isolated.
 	This agrees with our earlier observation.
 
-	The equilibrium points are stable when $\lambda_i < 0$ for $4 \leq i \leq 8$.
-	Since all the system parameters are positive, this implies $\lambda_i < 0$ for $4 \leq i \leq 7$.
-	Thus the stability depends on the sign of $\lambda_8$.
-	There are two cases when $\lambda_8 = - \chi + \chi \omega - \psi \omega < 0$ is true:
+	The equilibrium points are stable when $\lambda_i < 0$ for $5 \leq i \leq 7$.
+	Since all the system parameters are positive, this implies $\lambda_i < 0$ for $i = 5$ and $i = 6$.
+	Thus the stability depends on the sign of $\lambda_7$.
+	There are two cases when $\lambda_7 = - \theta + \theta \omega - \kappa \omega < 0$ is true:
 	\begin{enumerate}
 		\item $0 < \omega \leq 1$ implies $\lambda_8 < 0$, and
-		\item $\omega > 1$ and $\chi < \frac{\psi \omega}{\omega - 1}$ implies $\lambda_8 < 0$.
+		\item $\omega > 1$ and $\theta < \frac{\kappa \omega}{\omega - 1}$ implies $\lambda_8 < 0$.
 	\end{enumerate}

### 2021-11-12 version2/computemodel.m

@@ -16,7 +16,7 @@ function [T,y] = computemodel(p,q,trange,pops)
     s0 = n - (e0 + a0 + ar0 + i0 + h0 + r0 + d0);
     
     % Define the initial population vector
-    y0 = [s0, e0, a0, ar0, i0, h0, r0, d0];
+    y0 = [s0, e0, a0, ar0, i0, h0, r0];
 
     % Compute the solution vector
     [T,y] = ode45(@(t,y)sir(t,y,p1,q),trange,y0);

### 2021-11-12 version2/smoothTest.m

@@ -1,0 +1,89 @@
+function smoothTest
+
+n=100;          % number of intervals (i.e. parametric curve would be evaluted n+1 times)
+
+p =[0.5887    0.9819
+0.5887    0.9819
+0.5597    1.3551
+0.5565    1.7749
+0.4081    1.8274
+0.4500    1.4834
+0.3016    1.5300
+0.2210    1.3901
+0.2242    1.2152
+0.2790    1.0169
+0.4500    0.9295
+0.5790    0.9236
+0.5887    0.9819];
+Px = p(:,1); Py = p(:,2);
+% Note first and last points are repeated so that spline curve passes
+% through all points
+
+% when Tension=0 the class of Cardinal spline is known as Catmull-Rom spline
+Tension=0; 
+figure
+set1x=[];
+set1y=[];
+for k=1:length(Px)-3
+    [xvec,yvec]=EvaluateCardinal2DAtNplusOneValues([Px(k),Py(k)],[Px(k+1),Py(k+1)],[Px(k+2),Py(k+2)],[Px(k+3),Py(k+3)],Tension,n);
+    set1x=[set1x, xvec];
+    set1y=[set1y, yvec];    
+end
+
+Tension=0.5; 
+set2x=[];
+set2y=[];
+for k=1:length(Px)-3
+     [xvec,yvec]=EvaluateCardinal2DAtNplusOneValues([Px(k),Py(k)],[Px(k+1),Py(k+1)],[Px(k+2),Py(k+2)],[Px(k+3),Py(k+3)],Tension,n);   
+    set2x=[set2x, xvec];
+    set2y=[set2y, yvec];    
+end
+plot(set1x,set1y,'g',set2x,set2y,'b','linewidth',2)
+
+% no longer need. free the memory
+clear set1x, clear set1y, clear set2x, clear set2y, clear xvec, clear yvec 
+
+h = legend('Tension=0 (Catmull-Rom)','Tension=0.5');
+hold on
+
+plot(Px,Py,'ro','linewidth',2) % plot control points
+title('\bfCardinal Spline')
+hold off
+end
+
+function [xvec,yvec]=EvaluateCardinal2DAtNplusOneValues(P0,P1,P2,P3,T,N)
+
+xvec=[]; yvec=[];
+
+% u vareis b/w 0 and 1.
[…]

### 2021-11-12 version2/spatialModel.m

@@ -1,0 +1,45 @@
+clear
+clc
+close all;
+
+domain = shaperead('boundaries.shp','UseGeoCoords',true);
+
+pgon = simplify(polyshape(domain.Lon,domain.Lat));
+plot(pgon)
+axis equal
+
+tr = triangulation(pgon);
+
+model = createpde;
+
+tnodes = tr.Points';
+telements = tr.ConnectivityList';
+
+geometryFromMesh(model,tnodes,telements);
+
+dFig = figure(1);
+pdegplot(model)
+xlabel('$x$','interpreter','latex')
+ylabel('$y$','interpreter','latex')
+set(gca,'TickLabelInterpreter','latex')
+
+mFig = figure(2);
+pdemesh(model)
+xlabel('$x$','interpreter','latex')
+ylabel('$y$','interpreter','latex')
+set(gca,'TickLabelInterpreter','latex')
+
+% Export domain plot
+set(dFig,'Units','Inches');
+pos = get(dFig,'Position');
+set(dFig,'PaperPositionMode','Auto','PaperUnits','Inches','PaperSize',[pos(3), pos(4)])
+print(dFig,'~/Documents/GitHub/chicago-spatial-covid/update-10-22/domain','-dpdf','-r0')
+
+% Export mesh plot
+set(mFig,'Units','Inches');
+pos = get(mFig,'Position');
+set(mFig,'PaperPositionMode','Auto','PaperUnits','Inches','PaperSize',[pos(3), pos(4)])
+print(mFig,'~/Documents/GitHub/chicago-spatial-covid/update-10-22/mesh','-dpdf','-r0')
+
+
+%generateMesh(model)

### 2021-11-18 version3/version3.tex
version3 - new simplified model

\documentclass[11pt]{article}

%imports
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{footnote}
\usepackage{url}
\usepackage{hyperref}
\usepackage{subcaption}
\usepackage{tikz}

\usetikzlibrary{fit,shapes.geometric}

%images and image path
\usepackage{graphicx}
%\usepackage{subfig}
\graphicspath{ {./images/} }

%custom commands
\newcommand{\R}{\mathbf{R}}
\newcommand{\N}{\mathcal{N}}
\newcommand{\num}{\text{num}}
\newcommand{\obs}{\text{obs}}
\newcommand{\D}{\mathfrak{D}}
\newcommand{\Ro}{\mathcal{R}_0}
\newcommand{\lap}{{\mathcal{L}}}
\renewcommand\vec{\mathbf}
\newcommand{\mat}[1]{\mathsf{#1}}
\newcommand{\U}{\mathcal{U}}

%insitute command
\usepackage{etoolbox}
\makeatletter
\providecommand{\institute}[1]{% add institute to \maketitle
	\apptocmd{\@author}{\end{tabular}
	\par
	\begin{tabular}[t]{c}
		\small \textit{#1}}{}{}
}
\makeatother

%margins
\usepackage[letterpaper, total={6.5in, 9.5in}]{geometry}

%opening
\title{Reaction-diffusion spatial modeling of COVID-19 in Chicago}
\author{Trent Gerew\thanks{\texttt{tgerew@hawk.iit.edu}}}
\institute{Department of Applied Mathematics, Illinois Institute of Technology, Chicago, Illinois}

\begin{document}

\maketitle

\begin{abstract}
	content...
\end{abstract}

\section{Introduction}

\section{Materials and methods}
	\subsection{Confirmed and death data}
		In this study, we used the publicly available data set of COVID-19 provided by the City of Chicago Data Portal.
		This data set includes daily counts of confirmed infected cases, hospitalizations, and deaths.

	\subsection{Mathematical model}
		We focus our study on four components of the epidemic flow (Figure \ref{fig:model}).
		That is, the populations of Susceptible individuals ($S$), Asymptomatic infected individuals ($A$), symptomatic Infected individuals ($I$), and Removed individuals ($R$).
		\begin{figure}[h]
			\centering
			\begin{tikzpicture}
				\node[shape=rectangle, draw=black, fill=green] (s) at (-3,0) {$S$};
				\node[shape=rectangle, draw=black, fill=yellow] (a) at (0,1) {$A$};
				\node[shape=rectangle, draw=black, fill=orange] (i) at (0,-1) {$I$};
				\node[shape=rectangle, draw=black, fill=cyan] (r) at (3,0) {$R$};
				\path[thick,->] (s) edge node[above,swap] {$\omega \frac{(\beta_A A + \beta_I I)}{N}$} (a);
				\path[thick,->] (a) edge node[right,swap] {$\delta$} (i);
				\path[thick,->] (a) edge node[above,swap] {$\gamma_A$}	(r);
				\path[thick,->] (i) edge node[below,swap] {$\gamma_I$}	(r);
			\end{tikzpicture}
			\caption{Compartmental representation of the SAIR model.}
			\label{fig:model}
		\end{figure}
		Our model is known as the SAIR model \cite{s+t+spain}, which substitutes the $E$ compartment of the SEIR model by the $A$ compartment.
		This model is relevant when there are many undetected asymptomatic infectious individuals, which is known to be the case for COVID-19.

		We motivate the use of this simple model for the following reasons:
		\begin{itemize}
			\item
				We do not consider the $E$ group in this work because we feel this population can be merged with the $A$ group.
			\item
				We do not distinguish between quarantined, hospitalized, or nursing-home populations.
				Including these additional compartments would over parameterize the model.
				Additionally, the way in which data has been collected and is provided by authorities has varied over time, making its usage rather difficult.
				For the same reasons, we do not distinguish between deceased and recovered.
		\end{itemize}

		To build the mathematical model, we followed the standard strategy developed in the literature concerning SIR models.
		We assume that individuals in $S$ can be infected by both members of $A$ and $I$.
		We suppose that the individuals in the $A$ and $I$ compartments may have different contact rates $\beta_A$ and $\beta_I$, and different recovery rates $\gamma_A$ and $\gamma_I$.
		Furthermore, we consider a rate $\delta$ at which individuals in $A$ may develop symptoms or are otherwise detected and so will move to the $I$ compartment.
		Lastly, we assume that only members of $S$ and $A$ are mobile.

		The dynamics is governed by a system of two partial differential equations (PDE) and two ordinary differential equations (ODE) as follows, for $\vec{x} = (x,y) \in \Omega \subset \R^2, t > 0$,
		\begin{equation} \label{eq:model}
			\begin{aligned}
				\partial_t S - D(t) \Delta S &= - \omega(t) \frac{\beta_A A + \beta_I I}{N} S, \\
				\partial_t A - D(t) \Delta A &= \omega(t) \frac{\beta_A A + \beta_I I}{N} S - (\gamma_A - \delta) A, \\
				I' &= - \gamma_I I + \delta A, \\
				R' &= \gamma_A A + \gamma_I.
			\end{aligned}
		\end{equation}
		Since travel into the city of Chicago was heavily restricted for the early stages of the pandemic, the homogeneous Neumann boundary condition is imposed \cite{Mammeri+2020+102+113}.
		The total population is $N = S + A + I + R$, and is considered a constant.

	\subsection{Parameter estimation}
		To account for the lockdown, the average number of of contacts is updated as follows \cite{Kevrekidis-2021}
		\begin{equation} \label{eq:contacts}
			\omega (t) = \omega_0 \left[ \eta + (1 - \eta) \frac{1 - \tanh \left[2 (t - t_q) \right]}{2} \right],
		\end{equation}
		while the diffusion coefficient is similarly
		\begin{equation} \label{eq:diffusion}
			D (t) = D_0 \left[ \eta + (1 - \eta) \frac{1 - \tanh \left[2 (t - t_q) \right]}{2} \right].
		\end{equation}
		Here, $t_q = (t_{\mathrm{eol}} + t_{\mathrm{bol}}) / 2$, where $\mathrm{bol}$ denotes the beginning of the lockdown and $\mathrm{eol}$ denotes the end of the lockdown.
		The parameter $0 \leq \eta \leq 1$ is a varying coefficient translating respect for social distancing and other preventative measures.
		Note that $\omega_0$ is the average number of contacts before any intervention, and is a constant.

	\subsection{Numerical discretization}
\section{Results}
	\subsection{Existence of solutions and basic reproduction numbers}

	\subsection{Model resolution}

	\subsection{Spatial spread of COVID-19}

\section{Discussion}

	\textbf{Acknowledgments:} \\
	\textbf{Funding:}
	This research was funded by the Socially Responsible Modeling, Computation, and Design (SoReMo) Initiative, through the Illinois Institute of Technology. \\
	\textbf{Conflicts of Interest:}
	Author states no conflict of interest. \\

\bibliographystyle{abbrv}
\bibliography{version3}

\end{document}

### 2021-11-18 version3/version3.tex
added classical literature

@@
 
-		To build the mathematical model, we followed the standard strategy developed in the literature concerning SIR models.
+		To build the mathematical model, we followed the standard strategy developed in the literature concerning SIR models \cite{bio-models}.
 		We assume that individuals in $S$ can be infected by both members of $A$ and $I$.
@@
 	\textbf{Funding:}
-	This research was funded by the Socially Responsible Modeling, Computation, and Design (SoReMo) Initiative, through the Illinois Institute of Technology. \\
+	This research was funded by the Socially Responsible Modeling, Computation, and Design (SoReMo) Initiative, at the Illinois Institute of Technology. \\
 	\textbf{Conflicts of Interest:}

### 2021-11-18 version3/version3.tex
updates

@@
 \usepackage{tikz}
+\usepackage{amsthm}
 
@@
 \newcommand{\R}{\mathbf{R}}
+\newcommand{\C}{\mathbf{C}}
 \newcommand{\N}{\mathcal{N}}
@@
 
-%insitute command
+\newtheorem{theorem}{Theorem}
+\newtheorem{conjecture}{Conjecture}
+\newtheorem{definition}{Definition}
+
+%intsitute command
 \usepackage{etoolbox}
@@
 	\subsection{Numerical discretization}
+
 \section{Results}
 	\subsection{Existence of solutions and basic reproduction numbers}
+		We define $\vec{x} = (S,  A, I, R)$ and $\vec{x}_0 = (S_0, A_0, I_0, R_0)$.
+		\begin{theorem}
+			Let $0 \leq \vec{x}_0 \leq N$ be the initial datum.
+			Then there exists a unique in time solution $\vec{x}$ of the initial value problem (\ref{eq:model}) without diffusion over $C \subset U \subset \R^4 \times \R^1$ where $C$ is a compact set that contains $(\vec{x}_0, t_0)$.
+			Moreover, the solution is $\C^1$.
+		\end{theorem}
+		\begin{proof}
+			Since $\vec{x}_t = f(\vec{x},t)$ is $\C^1$ on the open set $U \subset \R^4 \times \R^1$, it follows that there exists a solution of (\ref{eq:model}) without diffusion through the point $\vec{x}_0$ at $t = t_0$ for $|t - t_0|$ sufficiently small.
+			Moreover, $\vec{x} (t, t_0,\vec{x_0})$ is a $\C^1$ function \cite{dynamics}.
+			Since $C$ is a compact set containing $(\vec{x}_0, t_0)$, then the solution $\vec{x} (t, t_0, \vec{x}_0)$ can be uniquely extended backward and forward in $t$ up to the boundary of $C$ \cite{dynamics}.
+		\end{proof}
+
+		\begin{conjecture}
+			Let $0 \leq \vec{x}_0 \leq N$ the initial datum.
+			Then there exists a unique global in time solution $\vec{x}$ of the initial boundary value problem (\ref{eq:model}).
+		\end{conjecture}
+
+		An important parameter in understanding the initial growth of the disease is the basic reproduction number.
+		\begin{definition}[Basic Reproduction Number]
+			The basic reproduction number $\Ro$ can be computed using the next generation of the model without diffusion.
+			Since the infected individuals are in $A$ and $I$, new infections ($\mathcal{F}$) and transitions between compartments ($\mathcal{V}$) can be rewritten as
+			\begin{align*}
+				\mathcal{F} &= \begin{pmatrix} \omega (\beta_A A + \beta_I I) \frac{S}{N} \\ 0 \end{pmatrix}, &
+				\mathcal{V} &= \begin{pmatrix} (\gamma_A + \delta) A \\ \gamma_I - \delta A \end{pmatrix}.
+			\end{align*}
+			Thus, $\Ro = \rho (\mat{F} \mat{V}^{-1})$ of the next generation matrix
+			\begin{equation*}
+				\mat{F} \mat{V}^{-1} = \begin{pmatrix} 
+				\frac{S_0 \omega_0 \beta_A}{N (\gamma_A + \delta)} + \frac{S_0 \omega_0 \beta_I \delta}{N \gamma_I (\gamma_A + \delta)} & 
+				\frac{S_0 \omega_0 \beta_I}{N \gamma_I} \\ 
+				0 & 0 \end{pmatrix}.
+			\end{equation*}
+			So,
+			\begin{equation} \label{eq:ro}
+				\Ro =  \frac{\omega_0}{\gamma_A + \delta}  \left( \beta_A + \beta_I \frac{\delta}{\gamma_I} \right) \frac{S_0}{N}.
+			\end{equation}
+		\end{definition}
+		This number is dimensionless and has an epidemiological meaning.
+		The first term represents the transmission rate by asymptomatic individuals, and the second term represents the transmission rate by symptomatic individuals.
 
@@
 
+\section*{Appendix}
+	\subsection*{Nomenclature (Units are indicated in brackets.)}
+		\begin{table}[h!]
+			\begin{tabular}{c c}
+				\textbf{Latin symbols} & \\
+				$A$ & Asymptomatic infected individuals [individuals] \\
+				$I$ & Symptomatic infected individuals [individuals] \\
+				$N$ & Total population of individuals [individuals] \\
+				$R$ & Removed individuals [individuals] \\
+				$S$ & Susceptible individuals [individuals] \\
+				\textbf{Greek symbols} & \\
+				$\beta_j$ & Contact rate of compartment $j$ [days\textsuperscript{-1}] \\
+				$\gamma_j$ & Recovery rate of compartment $j$ [days\textsuperscript{-1}] \\
+				$\delta$ & Rate of asymptomatic individuals that may develop symptoms [days\textsuperscript{-1}]
+			\end{tabular}
+		\end{table}
+
 \bibliographystyle{abbrv}

### 2021-11-19 version3/version3.tex
Added Sara's paper as a reference

@@
 \usepackage{amsthm}
+\usepackage{longtable}
 
@@
 	\subsection{Confirmed and death data}
-		In this study, we used the publicly available data set of COVID-19 provided by the City of Chicago Data Portal.
+		In this study, we used the publicly available data set of COVID-19 provided by the City of Chicago Data Portal \cite{Chicago-cases}.
 		This data set includes daily counts of confirmed infected cases, hospitalizations, and deaths.
@@
 				Including these additional compartments would over parameterize the model.
-				Additionally, the way in which data has been collected and is provided by authorities has varied over time, making its usage rather difficult.
+				Additionally, the way in which data has been collected and is provided by authorities has varied over time, making its usage rather difficult \cite{sara-data}.
 				For the same reasons, we do not distinguish between deceased and recovered.
@@
 	\subsection{Existence of solutions and basic reproduction numbers}
-		We define $\vec{x} = (S,  A, I, R)$ and $\vec{x}_0 = (S_0, A_0, I_0, R_0)$.
+		We show that we are justified in searching for suitable parameters to solve the model.
+		Let $\vec{x} = (S,  A, I, R)$ and $\vec{x}_0 = (S_0, A_0, I_0, R_0)$.
+		First we show that the solution of the initial value problem (\ref{eq:model}) exists for the case with no diffusion.
 		\begin{theorem}
@@
 		\end{proof}
-
+		\noindent We do provide a proof for the existence of solutions to the initial boundary value problem (\ref{eq:model}) with diffusion terms, but we assume it exists.
 		\begin{conjecture}
@@
 		\end{conjecture}
+		\noindent A proof for this conjecture will be similar to the one given in \cite{Mammeri+2020+102+113}.
 
-		An important parameter in understanding the initial growth of the disease is the basic reproduction number.
+		An important parameter in understanding the initial growth of the virus is the basic reproduction number.
 		\begin{definition}[Basic Reproduction Number]
@@
 		\end{definition}
-		This number is dimensionless and has an epidemiological meaning.
+		\noindent This number is dimensionless and has an epidemiological meaning.
 		The first term represents the transmission rate by asymptomatic individuals, and the second term represents the transmission rate by symptomatic individuals.
@@
 	\textbf{Conflicts of Interest:}
-	Author states no conflict of interest. \\
 
@@
 	\subsection*{Nomenclature (Units are indicated in brackets.)}
-		\begin{table}[h!]
-			\begin{tabular}{c c}
-				\textbf{Latin symbols} & \\
-				$A$ & Asymptomatic infected individuals [individuals] \\
-				$I$ & Symptomatic infected individuals [individuals] \\
-				$N$ & Total population of individuals [individuals] \\
-				$R$ & Removed individuals [individuals] \\
-				$S$ & Susceptible individuals [individuals] \\
-				\textbf{Greek symbols} & \\
-				$\beta_j$ & Contact rate of compartment $j$ [days\textsuperscript{-1}] \\
-				$\gamma_j$ & Recovery rate of compartment $j$ [days\textsuperscript{-1}] \\
-				$\delta$ & Rate of asymptomatic individuals that may develop symptoms [days\textsuperscript{-1}]
-			\end{tabular}
-		\end{table}
-
+		\begin{longtable}{c l l}
+			\textbf{Latin symbols} & & \\
+			$A$ & Asymptomatic infected individuals & [individuals] \\
+			$D$ & Diffusivity & \\
+			$I$ & Symptomatic infected individuals & [individuals] \\
+			$N$ & Total population of individuals & [individuals] \\
+			$R$ & Removed individuals & [individuals] \\
+			$S$ & Susceptible individuals & [individuals] \\
+			\textbf{Greek symbols} & & \\
+			$\beta_j$ & Contact rate of compartment $j$ & [days\textsuperscript{-1}] \\
+			$\gamma_j$ & Recovery rate of compartment $j$ & [days\textsuperscript{-1}] \\
+			$\delta$ & Rate of asymptomatic individuals that may develop symptoms & [days\textsuperscript{-1}] \\
+			$\omega$ & Contacts & 
+		\end{longtable}
 \bibliographystyle{abbrv}

### 2021-12-17 version3/version3.tex
Writing report

@@
 				\node[shape=rectangle, draw=black, fill=cyan] (r) at (3,0) {$R$};
-				\path[thick,->] (s) edge node[above,swap] {$\omega \frac{(\beta_A A + \beta_I I)}{N}$} (a);
+				\path[thick,->] (s) edge node[above,swap] {$\omega (\beta_A A + \beta_I I)$} (a);
 				\path[thick,->] (a) edge node[right,swap] {$\delta$} (i);
@@
 			\begin{aligned}
-				\partial_t S - D(t) \Delta S &= - \omega(t) \frac{\beta_A A + \beta_I I}{N} S, \\
-				\partial_t A - D(t) \Delta A &= \omega(t) \frac{\beta_A A + \beta_I I}{N} S - (\gamma_A - \delta) A, \\
-				I' &= - \gamma_I I + \delta A, \\
-				R' &= \gamma_A A + \gamma_I.
+				S_t - D(t) \Delta S &= - \omega(t) \frac{\beta_A A + \beta_I I}{N} S, \\
+				A_t - D(t) \Delta A &= \omega(t) \frac{\beta_A A + \beta_I I}{N} S - (\gamma_A - \delta) A, \\
+				I_t &= - \gamma_I I + \delta A, \\
+				R_t &= \gamma_A A + \gamma_I.
 			\end{aligned}
@@
 		Since travel into the city of Chicago was heavily restricted for the early stages of the pandemic, the homogeneous Neumann boundary condition is imposed \cite{Mammeri+2020+102+113}.
-		The total population is $N = S + A + I + R$, and is considered a constant.
+		The population compartments are considered fractions, such that $S + A + I + R = 1$.
 
 	\subsection{Parameter estimation}
-		To account for the lockdown, the average number of of contacts is updated as follows \cite{Kevrekidis-2021}
+		To account for the lockdown, the average number of of contacts is updated as in  \cite{Kevrekidis-2021}
 		\begin{equation} \label{eq:contacts}
-			\omega (t) = \omega_0 \left[ \eta + (1 - \eta) \frac{1 - \tanh \left[2 (t - t_q) \right]}{2} \right],
+			\omega (t) = \omega_0 \left[ \eta + (1 - \eta) \frac{1 - \tanh \left[2 (t - t_q) \right]}{2} \right].
 		\end{equation}
-		while the diffusion coefficient is similarly
-		\begin{equation} \label{eq:diffusion}
-			D (t) = D_0 \left[ \eta + (1 - \eta) \frac{1 - \tanh \left[2 (t - t_q) \right]}{2} \right].
-		\end{equation}
-		Here, $t_q = (t_{\mathrm{eol}} + t_{\mathrm{bol}}) / 2$, where $\mathrm{bol}$ denotes the beginning of the lockdown and $\mathrm{eol}$ denotes the end of the lockdown.
+		Here, $t_q = (t_{\mathrm{eol}} + t_{\mathrm{bol}}) / 5$, where $\mathrm{bol}$ denotes the beginning of the lockdown and $\mathrm{eol}$ denotes the end of the lockdown.
 		The parameter $0 \leq \eta \leq 1$ is a varying coefficient translating respect for social distancing and other preventative measures.
@@
 
+		The parameters $\omega_0$, $\beta_A$, and $\beta_I$ are not independently identifiable, so the optimization problem reduces to five parameters $\theta = (\omega_0 \beta_A, \omega_0 \beta_I, \delta, \gamma_A, \gamma_I)$.
+		Given, for $n$ days, the observations $I_\mathrm{obs} (t_i)$, the objective function is the least square function
+		\begin{equation} \label{eq:obj}
+			J (\theta) = \sum_{i=1}^n \left[ I_\mathrm{obs} (t_i) - I (t_i) \right]^2
+		\end{equation}
+		with constraints $0 \leq \theta \leq 1$.
+		$I (t_i)$ denotes the output of the mathematical model at time $t_i$ computed with parameters $\theta$.
+
+		The optimization problem is solved using the MATLAB nonlinear optimization function \verb|fmincon|.
+		The initial parameter guesses to seed \verb|fmincon| were randomly sampled from a uniform distribution over 1000 iterations.
+		The median of each resulting parameter was selected.
+		Figure \ref{fig:params} shows the estimated parameters.
+
 	\subsection{Numerical discretization}
 
+		while the diffusion coefficient is
+		\begin{equation} \label{eq:diffusion}
+		D (t) = D_0  - \eta H(t - t_q)
+		\end{equation}
+		where $H(t)$ is the Heaviside step function.
 \section{Results}
@@
 			\begin{align*}
-				\mathcal{F} &= \begin{pmatrix} \omega (\beta_A A + \beta_I I) \frac{S}{N} \\ 0 \end{pmatrix}, &
+				\mathcal{F} &= \begin{pmatrix} \omega (\beta_A A + \beta_I I) S \\ 0 \end{pmatrix}, &
 				\mathcal{V} &= \begin{pmatrix} (\gamma_A + \delta) A \\ \gamma_I - \delta A \end{pmatrix}.
@@
 				\mat{F} \mat{V}^{-1} = \begin{pmatrix} 
-				\frac{S_0 \omega_0 \beta_A}{N (\gamma_A + \delta)} + \frac{S_0 \omega_0 \beta_I \delta}{N \gamma_I (\gamma_A + \delta)} & 
-				\frac{S_0 \omega_0 \beta_I}{N \gamma_I} \\ 
+				\frac{S_0 \omega_0 \beta_A}{\gamma_A + \delta} + \frac{S_0 \omega_0 \beta_I \delta}{\gamma_I (\gamma_A + \delta)} & 
+				\frac{S_0 \omega_0 \beta_I}{\gamma_I} \\ 
 				0 & 0 \end{pmatrix}.
@@
 			\begin{equation} \label{eq:ro}
-				\Ro =  \frac{\omega_0}{\gamma_A + \delta}  \left( \beta_A + \beta_I \frac{\delta}{\gamma_I} \right) \frac{S_0}{N}.
+				\Ro =  \frac{S_0 \omega_0}{\gamma_A + \delta}  \left( \beta_A + \beta_I \frac{\delta}{\gamma_I} \right).
 			\end{equation}
@@
 	\subsection{Model resolution}
+		Calibration of the model is done from March 18, 2020 to June 24, 2020.
+		This range corresponds approximately to the first ``wave'' of cases seen in Chicago.
+		The lockdown time points match exactly to the imposed lockdown of the city.
+		That is, $t_\mathrm{bol} =$ March 21, 2020 and $t_\mathrm{eol} =$ May 29, 2020.
+		\begin{figure}[h!]
+			\centering
+			\begin{subfigure}{0.55\textwidth}
+				\begin{tabular}{c c c c}
+					\hline
+					$\omega_0 \beta_A$ & transmission rate due to $A$ & 0.1969 & days\textsuperscript{-1} \\
+					$\omega_0 \beta_I$ & transmission rate due to $I$ & 0.3061 & days\textsuperscript{-1} \\
+					$\eta$ & lockdown scale factor & 0.5514 & 1 \\
+					$\delta$ & symptom onset rate & 0.0939 & days\textsuperscript{-1} \\
+					$\gamma_A$ & removal rate of $A$ & 0.3632 & days\textsuperscript{-1} \\
+					$\gamma_I$ & removal rate of $I$ & 0.1385 & days\textsuperscript{-1} \\
+					\hline
+				\end{tabular}
+			\end{subfigure}%
+			\begin{subfigure}{0.45\textwidth}
+				\includegraphics[width=\textwidth]{box-plot.pdf}
+			\end{subfigure}
+			\caption{Parameters calibrated according to data from the Chicago Data Portal.
+				On the right is a boxplot of the parameter distribution from 1000 optimization iterations.}
+			\label{fig:params}
+		\end{figure}
+
+		\begin{figure}[h!]
+			\centering
+			\includegraphics[width=0.5\textwidth]{cases.pdf}
+			\caption{Fitting symptomatic infected by the median value.
+				Note the logarithmic scale on the y-axis.}
+			\label{fig:fit}
+		\end{figure}
 

### 2021-12-17 data/final_zip_data.py

@@ -6,11 +6,13 @@ df = pd.read_csv("COVID-19_Cases__Tests__and_Deaths_by_ZIP_Code.csv")
 cols = ["ZIP Code", "Week Start", "Cases - Weekly", "Deaths - Cumulative", "Population"]
 df = df[cols]
 
-df = df[df["Week Start"] == "04/19/2020"]
+#df = df[df["Week Start"] == "04/19/2020"]
+#df = df[df["Week Start"] == "03/29/2020"]
+df = df[df["Week Start"] == "06/21/2020"]
 df = df[df["ZIP Code"] != "Unknown"]
 df["Cases - Weekly"] = df["Cases - Weekly"].replace(np.nan, 0)
 
 cols = ["ZIP Code", "Cases - Weekly", "Deaths - Cumulative", "Population"]
 df = df[cols]
 
-df.to_csv("final_data.csv", index=False, header=False)
+df.to_csv("tend_data.csv", index=False, header=False)

### 2021-12-17 version3/box_plot.py

@@ -1,0 +1,23 @@
+# Imports
+import pandas as pd
+import matplotlib.pyplot as plt
+
+# LaTeX font for matplotlib
+import matplotlib
+matplotlib.rcParams['mathtext.fontset'] = 'stix'
+matplotlib.rcParams['font.family'] = 'STIXGeneral'
+
+# Read in csv and label the columns
+df = pd.read_csv("pEst1.csv")
+columns = [r'$\omega_0 \beta_A$', r'$\omega_0 \beta_I$', r'$\eta$', r'$\gamma_A$', r'$\gamma_I$', r'$\delta$']
+df.columns = columns
+
+# Create a boxplot
+boxplot = df.boxplot(column=columns)
+
+# Export plot
+plt.savefig('box-plot.pdf')
+
+# Plot the boxplot
+boxplot.plot()
+plt.show()

### 2021-12-17 version3/initialZips.m

@@ -7,7 +7,16 @@ h = height(zips);
 faceColors = makesymbolspec('Polygon',...
     {'INDEX',[1 h],'FaceColor',polcmap(h)});
 
+cFig = figure(1)
+
 mapshow(zips,'SymbolSpec',faceColors)
+
+set(cFig,'Units','Inches');
+pos = get(cFig,'Position');
+set(cFig,'PaperPositionMode','Auto','PaperUnits','Inches','PaperSize',[pos(3), pos(4)])
+print(cFig,'~/Documents/GitHub/chicago-spatial-covid/version3/zips','-dpng','-r0')
+
+
 
 initdata = load("initial_data.csv");
 

### 2021-12-17 version3/no-intervention.py

@@ -24,11 +24,6 @@ gammaA = 0.3632 # Removal rate of A
 gammaI = 0.1385 # Removal rate of I
 l = 0.0939 # Symptom onset rate
 
-# Time dependent beta function
-def omegabeta(omega0beta,eta,t,tq):
-    result = omega0beta * (eta + (1 - eta) * (1 - np.tanh(2 * (t - tq))) / 2)
-    return result
-
 
 ## PDE Model
 
@@ -44,18 +39,6 @@ model = Model(["D * (dxxs + dyys) - (omegaA * a + omegaI * i) * s",
 L = 40 # Space points
 x = y = np.linspace(0, L, L) # Domain space
 
-# Hook function for time dependent parameters
-def hook(t, fields):
-    global omegaA, omegaI, eta, tq
-    # Define parameters at current time
-    omegaA = omegabeta(omegaA,eta,t,tq)
-    omegaI = omegabeta(omegaI,eta,t,tq)
-
-    # Reassign fields
-    fields["omegaA"] = omegaA
-    fields["omegaI"] = omegaI
-
-    return fields
 
 ## Simulation 1: t=0 -> t=tq
 
@@ -64,45 +47,27 @@ s1 = np.genfromtxt('initialS.csv', dtype=float, delimiter=",")
 a1 = np.genfromtxt('initialA.csv', dtype=float, delimiter=",")
 i1 = np.genfromtxt('initialI.csv', dtype=float, delimiter=",")
 
-# Approximate the time dependent parameters
-omegaA1 = omegabeta(omegaA,eta,0,tq)
-omegaI1 = omegabeta(omegaA,eta,0,tq)
-
 # Populate the model fields
 initial_fields1 = model.Fields(x=x, y=y, s=s1, a=a1, i=i1, D=D, omegaA=omegaA, omegaI=omegaI, gammaA=gammaA, gammaI=gammaI, l=l)
 
 # Create and run the simulation
-simulation1 = Simulation(model, initial_fields1, dt=.1, tmax=tq, hook=hook)
+simulation1 = Simulation(model, initial_fields1, dt=.1, tmax=tq)
 container1 = simulation1.attach_container()
 tmax1, final_fields1 = simulation1.run()
 
 # Export results
-initial_fields1["a"].plot()
-plt.gca().invert_yaxis()
[…]

### 2021-12-17 version3/objfun.m

@@ -17,6 +17,7 @@ function obj = objfun(trange,Cobs,p,tq,y0,n)
     Cnum = i;
     
     % Compute the objective function
-    obj = sum(vecnorm(Cnum - Cobs / n));
-    disp(obj)
+    %obj = sum(vecnorm(Cnum - Cobs' / n));
+    obj = sum((Cnum - Cobs' / n).^2);
+    %disp(obj)
 end

### 2021-12-17 version3/optim.m

@@ -19,7 +19,8 @@ load('params.mat');
 lb = zeros(1,6);
 
 % Upper bounds
-ub = 1 * ones(1,6);
+%ub = 1 * ones(1,6);
+ub = [0.6, 0.6, 0.8, 0.7, 0.3, 0.3];
 
 %% Load Observed Data
 [time, Cobs, Dobs] = loadData("time_series.csv");
@@ -30,7 +31,7 @@ t_f = length(time);
 %% Optimum Solution
 
 % Define number of iterations
-iters = 1;
+iters = 1000;
     
 % Define initial populations
 y0 = [s0, a0, i0, r0];
@@ -43,22 +44,16 @@ tic % Start timer
 for iter=1:iters
 
     % Sample starting parameters
-    %p0 = sampleparams(lb,ub,s0);
-    p0 = (lb + ub) / 2;
+    p0 = sampleparams(lb,ub,s0);
+    %p0 = (lb + ub) / 2;
     
     % Compute parameter estimate
-    pEst(iter,:) = fmincon(@(p)objfun([1,t_f],Cobs,p,tq,y0,n),p0,[],[],[],[],lb,ub);
+    options = struct('MaxFunctionEvaluations', 1000);
+    pEst(iter,:) = fmincon(@(p)objfun(time,Cobs,p,tq,y0,n),p0,[],[],[],[],lb,ub);
 
     disp("Iteration: " + num2str(iter))
+
+    % Export results
+    writematrix(pEst(iter,:),'pEst8.csv','WriteMode','append')
 end
 toc % End timer
-
-%% Export Results
-% Compute the final solution vector from the median of the estimates
-pFin = median(pEst,1);
-
-% Compute the interquartile range
-iqrange = iqr(pEst,1);
-
-% Save the results to a file
-save('optimfile.mat','pFin','iqrange')

### 2021-12-17 version3/params.m

@@ -18,14 +18,14 @@ d0 = 3; % Initial deceased
 
 % Initial compartments
 i0 = (c0 + h0) / n;
-a0 = i0;
+a0 = 3 * i0;
 r0 = (d0 + 8) / n;
 s0 = 1 - (a0 + i0 + r0);
 
 % Lockdown times
 tbol = 4; % beginning of lockdown
 teol = 73; % end of lockdown
-tq = (tbol + teol) / 3;
+tq = (tbol + teol) / 5;
 
 % Export fixed parameters
 save('params.mat')

### 2021-12-17 version3/results.m

@@ -15,7 +15,11 @@ load('params.mat');
 y0 = [s0, a0, i0, r0];
 
 %% Load Optimized parameters
-load('optimfile.mat', 'pFin');
+pEst = load('pEst1.csv');
+pFin = median(pEst,1);
+
+% Compute the interquartile range
+iqrange = iqr(pEst,1);
 
 %% Load Observed Data
 [time, Cobs, Dobs] = loadData("time_series.csv");
@@ -41,13 +45,19 @@ Cnum = i;
 % Total cases
 cFig = figure(1);
 semilogy(T,Cnum,'r-',time,Cobs / n,'k.')
-legend('$C_\mathrm{num}(t)$','$C_\mathrm{obs}(t)$','Interpreter','latex','Location','southeast')
+legend('$I (t)$','$I_\mathrm{obs}(t)$','Interpreter','latex','Location','southeast')
 xlabel('$t$ (days from March 18)','interpreter','latex')
-ylabel('$C(t)$','interpreter','latex')
+ylabel('$I(t)$ (symptomatic infected)','interpreter','latex')
 set(gca,'TickLabelInterpreter','latex')
+
+cFig2 = figure(2);
+semilogy(T,Cnum,'r-',time,Cobs / n,'k.')
+legend('C_{num}','C_{obs}','Location','southeast')
+xlabel('t (days from March 18)')
+ylabel('C(t) (confirmed cases)')
 
 % Export cases plot
 set(cFig,'Units','Inches');
 pos = get(cFig,'Position');
 set(cFig,'PaperPositionMode','Auto','PaperUnits','Inches','PaperSize',[pos(3), pos(4)])
-print(cFig,'~/Documents/GitHub/chicago-spatial-covid/version3/cases','-dpng','-r0')
+print(cFig,'~/Documents/GitHub/chicago-spatial-covid/version3/cases','-dpdf','-r0')

### 2021-12-17 version3/ziptest.m

@@ -7,7 +7,7 @@ h = height(zips);
 faceColors = makesymbolspec('Polygon',...
     {'INDEX',[1 h],'FaceColor',polcmap(h)});
 
-findata = load("final_data.csv");
+findata = load("tend_data.csv");
 
 for i=1:length(zips)
     zip = str2num(zips(i).zip);
@@ -27,7 +27,7 @@ for i=1:length(zips)
 end
 
 maxinfected = max([zips.objectid]);
-fall = flipud(autumn(height(zips)));
+fall = parula(height(zips));
 densityColors = makesymbolspec('Polygon', {'objectid', [0 maxinfected], 'FaceColor', fall});
 
 cFig = figure(1)
@@ -40,4 +40,4 @@ colorbar
 set(cFig,'Units','Inches');
 pos = get(cFig,'Position');
 set(cFig,'PaperPositionMode','Auto','PaperUnits','Inches','PaperSize',[pos(3), pos(4)])
-print(cFig,'~/Documents/GitHub/chicago-spatial-covid/version3/final-cases','-dpng','-r0')
+print(cFig,'~/Documents/GitHub/chicago-spatial-covid/version3/tend-cases','-dpng','-r0')

### 2021-12-17 spatial-model/ZipGeometry.m
Clean repository

@@ -1,37 +1,0 @@
-clear
-clc
-close all;
-
-zips = shaperead('chicago_zips.shp','UseGeoCoords',true);
-h = height(zips);
-faceColors = makesymbolspec('Polygon',...
-    {'INDEX',[1 h],'FaceColor',polcmap(h)});
-
-geoshow(zips,'SymbolSpec',faceColors)
-
-% Create a polygon for each ZIP code
-for zip = 1:length(zips)
-    poly(zip) = simplify(polyshape(zips(zip).Lon, zips(zip).Lat));
-end
-
-
-% % Take the union of the polygons to create the entire domain
-% polyout = simplify(union(poly));
-% 
-% % Create a triangulation representation of the domain
-% tr = triangulation(polyout);
-% 
-% % Create a PDE model
-% model = createpde;
-% 
-% tnodes = tr.Points';
-% telements = tr.ConnectivityList';
-% 
-% geometryFromMesh(model,tnodes,telements);
-% pdegplot(model)
-% 
-% generateMesh(model,'Hgrad',1.3)
-% 
-% figure
-% pdemesh(model)
-% pdegplot(model)

### 2021-12-17 spatial-model/mapTest.m

@@ -1,19 +1,0 @@
-clear
-clc
-close all;
-
-zips = shaperead('chicago_zips.shp','UseGeoCoords',false);
-h = height(zips);
-faceColors = makesymbolspec('Polygon',...
-    {'INDEX',[1 h],'FaceColor',polcmap(h)});
-
-mapshow(zips,'SymbolSpec',faceColors)
-grid on
-
-series = load("initial_data.csv");
-
-for i = 1:61
-    disp(zips(i).BoundingBox);
-    disp(zips(i).zip);
-end
-

### 2021-12-17 spatial-model/maptest2.m

@@ -1,19 +1,0 @@
-MapLatLimit = [41 48];
-MapLonLimit = [-74 -66];
-
-states = readgeotable('usastatelo.shp');
-n = ["Maine" "New Hampshire" "Vermont" "Massachusetts" "Connecticut" "Rhode Island" "New York"];
-rows = ismember(states.Name,n);
-NEstates = states(rows,:);
-
-maxdensity = max([NEstates.PopDens2000])
-fall = flipud(autumn(height(NEstates)));
-
-densityColors = makesymbolspec('Polygon', {'PopDens2000', ...
-    [0 maxdensity], 'FaceColor', fall});
-
-axesm('MapProjection', 'eqaconic', 'MapParallels', [],...
-  'MapLatLimit', MapLatLimit, 'MapLonLimit', MapLonLimit,...
-  'GLineStyle', '-')
-geoshow(NEstates, 'DisplayType', 'polygon', ...
-    'SymbolSpec', densityColors)

### 2021-12-17 spatial-model/pdeTest2.m

@@ -1,65 +1,0 @@
-%% PDE Test 2
-% u_t = u_xx + u for 0 < x < pi, t > 0
-% u = 0          at x = 0, pi
-% u(x,0) = phi(x)
-
-clear
-clc
-
-%% Step parameters
-J = 100; % Number of discrete steps
-delta_x = pi/J; % Space mesh
-s = 5/11; % s = delta_t / (delta_x)^2 < 1/2
-
-%% Define matrix
-
-% Identity matrix
-I = eye(J+1);
-
-% Tridiagonal matrix
-D = diag(-2*s*ones(1,J+1)) + diag(0.5*s*delta_x^2*ones(1,J+1)) + diag(s*ones(1,J),1) + diag(s*ones(1,J),-1);
-D(1,:) = 0;
-D(J+1,:) = 0;
-
-% Full scheme matrix
-A = I + D;
-
-%% Define initial values
-
-for j=0:J
-    u(j+1,1) = phi(j*delta_x);
-end
-
-%% Define time steps
-
-% Define end time
-T = 3;
-
-% Calculate required time steps
-delta_t = s*delta_x^2; % calculate delta_t
-N = ceil(T/delta_t); % calculate required steps
-
-%% Iteration
-
-for n=2:N
-    u(:,n) = A*u(:,n-1);
-end
-
-%% Plot results
-[Y,X] = meshgrid(1:N,0:delta_x:pi);
-surf(X,Y,u)
-xlabel('$x$','interpreter','latex')
-ylabel('$t$','interpreter','latex')
-zlabel('$u(x,t)$','interpreter','latex')
-set(gca, 'TickLabelInterpreter','latex')
-
-%% Extra functions
-
-% Initial value function
-function u0 = phi(x)
-    if x >= 0 && x <= pi/2
-        u0 = x;
-    elseif x > pi/2 && x <= pi
-        u0 = pi - x;
-    end
-end

### 2021-12-17 version2/betafun.m

@@ -1,15 +1,0 @@
-function b = betafun(p,t_q,t)
-    % Define parameters
-    beta = p(1);
-    eta = p(2);
-
-    % Create a t_q vector
-    t_q = t_q * ones(length(t),1);
-
-    % Calculate beta at time t
-    b = 1 - tanh(2 * (t - t_q));
-    b = b / 2;
-    b = (1 - eta) * b;
-    b = eta + b;
-    b = beta * b;
-end

### 2021-12-17 version2/computemodel.m

@@ -1,23 +1,0 @@
-function [T,y] = computemodel(p,q,trange,pops)
-    % Separate the parameters from the population fractions
-    p1 = p(1:end-2); % Parameters
-    e0i0 = p(end-1); % Exposed/Infected
-    a0i0 = p(end); % Asymptotic/Infected
-
-    % Calculate the initial values
-    n = pops(1);
-    i0 = pops(2);
-    h0 = pops(3);
-    d0 = pops(4);
-    e0 = e0i0 * i0;
-    a0 = a0i0 * i0;
-    ar0 = 0;
-    r0 = 0;
-    s0 = n - (e0 + a0 + ar0 + i0 + h0 + r0 + d0);
-    
-    % Define the initial population vector
-    y0 = [s0, e0, a0, ar0, i0, h0, r0];
-
-    % Compute the solution vector
-    [T,y] = ode45(@(t,y)sir(t,y,p1,q),trange,y0);
-end

### 2021-12-17 version2/loadData.m

@@ -1,16 +1,0 @@
-function [time, cases, deceased] = loadData(file)
-% LOADDATA   Imports time series data.
-%   [time, cases, deceased] = loadData(file) imports the first two
-%   columns from file.csv as cases and deceased, and calculates the
-%   time steps.
-
-    % Import the time series csv file
-    series = load(file);
-
-    % Define the caes
-    cases = series(:,1)';
-    deceased = series(:,2)';
-
-    % Create time steps
-    time =  1:length(cases);
-end

### 2021-12-17 version2/objfun.m

@@ -1,22 +1,0 @@
-function obj = objfun(trange,Cobs,Dobs,p,q,pops)
-% OBJFUN   Compute objective function.
-%   obj = objfun(trange,Yobs,p,q,y0) compute norm2(Ynum - Yobs)
-%   where Ynum is the solution to sir(t,y,p1) over trange with
-%   initial conditions y0 and p1 = [p,q].
-
-    % Compute the solution vector
-    [~,y] = computemodel(p,q,trange,pops);
-    
-    % Convert to system vectors
-    i = y(:,5);
-    h = y(:,6);
-    r = y(:,7);
-    d = y(:,8);
-    
-    % Define model cases and deaths
-    Cnum = i + h + r + d;
-    Dnum = d;
-    
-    % Compute the objective function
-    obj = sum(vecnorm(log(Cnum) - log(Cobs)) + vecnorm(log(Dnum) - log(Dobs)));
-end

### 2021-12-17 version2/optim.m

@@ -1,69 +1,0 @@
-%% Optimization Program
-%
-%
-%% 
-clear
-close all
-clc
-
-%% Fixed Parameters
-
-% Load fixed parameters
-load('params.mat');
-
-%% Optimization Parameter Bounds
-% Bounds are in the form:
-% [beta_si, beta_sa, eta_si, eta_sa, sigma_i, sigma_a, m, m_ar, chi, psi,
-% gamma, omega, e0/i0, a0/i0]
-
-% Lower bounds
-lb = [0, 0, 0, 0, 2, 2, 5, 5, 5, 5, 0.25, 0.1, 1, 1];
-
-% Upper bounds
-ub = [1, 1, 1, 1, 7, 7, 12, 12, 20, 20, 0.75, 0.5, 5, 5];
-
-%% Load Observed Data
-[time, Cobs, Dobs] = loadData("time_series.csv");
-
-% Create y vector
-Yobs = [Cobs(:); Dobs(:)];
-
-% Define end of simulation time
-t_f = length(time);
-
-%% Optimum Solution
-
-% Define number of iterations
-iters = 1;
-    
-% Define fixed parameters
-q = [mu, t_q];
-pops = [n, i0, h0, d0];
-
-% Preallocate the estimate matrix
-pEst = zeros(iters,length(lb));
-
-% Perform optimization iterations
-tic % Start timer
-for iter=1:iters
-
-    % Sample starting parameters
-    %p0 = sampleparams(lb,ub,n);
-    p0 = (lb + ub) / 2;
-    
-    % Compute parameter estimate
-    pEst(iter,:) = fmincon(@(p)objfun(time,Cobs,Dobs,p,q,pops),p0,[],[],[],[],lb,ub);
-
-    disp("Iteration: " + num2str(iter))
-end
-toc % End timer
-
-%% Export Results
-% Compute the final solution vector from the median of the estimates
-pFin = median(pEst,1);
-
-% Compute the interquartile range
-iqrange = iqr(pEst,1);
-
-% Save the results to a file
-save('optimfile.mat','pFin','iqrange')

### 2021-12-17 version2/params.m

@@ -1,25 +1,0 @@
-%% Define Fixed Parameters
-% Set the known parmaeters here.
-% These are exported to the other main programs.
-
-%% Clear MATLAB
-clear
-close all
-clc
-
-%% Set Values
-
-% Population parameters
-n = 2695598; % Total population
-i0 = 160; % Initial infected
-h0 = 38; % Initial hospitalized
-d0 = 3; % Initial deceased
-
-% Non-COVID deaths
-mu = 1.8997e-5;
-
-% Lockdown start
-t_q = 4;
-
-% Export fixed parameters
-save('params.mat')

### 2021-12-17 version2/results.m

@@ -1,73 +1,0 @@
-%% Results Program
-%
-%
-%% Clear MATLAB
-clear
-close all
-clc
-
-%% Fixed Parameters
-
-% Load fixed parameters
-load('params.mat');
-
-% Define fixed parameters
-q = [mu, t_q];
-pops = [n, i0, h0, d0];
-
-%% Load Optimized parameters
-load('optimfile.mat', 'pFin');
-
-%% Load Observed Data
-[time, Cobs, Dobs] = loadData("time_series.csv");
-
-% Define end of simulation time
-t_f = length(time);
-
-%% Compute solution
-
-% Compute the solution vector
-[T,y] = computemodel(pFin,q,time,pops);
-
-% Convert to system vectors
-s = y(:,1);
-e = y(:,2);
-ar = y(:,3);
-a = y(:,4);
-i = y(:,5);
-h = y(:,6);
-r = y(:,7);
-d = y(:,8);
-
-% Define model cases and deaths
-Cnum = i + h + r + d;
-Dnum = d;
-
-%% Plot Results
-% Total cases
-cFig = figure(1);
-semilogy(T,Cnum,'r-',time,Cobs,'k.')
-legend('$C_\mathrm{num}(t)$','$C_\mathrm{obs}(t)$','Interpreter','latex','Location','southeast')
-xlabel('$t$ (days from March 18)','interpreter','latex')
-ylabel('$C(t)$','interpreter','latex')
-set(gca,'TickLabelInterpreter','latex')
-
-% Total deaths
-dFig = figure(2);
-semilogy(T,Dnum,'r-',time,Dobs,'k.')
-legend('$D_\mathrm{num}(t)$','$D_\mathrm{obs}(t)$','Interpreter','latex','Location','southeast')
-xlabel('$t$ (days from March 18)','interpreter','latex')
-ylabel('$D(t)$','interpreter','latex')
-set(gca,'TickLabelInterpreter','latex')
-
-% Export cases plot
-set(cFig,'Units','Inches');
-pos = get(cFig,'Position');
-set(cFig,'PaperPositionMode','Auto','PaperUnits','Inches','PaperSize',[pos(3), pos(4)])
-print(cFig,'~/Documents/GitHub/chicago-spatial-covid/update-10-22/cases','-dpdf','-r0')
-
-% Export deaths plot
-set(dFig,'Units','Inches');
-pos = get(dFig,'Position');
[…]

### 2021-12-17 version2/rndSample.m

@@ -1,10 +1,0 @@
-function r = rndSample(range,n)
-% RNDSAMPLE   Calculate a normally distributed random sample.
-%   r = rndSample(range,n) cacluates a size-n sample from
-%   range = [range(1),range(2)].
-
-    a = range(1);
-    b = range(2);
-    
-    r = (b - a) .* rand(n,1) + a;
-end

### 2021-12-17 version2/sampleparams.m

@@ -1,43 +1,0 @@
-function p0 = sampleparams(lb,ub,n)
-    % Transmission rates
-    beta_si = unifrnd(lb(1),ub(1)); % S -> I [per day]
-    beta_sa = unifrnd(lb(2),ub(2)); % S -> A [per day]
-    
-    % Lockdown effects
-    eta_si = unifrnd(lb(3),ub(3)); % S -> I
-    eta_sa = unifrnd(lb(4),ub(4)); % S -> A
-    
-    % Incubation period
-    sigma_i = 1 / unifrnd(lb(5),ub(5)); % E -> I [days]
-    
-    % Latent period
-    sigma_a = 1 / unifrnd(lb(6),ub(6)); % E -> I [days]
-    
-    % Infectivity period
-    m = 1 / unifrnd(lb(7),ub(7)); % [days]
-    
-    % Recovery periods
-    m_ar = 1 / unifrnd(lb(8),ub(8)); % A -> AR [days]
-    chi = 1 / unifrnd(lb(9),ub(9)); % H -> R [days]
-    
-    % Period to deceased
-    psi = 1 / unifrnd(lb(10),ub(10)); % H -> D [days]
-    
-    % Conversion fractions
-    gamma = unifrnd(lb(11),ub(11)); % I -> H, I -> R
-    omega = unifrnd(lb(12),ub(12)); % H -> D, H -> R
-    
-    % Initial population fractions
-    e0i0 = unifrnd(lb(13),ub(13)); % Exposed/Infected
-    a0i0 = unifrnd(lb(13),ub(13)); % Asymptomatic/Infected
-
-    if 0 < omega && omega <= 1
-        %disp('Condition 1 stable')
-    elseif 1 < omega && chi < (psi * omega) / (omega - 1)
-        %disp('Conditon 2 stable')
-    else
-        disp('Unstable')
-    end
-
-    p0 = [sigma_a, sigma_i, m_ar, m, gamma, omega, chi, psi, beta_sa / n, beta_si / n, eta_sa, eta_si, e0i0, a0i0];
-end

### 2021-12-17 version2/sir.m

@@ -1,39 +1,0 @@
-function dydt = sir(t,y,p,t_q)
-% SIR   SIR model function.
-
-    % Derivatives with respect to time
-    s = y(1);
-    e = y(2);
-    a = y(3);
-    i = y(5);
-    h = y(6);
-    
-    % System parameters
-    % Optimized parameters (p vector)
-    alpha_a = p(1);
-    alpha_i = p(2);
-    eta_a = p(3);
-    eta_i = p(4);
-    lambda_a = p(5);
-    lambda_i = p(6);
-    xi = p(7);
-    gamma = p(8);
-    omega = p(9);
-    theta = p(10);
-    kappa = p(11);
-    
-    % Calculate time dependent betas
-    alpha_i_t = betafun([alpha_i,eta_i],t_q,t);
-    alpha_a_t = betafun([alpha_a,eta_a],t_q,t);
-
-    % Define system
-    dydt = [
-        -alpha_a_t .* (s * a) - alpha_i_t .* (s * i); % s_t
-        alpha_a_t .* (s * a) + alpha_i_t .* (s * i) - (lambda_a + lambda_i) * e; % e_t
-        lambda_a * e - xi * a; % a_t
-        xi * a; % ar_t
-        lambda_i * e - i; % i_t
-        gamma * i - (1 - omega) * theta * h - omega * kappa * h; % h_t
-        (1 - gamma) * i + (1 - omega) * theta * h % r_t
-        ];
-end

### 2021-12-17 version2/smoothTest.m

@@ -1,89 +1,0 @@
-function smoothTest
-
-n=100;          % number of intervals (i.e. parametric curve would be evaluted n+1 times)
-
-p =[0.5887    0.9819
-0.5887    0.9819
-0.5597    1.3551
-0.5565    1.7749
-0.4081    1.8274
-0.4500    1.4834
-0.3016    1.5300
-0.2210    1.3901
-0.2242    1.2152
-0.2790    1.0169
-0.4500    0.9295
-0.5790    0.9236
-0.5887    0.9819];
-Px = p(:,1); Py = p(:,2);
-% Note first and last points are repeated so that spline curve passes
-% through all points
-
-% when Tension=0 the class of Cardinal spline is known as Catmull-Rom spline
-Tension=0; 
-figure
-set1x=[];
-set1y=[];
-for k=1:length(Px)-3
-    [xvec,yvec]=EvaluateCardinal2DAtNplusOneValues([Px(k),Py(k)],[Px(k+1),Py(k+1)],[Px(k+2),Py(k+2)],[Px(k+3),Py(k+3)],Tension,n);
-    set1x=[set1x, xvec];
-    set1y=[set1y, yvec];    
-end
-
-Tension=0.5; 
-set2x=[];
-set2y=[];
-for k=1:length(Px)-3
-     [xvec,yvec]=EvaluateCardinal2DAtNplusOneValues([Px(k),Py(k)],[Px(k+1),Py(k+1)],[Px(k+2),Py(k+2)],[Px(k+3),Py(k+3)],Tension,n);   
-    set2x=[set2x, xvec];
-    set2y=[set2y, yvec];    
-end
-plot(set1x,set1y,'g',set2x,set2y,'b','linewidth',2)
-
-% no longer need. free the memory
-clear set1x, clear set1y, clear set2x, clear set2y, clear xvec, clear yvec 
-
-h = legend('Tension=0 (Catmull-Rom)','Tension=0.5');
-hold on
-
-plot(Px,Py,'ro','linewidth',2) % plot control points
-title('\bfCardinal Spline')
-hold off
-end
-
-function [xvec,yvec]=EvaluateCardinal2DAtNplusOneValues(P0,P1,P2,P3,T,N)
-
-xvec=[]; yvec=[];
-
-% u vareis b/w 0 and 1.
[…]

### 2021-12-17 version2/spatialModel.m

@@ -1,45 +1,0 @@
-clear
-clc
-close all;
-
-domain = shaperead('boundaries.shp','UseGeoCoords',true);
-
-pgon = simplify(polyshape(domain.Lon,domain.Lat));
-plot(pgon)
-axis equal
-
-tr = triangulation(pgon);
-
-model = createpde;
-
-tnodes = tr.Points';
-telements = tr.ConnectivityList';
-
-geometryFromMesh(model,tnodes,telements);
-
-dFig = figure(1);
-pdegplot(model)
-xlabel('$x$','interpreter','latex')
-ylabel('$y$','interpreter','latex')
-set(gca,'TickLabelInterpreter','latex')
-
-mFig = figure(2);
-pdemesh(model)
-xlabel('$x$','interpreter','latex')
-ylabel('$y$','interpreter','latex')
-set(gca,'TickLabelInterpreter','latex')
-
-% Export domain plot
-set(dFig,'Units','Inches');
-pos = get(dFig,'Position');
-set(dFig,'PaperPositionMode','Auto','PaperUnits','Inches','PaperSize',[pos(3), pos(4)])
-print(dFig,'~/Documents/GitHub/chicago-spatial-covid/update-10-22/domain','-dpdf','-r0')
-
-% Export mesh plot
-set(mFig,'Units','Inches');
-pos = get(mFig,'Position');
-set(mFig,'PaperPositionMode','Auto','PaperUnits','Inches','PaperSize',[pos(3), pos(4)])
-print(mFig,'~/Documents/GitHub/chicago-spatial-covid/update-10-22/mesh','-dpdf','-r0')
-
-
-%generateMesh(model)

### 2021-12-20 version3/version3.tex
Report updates

@@
 \usepackage{longtable}
+\usepackage{float}
 
@@
 \usepackage{graphicx}
-%\usepackage{subfig}
 \graphicspath{ {./images/} }
@@
 	\subsection{Confirmed and death data}
-		In this study, we used the publicly available data set of COVID-19 provided by the City of Chicago Data Portal \cite{Chicago-cases}.
-		This data set includes daily counts of confirmed infected cases, hospitalizations, and deaths.
+		In this study, we used the publicly available data sets of COVID-19 metrics provided by the City of Chicago Data Portal.
+		\cite{Chicago-cases} includes daily counts of city-wide confirmed infected cases, hospitalizations, and deaths.
+		Weekly cases, tests and deaths by ZIP code are logged in \cite{Chicago-zips}.
 
@@
 
-		We motivate the use of this simple model for the following reasons:
+		A few notes motivating this very simple model are necessary.
 		\begin{itemize}
 			\item
-				We do not consider the $E$ group in this work because we feel this population can be merged with the $A$ group.
+				We do not consider the ``exposed'' $E$ group in this work.
+				Because the members of $A$ have separate infection and recovery rates, and because there is a possibility they are detected and move to $I$, we consider $E$ to be merged with $A$.
 			\item
-				We do not distinguish between quarantined, hospitalized, or nursing-home populations.
-				Including these additional compartments would over parameterize the model.
-				Additionally, the way in which data has been collected and is provided by authorities has varied over time, making its usage rather difficult \cite{sara-data}.
-				For the same reasons, we do not distinguish between deceased and recovered.
+				We do not distinguish between quarantined, hospitalized, or nursing home populations.
+				Nursing homes are now known to be epicenters of the virus, however the true case rates are notoriously difficult to track \cite{nursing-homes}.
+				Additionally, these three population groups can be assumed to be stationary in nature, and in the case of quarantined and hospitalized, isolated.
+				Therefore we ignore their individual contributions.
+			\item
+				In general, the way in which data has been collected and is provided by authorities has varied over time, making its usage rather difficult\footnote{For an in-depth analysis of the data collection problem, see \cite{sara-soremo}.} \cite{messy-data1}, \cite{messy-data2}, \cite{challenges}.
+				There is also the question of case counts being influenced by access to testing, especially early in the pandemic.
+				Including additional compartments would over-parameterize the model making it more difficult to verify, and would not provide useful information.
+			\item
+				Lastly, adding additional parameters reduces the computational feasibility of the optimization problem.
 		\end{itemize}
+		The goal of this model is to strike a reasonable balance between representation of the pandemic and the populations involved, and feasibility of data collection and computation.
 
@@
 
-	\subsection{Numerical discretization}
-
-		while the diffusion coefficient is
+		The diffusion coefficient is also assumed to be altered by the lockdown, and is updated as a simple step function
 		\begin{equation} \label{eq:diffusion}
-		D (t) = D_0  - \eta H(t - t_q)
+			D (t) = D_0 \left[ 1 - \eta H(t - t_q) \right]
 		\end{equation}
 		where $H(t)$ is the Heaviside step function.
+		The average one-way commute in Chicago is about 5 miles \cite{travel}.
+		Thus, we set $D_0$ to the fixed value of $(5 / 0.72)^2 / 4 \Delta t$ where $\Delta t$ is the time step \cite{methods}.
+
+	\subsection{Numerical discretization}
+		From the map of the city, the computational domain $\Omega$ is defined as the minimum square enclosing the city.
+		The city is contained by $\Omega' = \left\{ (X,Y) \mid X \in [-87.9397, -87.5245], \ Y \in [41.6447, 42.023] \right\}$ where $X$ represents degrees latitude and $Y$ represents degrees longitude.
+		Note that degrees latitude and longitude are not equal when converted to miles.
+		Therefore, we define the computational domain such that the grid spacing is approximately 0.72 miles in both axes.
+		This gives the domain
+		\begin{equation} \label{eq:domain}
+			\Omega = \left\{ \vec{x} \in \R^2 \mid x \in [0, 40], \ y \in [0, 29] \right\}.
+		\end{equation}
+
+		The initial spatial distribution of the infected population is determined by sampling the ZIP code data of confirmed cases \cite{Chicago-zips} at the grid points, as shown in Figure \ref{fig:seeding}.
+
+		The model is solved via finite differences, using the \verb|scikit-fdiff| Python module \cite{nicolas_cellier_2019_3236970}.
+
+		\begin{figure}[h!]
+			\centering
+			\begin{subfigure}{0.5\textwidth}
+				\includegraphics[width=\textwidth]{t0-cases}
+				\caption{ZIP data.}
+			\end{subfigure}%
+			\begin{subfigure}{0.5\textwidth}
+				\includegraphics[width=\textwidth]{infected_0}
+				\caption{Seeded model.}
+			\end{subfigure}
+			\caption{Initial seeding of the model with data from March 18, 2020.}
+			\label{fig:seeding}
+		\end{figure}
+
 \section{Results}
@@
 		That is, $t_\mathrm{bol} =$ March 21, 2020 and $t_\mathrm{eol} =$ May 29, 2020.
+		In Figure \ref{fig:params}, the table shows the estimated parameters.
+		Figure \ref{fig:fit} compares the data and the fitted symptomatic infected populations.
 		\begin{figure}[h!]
@@
 
-	\subsection{Spatial spread of COVID-19}
+	\subsection{Spatial spread of COVID-19} \label{sec:spread}
+		Simulations are performed from March 18, 2020 to June 24, 2020.
+		The time step is $\Delta t = 0.2$ [days].
+
+		Figure \ref{fig:spatial-results} presents three of the days from the simulation time: the effective lockdown $t_q$ day, the last day of simulation, and arbitrary day from the period in between.
+		The observed data is shown on the left, while the model results are on the right.
+		The data shows that the main progression of the disease is from the lake, to the southwest, then to center-west, then slowly decreases through most of the city.
+		In contrast, the model infected population is an order of magnitude different than the true population after only 15 days, and the difference increases as the simulation continues.
+		Furthermore, the hoped-for diffusion behavior is not present in the model results.
+
+		In short, the proposed model does \textit{not} reproduce the spatial propagation of the virus.
+		\begin{figure}[H]
+			\centering
+			\begin{subfigure}{0.5\textwidth}
+				\includegraphics[width=\textwidth]{tq-cases}
+				\caption{Observed Day 15 ($t_q$): April 2, 2020}
+			\end{subfigure}%
+			\begin{subfigure}{0.5\textwidth}
+				\includegraphics[width=\textwidth]{infected_15}
+				\caption{Model Day 15 ($t_q$): April 2, 2020}
+			\end{subfigure}
+
+			\begin{subfigure}{0.5\textwidth}
+				\includegraphics[width=\textwidth]{tmid-cases}
+				\caption{Observed Day 35: April 22, 2020}
+			\end{subfigure}%
+			\begin{subfigure}{0.5\textwidth}
+				\includegraphics[width=\textwidth]{infected_35}
+				\caption{Model Day 35: April 22, 2020}
+			\end{subfigure}
+
+			\begin{subfigure}{0.5\textwidth}
+				\includegraphics[width=\textwidth]{tfin-cases}
+				\caption{Observed Day 99: June 24, 2020}
+			\end{subfigure}%
+			\begin{subfigure}{0.5\textwidth}
+				\includegraphics[width=\textwidth]{infected_99}
+				\caption{Model Day 99: June 24, 2020}
+			\end{subfigure}
+
+			\caption{Comparison of observed infected cases and the model results.}
+			\label{fig:spatial-results}
+		\end{figure}
 
 \section{Discussion}
+	As mentioned in \S \ref{sec:spread}, the proposed model does not reproduce the spatial propagation of the virus.
+	In this section, we address possible causes of these inaccuracies and discuss proposals for improving the results.
 
-	\textbf{Acknowledgments:} \\
-	\textbf{Funding:}
-	This research was funded by the Socially Responsible Modeling, Computation, and Design (SoReMo) Initiative, at the Illinois Institute of Technology. \\
-	\textbf{Conflicts of Interest:}
+	The selected model populations may not, contrary to the assumption, be sufficient to describe the virus dynamics.
+	Figure \ref{fig:fit} suggests this conclusion, since the growth of the model population only roughly describes the observed data.
+	Likely, the Exposed compartment describing the latent period is necessary, as in \cite{Mammeri+2020+102+113}.
+	It is possible that the assumption that Deceased and Recovered populations can be modeled by the same population compartment is errant.
+	On the other hand, in either case these populations have no effect on the act of transmission, so the model dynamics would likely be similar for both cases.
+	A Deceased population may be useful for other reasons, as discussed below.
 
@@
 			\textbf{Latin symbols} & & \\
-			$A$ & Asymptomatic infected individuals & [individuals] \\
+			$A$ & Asymptomatic infected individuals & [1] \\
 			$D$ & Diffusivity & \\
-			$I$ & Symptomatic infected individuals & [individuals] \\
-			$N$ & Total population of individuals & [individuals] \\
-			$R$ & Removed individuals & [individuals] \\
-			$S$ & Susceptible individuals & [individuals] \\
+			$I$ & Symptomatic infected individuals & [1] \\
+			$R$ & Removed individuals & [1] \\
+			$S$ & Susceptible individuals & [1] \\
 			\textbf{Greek symbols} & & \\
@@
 			$\delta$ & Rate of asymptomatic individuals that may develop symptoms & [days\textsuperscript{-1}] \\
-			$\omega$ & Contacts & 
+			$\omega$ & Contacts & [1]
 		\end{longtable}
+
+	\newpage
+	\subsection*{Model Populations}
+		\begin{figure}[H]
+			\centering
+			\begin{subfigure}{0.5\textwidth}
+				\includegraphics[width=\textwidth]{susceptible_15}
+				\caption{Susceptible Day 15 ($t_q$): April 2, 2020}
+			\end{subfigure}%
+			\begin{subfigure}{0.5\textwidth}
+				\includegraphics[width=\textwidth]{asymptomatic_15}
+				\caption{Asymptomatic Day 15 ($t_q$): April 2, 2020}
+			\end{subfigure}
+
+			\begin{subfigure}{0.5\textwidth}
+				\includegraphics[width=\textwidth]{susceptible_35}
+				\caption{Susceptible Day 35: April 22, 2020}
+			\end{subfigure}%
+			\begin{subfigure}{0.5\textwidth}
+				\includegraphics[width=\textwidth]{asymptomatic_35}
+				\caption{Asymptomatic Day 35: April 22, 2020}
+			\end{subfigure}
+
+			\begin{subfigure}{0.5\textwidth}
+				\includegraphics[width=\textwidth]{susceptible_99}
+				\caption{Susceptible Day 99: June 24, 2020}
+			\end{subfigure}%
+			\begin{subfigure}{0.5\textwidth}
+				\includegraphics[width=\textwidth]{asymptomatic_99}
+				\caption{Asymptomatic Day 99: June 24, 2020}
+			\end{subfigure}
+
+			\caption{Model results for susceptible and asymptomatic populations.}
+			\label{fig:app-pops}
+		\end{figure}
+
 \bibliographystyle{abbrv}

### 2021-12-20 data/final_zip_data.py

@@ -1,6 +1,8 @@
 import pandas as pd
 import numpy as np
 
+path = "/home/trent/Documents/GitHub/chicago-spatial-covid/version3/"
+
 df = pd.read_csv("COVID-19_Cases__Tests__and_Deaths_by_ZIP_Code.csv")
 
 cols = ["ZIP Code", "Week Start", "Cases - Weekly", "Deaths - Cumulative", "Population"]
@@ -8,11 +10,12 @@ df = df[cols]
 
 #df = df[df["Week Start"] == "04/19/2020"]
 #df = df[df["Week Start"] == "03/29/2020"]
-df = df[df["Week Start"] == "06/21/2020"]
+#df = df[df["Week Start"] == "06/21/2020"]
+df = df[df["Week Start"] == "03/15/2020"]
 df = df[df["ZIP Code"] != "Unknown"]
 df["Cases - Weekly"] = df["Cases - Weekly"].replace(np.nan, 0)
 
 cols = ["ZIP Code", "Cases - Weekly", "Deaths - Cumulative", "Population"]
 df = df[cols]
 
-df.to_csv("tend_data.csv", index=False, header=False)
+df.to_csv(path + "t0_data.csv", index=False, header=False)

### 2021-12-20 version3/box_plot.py

@@ -14,10 +14,10 @@ df.columns = columns
 
 # Create a boxplot
 boxplot = df.boxplot(column=columns)
+boxplot.plot()
 
 # Export plot
-plt.savefig('box-plot.pdf')
+plt.savefig('/home/trent/Documents/GitHub/chicago-spatial-covid/version3/images/box-plot.pdf')
 
 # Plot the boxplot
-boxplot.plot()
 plt.show()

### 2021-12-20 version3/spatial.py

@@ -4,6 +4,9 @@ import numpy as np
 import csv
 import matplotlib.pyplot as plt
 
+# Image path
+impath = '/home/trent/Documents/GitHub/chicago-spatial-covid/version3/images/'
+
 # LaTeX font for matplotlib
 import matplotlib
 matplotlib.rcParams['mathtext.fontset'] = 'stix'
@@ -11,13 +14,18 @@ matplotlib.rcParams['font.family'] = 'STIXGeneral'
 
 
 # Time points
-tq = 11 # Lockdown point tq = (tbol + teol) / 5
+tbol = 4 # Beginning of lockdown
+teol = 73 # End of lockdown
 tmid = 35 # Test point in the middle of fitting
 tfin = 99 # Last day of fitting
+tq = (tbol + teol) / 5 # Calculate tq
+
+# Time step
+dt = 0.2
 
 # Parameters
 eta = 0.5514 # Lockdown scale factor
-D = 1.105 # Diffusivity
+D = (5 / 0.72) ** 2 / (4 * dt) # Diffusivity
 omegaA = 0.1969 # Transmission rate due to A (initial)
 omegaI = 0.3061 # Transmission rate due to I (initial)
 gammaA = 0.3632 # Removal rate of A
@@ -26,8 +34,11 @@ l = 0.0939 # Symptom onset rate
 
 # Time dependent beta function
 def omegabeta(omega0beta,eta,t,tq):
-    result = omega0beta * (eta + (1 - eta) * (1 - np.tanh(2 * (t - tq))) / 2)
-    return result
+    return omega0beta * (eta + (1 - eta) * (1 - np.tanh(2 * (t - tq))) / 2)
+
+# Time dependent D function
+def diffusion(D,eta,t,tq):
+    return D * (1 - eta * np.heaviside(t - tq,1))
 
 
 ## PDE Model
@@ -41,19 +52,24 @@ model = Model(["D * (dxxs + dyys) - (omegaA * a + omegaI * i) * s",
               boundary_conditions="noflux") # No flux boundary condition (homogeneous Neumann)
 
 # Computational domain
-L = 40 # Space points
-x = y = np.linspace(0, L, L) # Domain space
+xpoints = 40
+ypoints = 29
+x = np.linspace(0, xpoints, xpoints)
+y = np.linspace(0, ypoints, ypoints)
 
 # Hook function for time dependent parameters
 def hook(t, fields):
[…]

### 2021-12-20 version3/ziptest.m

@@ -7,7 +7,7 @@ h = height(zips);
 faceColors = makesymbolspec('Polygon',...
     {'INDEX',[1 h],'FaceColor',polcmap(h)});
 
-findata = load("tend_data.csv");
+findata = load("t0_data.csv");
 
 for i=1:length(zips)
     zip = str2num(zips(i).zip);
@@ -32,12 +32,16 @@ densityColors = makesymbolspec('Polygon', {'objectid', [0 maxinfected], 'FaceCol
 
 cFig = figure(1)
 mapshow(zips,'SymbolSpec',densityColors)
+xlabel('$x$','interpreter','latex')
+ylabel('$y$','interpreter','latex')
+set(gca,'TickLabelInterpreter','latex')
 
 caxis([0 maxinfected])
 colormap(fall)
-colorbar
+c = colorbar;
+set(c,'TickLabelInterpreter','latex')
 
 set(cFig,'Units','Inches');
 pos = get(cFig,'Position');
 set(cFig,'PaperPositionMode','Auto','PaperUnits','Inches','PaperSize',[pos(3), pos(4)])
-print(cFig,'~/Documents/GitHub/chicago-spatial-covid/version3/tend-cases','-dpng','-r0')
+print(cFig,'~/Documents/GitHub/chicago-spatial-covid/version3/images/t0-cases','-dpdf','-r0')

### 2021-12-20 version3/version3.tex
spatial model update

@@
 
+	Even if the model is based on faulty assumptions, we would still expect to see some evidence of diffusion in the spatial simulations.
+	Examining the simulated populations of $S$ and $A$ (Figure \ref{fig:app-pops}), we immediately see there is a problem with the spatial simulation.
+	The bounds on the computation domain are clearly not sufficient to describe the boundaries of the city, since the $S$ and $A$ populations almost entirely diffuse into Lake Michigan.
+	As a result, there are very few int
+
 \section*{Appendix}

### 2021-12-20 version3/spatial.py

@@ -5,7 +5,8 @@ import csv
 import matplotlib.pyplot as plt
 
 # Image path
-impath = '/home/trent/Documents/GitHub/chicago-spatial-covid/version3/images/'
+#impath = '/home/trent/Documents/GitHub/chicago-spatial-covid/version3/images/'
+impath = '/Users/trentgerew/GitHub/chicago-spatial-covid/version3/images/'
 
 # LaTeX font for matplotlib
 import matplotlib
@@ -21,7 +22,7 @@ tfin = 99 # Last day of fitting
 tq = (tbol + teol) / 5 # Calculate tq
 
 # Time step
-dt = 0.2
+dt = 0.1
 
 # Parameters
 eta = 0.5514 # Lockdown scale factor
@@ -84,7 +85,7 @@ i1 = np.genfromtxt('initialI.csv', dtype=float, delimiter=",")
 initial_fields = model.Fields(x=y, y=x, s=s1, a=a1, i=i1, D=D, omegaA=omegaA, omegaI=omegaI, gammaA=gammaA, gammaI=gammaI, l=l)
 
 # Create and run the simulation
-simulation = Simulation(model, initial_fields, dt=dt, tmax=tfin, hook=hook)
+simulation = Simulation(model, initial_fields, dt=dt, tmax=tfin)
 container = simulation.attach_container()
 t, fields = simulation.run()
 

### 2021-12-21 version3/version3.tex
Draft 1 completed

--- 2021-12-20 version3/version3.tex
@@
 \begin{abstract}
-	content...
+	We investigate whether a reaction-diffusion model considering only meanly daily movement is sufficient to describe the spread of the COVID-19 virus in Chicago.
+	The model is calibrated using publicly available health data published by the city.
+	We first study the system of partial differential equations, then derive the basic reproduction number $\Ro$.
+	Then we consider the numerical simulations conducted from March 18 to June 24, 2020.
+	Finally, we discuss shortcomings of the model, and offer some potential solutions.
 \end{abstract}
@@
 \section{Introduction}
+	Emerging from Wuhan, China in late 2019, the COVID-19 virus rapidly spread throughout the globe \cite{covid-name}.
+	Since the outbreak, governments everywhere have been desperately trying to contain the pandemic.
+	Now, with safe and effective vaccines being distributed, it is hoped the pandemic can be overcome \cite{vaccine}.
+	Before the introduction of the vaccines, so-called ``non-therapeutic'' interventions \cite{interventions1} were frequently deployed.
+	Prominent among these are social distancing, self quarantining, lockdowns, gathering limitations, and the use of personal protective equipment.
+	Now with the Omicron variant on the rise \cite{omicron}, these interventions are once again becoming prominent with some European countries reintroducing lockdowns and travel restrictions \cite{lockdowns}.
+
+	Many mathematical models have been proposed to study the spread of the virus.
+	Most are compartment models, where a population is divided into groups according to the state of individuals.
+	These are generally known as SIR type models.
+	The majority of such models ignore any spatial components.
+	\cite{s+t+spain} employs a SAIR model and includes mobility terms based on the position of cell phones to model the propagation of the disease through Spain.
+	Similarly, \cite{Danon2020.02.12.20022566} incorporates daily movements into an SEIR model.
+	In a different approach, \cite{giuliani2020modelling} considers a statistical model to handle diffusion of COVID-19 in Italy.
+
+	Here, we follow \cite{Kevrekidis-2021} and \cite{Mammeri+2020+102+113} in building a virus model where the spatial propagation is modeled by a diffusion and the reactions are derived from an SAIR model.
+	Specifically, we examine the spread of COVID-19 in the city of Chicago.
+	Does a reaction-diffusion model, which considers only the average daily movements of the population, correctly describe the spread of the virus in Chicago?
+	If it does, we can use the model to assess possible vaccination strategies or reopening scenarios.
+
+	The paper is organized as follows.
+	Section 2 outlines the methodology, including data collection, modeling assumptions, and parameter estimates.
+	Section 3 contains a brief analysis of the model and the results.
+	In Section 4, we offer a discussion of the shortcomings of the model, and how they can be improved.
 
@@
 			\item
-				Lastly, adding additional parameters reduces the computational feasibility of the optimization problem.
+				Lastly, adding additional compartments and parameters reduces the computational feasibility of the optimization problem.
 		\end{itemize}
@@
 			\begin{aligned}
-				S_t - D(t) \Delta S &= - \omega(t) \frac{\beta_A A + \beta_I I}{N} S, \\
-				A_t - D(t) \Delta A &= \omega(t) \frac{\beta_A A + \beta_I I}{N} S - (\gamma_A - \delta) A, \\
+				S_t - D(t) \Delta S &= - \omega(t) \left( \beta_A A + \beta_I I \right) S, \\
+				A_t - D(t) \Delta A &= \omega(t) \left( \beta_A A + \beta_I I \right) S - (\gamma_A - \delta) A, \\
 				I_t &= - \gamma_I I + \delta A, \\
@@
 			\begin{equation} \label{eq:diffusion}
-			D (t) = D_0 \left[ 1 - \eta H(t - t_q) \right]
+				D(t) = D_0 \left[ 1 - (1 - \eta) H (t - t_\mathrm{bol}) \right]
 			\end{equation}
@@
 		The average one-way commute in Chicago is about 5 miles \cite{travel}.
-		Thus, we set $D_0$ to the fixed value of $(5 / 0.72)^2 / 4 \Delta t$ where $\Delta t$ is the time step \cite{methods}.
+		Thus we set the diffusion coefficient $D_0$ to the fixed value of $5 / 0.72$ to convert to the spacing used for the discretization.
 
@@
 		The model is solved via finite differences, using the \verb|scikit-fdiff| Python module \cite{nicolas_cellier_2019_3236970}.
+		It is important to note that due to the limitations of the solver, the simulation had to be broken into time blocks where the values of $\omega (t) \beta_A$, $\omega (t) \beta_I$, and $D(t)$ are evaluated for the initial time in each block and held constant.
 
@@
 		\begin{theorem}
-			Let $0 \leq \vec{x}_0 \leq N$ be the initial datum.
+			Let $0 \leq \vec{x}_0 \leq 1$ be the initial datum.
 			Then there exists a unique in time solution $\vec{x}$ of the initial value problem (\ref{eq:model}) without diffusion over $C \subset U \subset \R^4 \times \R^1$ where $C$ is a compact set that contains $(\vec{x}_0, t_0)$.
@@
 		\begin{conjecture}
-			Let $0 \leq \vec{x}_0 \leq N$ the initial datum.
+			Let $0 \leq \vec{x}_0 \leq 1$ the initial datum.
 			Then there exists a unique global in time solution $\vec{x}$ of the initial boundary value problem (\ref{eq:model}).
@@
 		Simulations are performed from March 18, 2020 to June 24, 2020.
-		The time step is $\Delta t = 0.2$ [days].
+		The time step is $\Delta t = 0.2$ [days], chosen to satisfy the convergence condition \cite{pde}.
 
@@
 		The observed data is shown on the left, while the model results are on the right.
-		The data shows that the main progression of the disease is from the lake, to the southwest, then to center-west, then slowly decreases through most of the city.
-		In contrast, the model infected population is an order of magnitude different than the true population after only 15 days, and the difference increases as the simulation continues.
-		Furthermore, the hoped-for diffusion behavior is not present in the model results.
+		Comparing the model results to the data, we see that the model's diffusion mostly misses the diseases west-ward movement between days 15 and 35.
+		Additionally, the model shows significantly fewer cases than are seen in the data (note the difference in the scale of the colorbars).
 
-		In short, the proposed model does \textit{not} reproduce the spatial propagation of the virus.
+		In short, this model does not produce a reasonable reproduction of the spatial spread of COVID-19.
 		\begin{figure}[H]
@@
 \section{Discussion}
-	As mentioned in \S \ref{sec:spread}, the proposed model does not reproduce the spatial propagation of the virus.
+	As mentioned in Section \ref{sec:spread}, the proposed model does not reproduce the spatial propagation of the virus.
 	In this section, we address possible causes of these inaccuracies and discuss proposals for improving the results.
@@
 	Figure \ref{fig:fit} suggests this conclusion, since the growth of the model population only roughly describes the observed data.
+	Note that after Day 20, the model and data begin to diverge substantially.
+	This seems to correspond to the discrepancy in the number of cases seen in the spatial results in Figure \ref{fig:spatial-results}.
 	Likely, the Exposed compartment describing the latent period is necessary, as in \cite{Mammeri+2020+102+113}.
@@
 	A Deceased population may be useful for other reasons, as discussed below.
+
+	The boundaries imposed on the computational domain (Equation \ref{eq:domain}) are likely not sufficient to force the diffusion process to replicate the true virus propagation.
+	Figure \ref{fig:app-pops} shows the $S$ and $A$ populations at the same time points.
+	There is a notable amount of diffusion of the $S$ population out of the boundaries of the city, including into Lake Michigan.
+	Since the $S$ population diffuses into areas not reachable by the $I$ population, the reaction rates as described in Equation \ref{eq:model} are small.
+	This may also explain the differences in case numbers seen in Figure \ref{fig:spatial-results}.
+	It seems a tightening of the bounds of the computational domain is in order.
+	Ideally, we would define the computational domain to be exactly the boundaries of the city.
+	In practice, however, this is infeasible due to the irregularities of the shape.
+	A better alternative is to define a polygon that approximates the shape of the city to be the computational domain.
+	In any case, a new solver will be needed as the \verb|scikit-fdiff| module only allows rectangular domains.
+
+	The optimization method can be improved in two ways.
+	First, by adding a Deceased population we can modify the objective function (Equation \ref{eq:obj}) to compare both cases and deaths, as in \cite{Kevrekidis-2021}.
+	COVID-19 death data is readily available, though requires caution to work with due to reporting inconsistencies.
+	Second, we can perform the optimization directly on the spatial model, as in \cite{Mammeri+2020+102+113}, rather than relying on parameters from a temporal-only model.
 
@@
 			$I$ & Symptomatic infected individuals & [1] \\
+			$J$ & Objective function & N/A \\
 			$R$ & Removed individuals & [1] \\
+			$\Ro$ & Basic reproduction number & [1] \\
 			$S$ & Susceptible individuals & [1] \\
@@
 			$\gamma_j$ & Recovery rate of compartment $j$ & [days\textsuperscript{-1}] \\
+			$\Delta$ & Laplacian operator & N/A \\
 			$\delta$ & Rate of asymptomatic individuals that may develop symptoms & [days\textsuperscript{-1}] \\
+			$\eta$ & Lockdown scale factor & [1] \\
+			$\theta$ & Parameter vector & N/A \\
+			$\rho$ & Spectral radius & N/A \\
+			$\Omega$ & Computational domain & N/A \\
 			$\omega$ & Contacts & [1]

### 2021-12-21 version3/betafun.m

@@ -6,10 +6,15 @@ function b = betafun(p,t_q,t)
     % Create a t_q vector
     t_q = t_q * ones(length(t),1);
 
-    % Calculate beta at time t
-    b = 1 - tanh(2 * (t - t_q));
-    b = b / 2;
-    b = (1 - eta) * b;
-    b = eta + b;
-    b = beta * b;
+%     % Calculate beta at time t
+%     b = 1 - tanh(2 * (t - t_q));
+%     b = b / 2;
+%     b = (1 - eta) * b;
+%     b = eta + b;
+%     b = beta * b;
+    if t < t_q
+        b = beta;
+    else
+        b = eta * beta;
+    end
 end

### 2021-12-21 version3/spatial.py

@@ -5,8 +5,8 @@ import csv
 import matplotlib.pyplot as plt
 
 # Image path
-#impath = '/home/trent/Documents/GitHub/chicago-spatial-covid/version3/images/'
-impath = '/Users/trentgerew/GitHub/chicago-spatial-covid/version3/images/'
+impath = '/home/trent/Documents/GitHub/chicago-spatial-covid/version3/images/'
+#impath = '/Users/trentgerew/GitHub/chicago-spatial-covid/version3/images/'
 
 # LaTeX font for matplotlib
 import matplotlib
@@ -26,7 +26,8 @@ dt = 0.1
 
 # Parameters
 eta = 0.5514 # Lockdown scale factor
-D = (5 / 0.72) ** 2 / (4 * dt) # Diffusivity
+D = 5 / 0.72 # Diffusivity
+#D = (5 / 0.72) ** 2 / (4 * dt) # Diffusivity
 omegaA = 0.1969 # Transmission rate due to A (initial)
 omegaI = 0.3061 # Transmission rate due to I (initial)
 gammaA = 0.3632 # Removal rate of A
@@ -39,7 +40,7 @@ def omegabeta(omega0beta,eta,t,tq):
 
 # Time dependent D function
 def diffusion(D,eta,t,tq):
-    return D * (1 - eta * np.heaviside(t - tq,1))
+    return D * (1 - (1 - eta) * np.heaviside(t - tq,1))
 
 
 ## PDE Model
@@ -55,68 +56,83 @@ model = Model(["D * (dxxs + dyys) - (omegaA * a + omegaI * i) * s",
 # Computational domain
 xpoints = 40
 ypoints = 29
-x = np.linspace(0, xpoints, xpoints)
-y = np.linspace(0, ypoints, ypoints)
+x = np.linspace(0, 100, xpoints)
+y = np.linspace(0, 100, ypoints)
+
+# Check computational stability
+s1 = dt / x[1] ** 2
+s2 = dt / y[1] ** 2
+if not s1 < 0.5 or not s2 < 0.5:
+    print('Bad time step.')
+    sys.exit()
 
 # Hook function for time dependent parameters
 def hook(t, fields):
-    global omegaA, omegaI, D, eta, tq
-    
-    # Define parameters at current time
-    omegaA = diffusion(omegaA,eta,t,tq)
-    omegaI = diffusion(omegaI,eta,t,tq)
-    D = diffusion(D,eta,t,tq)
[…]

### 2022-01-19 response/referee-response.tex
Finalized document
Reorganized folders.
Included referee requests.

\documentclass[11pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{ragged2e}

\pagenumbering{gobble}

%intsitute command
\usepackage{etoolbox}
\makeatletter
\providecommand{\institute}[1]{% add institute to \maketitle
	\apptocmd{\@author}{\end{tabular}
	\par
	\begin{tabular}[t]{c}
		\small \textit{#1}}{}{}
}
\makeatother

%margins
\usepackage[letterpaper, total={6.5in, 9.5in}]{geometry}

\begin{document}
	\vspace{1in}

	\begin{center}
		{\Large Response to Referee Requests for ``Reaction-diffusion spatial modeling of COVID-19 in Chicago''}
	\end{center}

	\vspace{0.25in}

	I would like to thank the anonymous reviewers of my project writeup, ``Reaction-diffusion spatial modeling of COVID-19 in Chicago'' for their comments and critiques.

	I've revised the paper to address multiple mathematical and form errors, including the presentation of the optimization problem as well as the vector notation.
	I have also included phrases to better introduce some of the mathematics.
	For example, the number of contacts and the diffusion coefficient are now introduced before the presentation of the dynamical system.
	The acronyms for the various epidemic models discussed are now spelled out, as requested.

	Additionally, I've introduced language to hopefully clarify the connection between my project and the values of the SoReMo initiative.
	In truth, I was hoping to obtain better mathematical results, in which case I had planned on including visualizations of inequalities in the city overlayed on the model results.
	I had also hoped to be able to offer suggestions for further use of the model to study how the pandemic interacted with social justice; unfortunately I was forced to offer thoughts on the failure of the model to reproduce the data.

	Lastly, in light of the overall failure of the model, I believe the request to include a discussion of the reasonableness of the parameters is beyond the scope of this project.
	Had the model been successful in describing the propagation of the disease, a discussion of the biological implication of the model would certainly be warranted.
	Since that did not happen, I'm not sure how much value such a discussion would add.
	I think the question of reasonableness is already addressed within the current scope of the project, in that the underlying assumptions about the transmission of the virus need to be revisited.

	\vspace{0.25in}

	\begin{FlushRight}
		Trent Gerew

		\today
	\end{FlushRight}
\end{document}

### 2022-01-19 report/version3.tex
Finalized document
Reorganized folders.
Included referee requests.

--- 2021-12-21 version3/version3.tex
@@
 	Then we consider the numerical simulations conducted from March 18 to June 24, 2020.
+	These simulations show that this model may not be sufficient to describe COVID-19 in Chicago.
 	Finally, we discuss shortcomings of the model, and offer some potential solutions.
@@
 	Emerging from Wuhan, China in late 2019, the COVID-19 virus rapidly spread throughout the globe \cite{covid-name}.
-	Since the outbreak, governments everywhere have been desperately trying to contain the pandemic.
+	Since the outbreak, governments have been trying to contain the pandemic.
 	Now, with safe and effective vaccines being distributed, it is hoped the pandemic can be overcome \cite{vaccine}.
@@
 	Most are compartment models, where a population is divided into groups according to the state of individuals.
-	These are generally known as SIR type models.
+	These are generally known as SIR (susceptible-infected-removed) type models.
 	The majority of such models ignore any spatial components.
-	\cite{s+t+spain} employs a SAIR model and includes mobility terms based on the position of cell phones to model the propagation of the disease through Spain.
-	Similarly, \cite{Danon2020.02.12.20022566} incorporates daily movements into an SEIR model.
+	\cite{s+t+spain} employs a SAIR (asymptomatic) model and includes mobility terms based on the position of cell phones to model the propagation of the disease through Spain.
+	Similarly, \cite{Danon2020.02.12.20022566} incorporates daily movements into an SEIR (exposed) model.
 	In a different approach, \cite{giuliani2020modelling} considers a statistical model to handle diffusion of COVID-19 in Italy.
 
-	Here, we follow \cite{Kevrekidis-2021} and \cite{Mammeri+2020+102+113} in building a virus model where the spatial propagation is modeled by a diffusion and the reactions are derived from an SAIR model.
+	The goal of our model is to strike a reasonable balance between representation of the pandemic and the populations involved, and feasibility of data collection and computation.
+
+	To do this, we follow \cite{Kevrekidis-2021} and \cite{Mammeri+2020+102+113} in building a virus model where the spatial propagation is modeled by a diffusion and the reactions are derived from an SAIR model.
 	Specifically, we examine the spread of COVID-19 in the city of Chicago.
@@
 	If it does, we can use the model to assess possible vaccination strategies or reopening scenarios.
+
+	A successful model can also be used to examine how social and economic disparities play a role in the transmission of COVID-19.
+	One can merge race, income, or health-care accessibility data from the Chicago area with the results from the model to ask and hopefully answer questions like the following:
+	Does community structure play a role?
+	How effective are vaccines? (Or what is thwarting them?)
+	Are care facilities properly allocated?
 
@@
 		\end{figure}
+
 		Our model is known as the SAIR model \cite{s+t+spain}, which substitutes the $E$ compartment of the SEIR model by the $A$ compartment.
@@
 			\item
-				We do not distinguish between quarantined, hospitalized, or nursing home populations.
+				We do not distinguish between quarantined, incarcerated, hospitalized, or nursing home populations.
 				Nursing homes are now known to be epicenters of the virus, however the true case rates are notoriously difficult to track \cite{nursing-homes}.
-				Additionally, these three population groups can be assumed to be stationary in nature, and in the case of quarantined and hospitalized, isolated.
+				Additionally, these four population groups can be assumed to be stationary in nature, and in the case of quarantined and hospitalized, isolated.
 				Therefore we ignore their individual contributions.
@@
 		\end{itemize}
-		The goal of this model is to strike a reasonable balance between representation of the pandemic and the populations involved, and feasibility of data collection and computation.
 
@@
 		Lastly, we assume that only members of $S$ and $A$ are mobile.
+
+		The average number of contacts $\omega$ is described in Equation \ref{eq:contacts}.
+		The diffusion coefficient $D$ is assumed to be the same for both $A$ and $I$, and is defined by Equation \ref{eq:diffusion}.
 
@@
 
-		The parameters $\omega_0$, $\beta_A$, and $\beta_I$ are not independently identifiable, so the optimization problem reduces to five parameters $\theta = (\omega_0 \beta_A, \omega_0 \beta_I, \delta, \gamma_A, \gamma_I)$.
-		Given, for $n$ days, the observations $I_\mathrm{obs} (t_i)$, the objective function is the least square function
+		The parameters $\omega_0$, $\beta_A$, and $\beta_I$ are not independently identifiable, so the optimization problem reduces to five parameters $\vec{\theta} = (\omega_0 \beta_A, \omega_0 \beta_I, \delta, \gamma_A, \gamma_I)$.
+		Given the observations $I_\mathrm{obs} (t_i)$ over $n$ days, we have the minimization problem
 		\begin{equation} \label{eq:obj}
-			J (\theta) = \sum_{i=1}^n \left[ I_\mathrm{obs} (t_i) - I (t_i) \right]^2
+			\begin{aligned}
+				\min_\theta \quad & \sum_{i=1}^n \left[ I_\mathrm{obs} (t_i) - I (t_i) \right]^2 \\
+				\textrm{s.t.} \quad & \vec{0} \leq \theta \leq \vec{1}
+			\end{aligned}
 		\end{equation}
-		with constraints $0 \leq \theta \leq 1$.
-		$I (t_i)$ denotes the output of the mathematical model at time $t_i$ computed with parameters $\theta$.
+		$I (t_i)$ denotes the output of the mathematical model at time $t_i$ computed with parameters $\vec{\theta}$.
 
@@
 		This gives the domain
-		\begin{equation} \label{eq:domain}
+		\begin{equation*} \label{eq:domain}
 			\Omega = \left\{ \vec{x} \in \R^2 \mid x \in [0, 40], \ y \in [0, 29] \right\}.
-		\end{equation}
+		\end{equation*}
 
@@
 		We show that we are justified in searching for suitable parameters to solve the model.
-		Let $\vec{x} = (S, A, I, R)$ and $\vec{x}_0 = (S_0, A_0, I_0, R_0)$.
+		Let $\vec{x} = (S, A, I, R)^\intercal$ and $\vec{x}_0 = (S_0, A_0, I_0, R_0)^\intercal$.
 		First we show that the solution of the initial value problem (\ref{eq:model}) exists for the case with no diffusion.
 		\begin{theorem}
-			Let $0 \leq \vec{x}_0 \leq 1$ be the initial datum.
+			Let $\vec{0} \leq \vec{x}_0 \leq \vec{1}$ be the initial datum.
 			Then there exists a unique in time solution $\vec{x}$ of the initial value problem (\ref{eq:model}) without diffusion over $C \subset U \subset \R^4 \times \R^1$ where $C$ is a compact set that contains $(\vec{x}_0, t_0)$.
@@
 		\end{proof}
-		\noindent We do provide a proof for the existence of solutions to the initial boundary value problem (\ref{eq:model}) with diffusion terms, but we assume it exists.
+		\noindent We do not provide a proof for the existence of solutions to the initial boundary value problem (\ref{eq:model}) with diffusion terms, but we assume it exists.
 		\begin{conjecture}
-			Let $0 \leq \vec{x}_0 \leq 1$ the initial datum.
+			Let $\vec{0} \leq \vec{x}_0 \leq \vec{1}$ the initial datum.
 			Then there exists a unique global in time solution $\vec{x}$ of the initial boundary value problem (\ref{eq:model}).
@@
 			\end{align*}
+			Then $\mat{F}$ and $\mat{V}$ are the Jacobians of $\mathcal{F}$ and $\mathcal{V}$ respectively evaluated at the disease free equilibrium.
 			Thus, $\Ro = \rho (\mat{F} \mat{V}^{-1})$ of the next generation matrix
@@
 			So,
-			\begin{equation} \label{eq:ro}
+			\begin{equation*} \label{eq:ro}
 				\Ro =  \frac{S_0 \omega_0}{\gamma_A + \delta}  \left( \beta_A + \beta_I \frac{\delta}{\gamma_I} \right).
-			\end{equation}
+			\end{equation*}
 		\end{definition}
@@
 					$\eta$ & lockdown scale factor & 0.5514 & 1 \\
-					$\delta$ & symptom onset rate & 0.0939 & days\textsuperscript{-1} \\
 					$\gamma_A$ & removal rate of $A$ & 0.3632 & days\textsuperscript{-1} \\
 					$\gamma_I$ & removal rate of $I$ & 0.1385 & days\textsuperscript{-1} \\
+					$\delta$ & symptom onset rate & 0.0939 & days\textsuperscript{-1} \\
 					\hline
@@
 
+	\vspace{0.5cm}
+
+	\noindent \textbf{Acknowledgments:} The author would like to thank the anonymous reviews for their comments.
+
 \section*{Appendix}
@@
 			$D$ & Diffusivity & \\
+			$\mathcal{F}$ & Rate of appearance of new infections & [days\textsuperscript{-1}] \\
+			$\mat{F}$ & Jacobian of $\mathcal{F}$ evaluated at the disease-free equilibrium & N/A \\
 			$I$ & Symptomatic infected individuals & [1] \\
-			$J$ & Objective function & N/A \\
 			$R$ & Removed individuals & [1] \\
@@
 			$S$ & Susceptible individuals & [1] \\
+			$\mathcal{V}$ & Rate of transfer of individuals & [days\textsuperscript{-1}] \\
+			$\mat{V}$ & Jacobian of $\mathcal{V}$ evaluated at the disease-free equilibrium & N/A \\
 			\textbf{Greek symbols} & & \\

### 2022-01-19 version3/no-intervention.py

@@ -1,118 +1,0 @@
-# Imports
-from skfdiff import Model, Simulation
-import numpy as np
-import csv
-import matplotlib.pyplot as plt
-
-# LaTeX font for matplotlib
-import matplotlib
-matplotlib.rcParams['mathtext.fontset'] = 'stix'
-matplotlib.rcParams['font.family'] = 'STIXGeneral'
-
-
-# Time points
-tq = 11 # Lockdown point tq = (tbol + teol) / 5
-tmid = 35 # Test point in the middle of fitting
-tfin = 99 # Last day of fitting
-
-# Parameters
-eta = 0.5514 # Lockdown scale factor
-D = 1.105 # Diffusivity
-omegaA = 0.1969 # Transmission rate due to A (initial)
-omegaI = 0.3061 # Transmission rate due to I (initial)
-gammaA = 0.3632 # Removal rate of A
-gammaI = 0.1385 # Removal rate of I
-l = 0.0939 # Symptom onset rate
-
-
-## PDE Model
-
-# Model setup
-model = Model(["D * (dxxs + dyys) - (omegaA * a + omegaI * i) * s",
-               "D * (dxxa + dyya) + (omegaA * a + omegaI * i) * s - (gammaA - l) * a",
-               "- gammaI * i + l * a"], # Governing equations
-              ["s(x, y)", "a(x, y)", "i(x, y)"], # Population variables
-              parameters=["D", "omegaA", "omegaI", "gammaA", "gammaI", "l"], # Parameters
-              boundary_conditions="noflux") # No flux boundary condition (homogeneous Neumann)
-
-# Computational domain
-L = 40 # Space points
-x = y = np.linspace(0, L, L) # Domain space
-
-
-## Simulation 1: t=0 -> t=tq
-
-# Import the initial populations
-s1 = np.genfromtxt('initialS.csv', dtype=float, delimiter=",")
-a1 = np.genfromtxt('initialA.csv', dtype=float, delimiter=",")
-i1 = np.genfromtxt('initialI.csv', dtype=float, delimiter=",")
-
-# Populate the model fields
-initial_fields1 = model.Fields(x=x, y=y, s=s1, a=a1, i=i1, D=D, omegaA=omegaA, omegaI=omegaI, gammaA=gammaA, gammaI=gammaI, l=l)
-
-# Create and run the simulation
[…]
